### 2020-02-19 paper.tex
Paper

%% This is an article based on superfri.cls class file of
%% ``Supercomputing Frontiers and Innovations. An International Journal''
%% http://superfri.org/.

\documentclass{superfri}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}

\usepackage{array}
\usepackage{todonotes}
\usepackage[capitalize]{cleveref}
\Crefname{lstlisting}{Listing}{Listings}
\usepackage{subcaption}
\usepackage{comment}
\usepackage{float}
\usepackage{listings}
\usepackage{textcomp} %for upquote=true (straight quotes)
\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
\definecolor{pred}{rgb}{0.9,0,0}
\definecolor{pgrey}{rgb}{0.46,0.45,0.48}
\lstset{
    basicstyle=\ttfamily\footnotesize,
    numberstyle=\ttfamily\tiny,
    frame=single,
    numbers=left,
    numbersep=1em,
    xleftmargin=2em,
    language=XML, % to avoid type being blue
    breaklines=true,
    breakatwhitespace=true,
    postbreak=\hbox{$\hookrightarrow$ },
    showstringspaces=false,
    tabsize=2,
    commentstyle=\color{pgreen},
    keywordstyle=\color{pblue},
    stringstyle=\color{pred},
    moredelim=[il][\textcolor{pgrey}]{\[\]},
    moredelim=[is][\textcolor{pgrey}]{\%\%}{\%\%},
  columns=fullflexible,
    upquote=true
}

\newcommand{\jk}[1]{\todo[inline]{JK: #1}}
\newcommand{\lr}[1]{\textcolor{cyan}{LR: #1}}
\newcommand{\hh}[1]{\textcolor{blue}{HuangHua: #1}}
\newcommand{\bnl}[1]{\textcolor{green}{BNL: #1}}
\pagestyle{plain}

% ------------
\bibliographystyle{superfri}

\begin{document}

\author{Julian M. Kunkel\footnote{\label{uread}University of Reading, Reading, United Kingdom}\orcidID{0000-0002-6915-1179} \and Luciana R. Pedro\footnoteref{uread}\orcidID{0000-0001-8365-6264} \and \textcolor{red}{Please add yourself...}
}

\title{Potential of I/O-Aware Workflows in Climate and Weather}

\maketitle{}

\begin{abstract}
The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other and the needs of the end-to-end user.

As climate and weather workflows become increasingly complex and blended beyond data centres while, at the same time, storage hierarchies become deeper, the community investigates ways to reorganise storage access to utilise such systems fully.

The key contributions of this paper are:
1) we sketch the vision of an integrated data-driven approach and discuss the challenges and implications of this strategy, and 2) we illustrate architecture and roadmap that allows the seamless integration into current climate and weather workflows.
The tools employed here to achieve an extended workflow are Cylc, XIOS, DDN IME, and ESDM.

We believe workflows composed of data, computing, and communication-intensive tasks should drive the interfaces and hardware configurations to best support the programming models.
The changes proposed here increase the opportunity of implementations for smarter scheduling of computing and storage in heterogeneous storage systems.

\keywords{workflow, climate, weather, heterogeneous storage, data-driven}
\end{abstract}

% -----------------------------------------------------------------------

\section*{Introduction}
\label{sec:intro}

High-Performance Computing (HPC) harnesses the fastest available hardware components to enable the execution of tightly coupled applications from science and industry.
Typical use-cases include numerical simulation of physical systems and analysis of large-scale observational data.
In the domain of climate and weather, there is a considerable demand for the orchestration of ensembles of simulation models and the generation of data products.
A weather forecast service such as the operational weather forecast workflow in Met Office writes around 200TB every day and reads around 600\,TB.
%For a typical Met Office Unified Model (UM) from the Coupled Model Intercomparison Project (CMIP) run, some %rough estimates of data volumes are:
%\begin{itemize}
%\item N96-ORCA1 -- 200Gb per model year
%\item N216-ORCA025 -- 700Gb per model year
%\end{itemize}
%Historical data spans over 170 years, generating 34Tb for low resolution and 119Tb for high resolution, while %projections will generate approximately half of this.
In total, at the Met Office, on average 1.5\,PBs and 14\,PBs per day are written and read, respectively, for all climate and weather forecasts across all HPC clusters.
%The above supports the view of how important data management already is and these numbers will grow with resolution and more ensembles.
%Existing climate experiments do ensemble experiments sequentially as few institutions have resources for parallel climate ensembles.
%That's an exascale use case requiring not only a big machine but the ability to schedule big jobs.

Based on their needs, the HPC community has developed a software ecosystem that supports climate and weather scientists to execute their large-scale workflows.
While the current advances correspond to a big leap forward, many processes still require experts.
For example, porting a workflow from one system to another still requires adjusting runtime parameters of applications and deciding on how data is managed.

%\bnl{I don't believe people port workflows, they port simulations, and they write and move data ... we want to be able to port workflows ... }
%\jk{I believe they do port aspects of workflows as scripts must be adjusted (where is data and such). Cylc team to reply}
%\lr{It's not clear to me the difference between workflows and simulations. I can imagine that a simulation is a subset of a workflow, right? So, jobs, in terms of computer science?}

Since performance is of crucial importance to large-scale workflows, careful attention must be paid to exploit the system characteristics of the target supercomputer.
A specific supercomputer typically requires substantial changes to the workflow to tailor it to the particular machine and obtain the best performance.
For instance, a data-driven workflow may benefit from using a heterogeneous set of computing and storage technology at the same time.

Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
As the complexity of systems expands and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise resource usage in workflows.
In many cases, modifications are not performed because: 1) They are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) Users are not aware of the potential of the complex system; and, 3) because it is a one-time explorative workflow.
In this paper, we illustrate how knowing the Input/Output (I/O) characteristics of workflow tasks and overall experimental design helps to optimise the execution of climate and weather workflows.
Exploiting this information automatically may increase the performance, throughput and cost-efficiency of the systems, providing an incentive to users and data-centres that cannot be neglected any longer.
Our approach intends to reduce the burden on researchers and, at the same time, optimise the decisions about jobs running on HPC systems.

This paper is structured as follows:
First, we describe the software stack involved in executing workflows in climate and weather in \Cref{sec:workflows}.
Next, the vision for including knowledge about data requirements and characteristics is sketched in  \Cref{sec:vision} outlining the potential benefit the automatic exploitation may bring.
Our design based on existing components in climate and weather is described in \Cref{sec:design}.
The paper is concluded in \Cref{sec:conclusions}.

\section{Workflows in Climate/Weather}
\label{sec:workflows}

In this section, we describe how workflows are executed in a typical software stack and the typical IO stack involved in running a single application.

\subsection{Cylc}

Cylc~\cite{8675433} is in charge of executing and monitoring cyclic workflows in which each step is submitted to the batch scheduler of a data centre.
%Cylc development began in 2008 in response to deficiencies of existing NWP workflow schedulers related to the creation of a single never-ending workflow rather than a never-ending sequence of single-cycle workflows.
%With the current techniques, the evolution of the workflow could be a significant impediment
With Cylc, tasks from multiple cycles may be able to run concurrently without violating dependencies preventing the issue of delays that cause one cycle to run into another.
Cylc was written in Python and built around a new scheduling algorithm that can manage infinite workflows of cycling tasks without a sequential cycle loop.
At any point during workflow execution, only the dependence between the individual tasks matters, regardless of their particular cycle points.

Consider the Cylc workflow for a toy monthly cycling workflow in \cref{pic/cylc1}.
In this workflow, an atmospheric model (labelled as \textbf{model} in the figure) simulates the physics from a current state to predict the future, for example, a month later.
In climate research, this process is repeated in the model to simulate years into the future.
Once the simulation of any month is computed, this task completes and implicitly the data for this month becomes available and can now be analysed.
In this workflow, the \textbf{model} task is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks, all specified as a workflow in a Cylc configuration file (\texttt{conf.rc}).

\fig{width=0.6\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file}

The information Cylc uses to control a given workflow is the task dependency.
The scientists make the configuration about parallelism and data location in the script file that is executed for each specific task.
%Decisions about storage must take into consideration the system architecture and the relevance of heterogeneous infrastructure in which the workflow will run.

\subsection{Workflow Execution}

While Cylc is directing the execution of workflows, several components are involved in the execution.
The software stack involved in executing a general workflow is depicted in \cref{pic/stages}.
%As a preliminary, the user has to specify a configuration representing the Cylc workflow and provide scripts for the individual tasks/applications to be executed.
In the following, each stage of the execution is further described.

\fig{width=0.7\textwidth}{pic/stages}{Software stack and stages of execution}

\begin{enumerate}

  \item \textbf{Scientist} specifies the workflow and provides scripts for each task.
  After that, the user enacts Cylc to start the workflow.

  \item \textbf{Cylc} parses the workflow configuration file, generates tasks dependencies, defines a schedule for the execution, and monitors the progress of the workflow.
  Once a task can be executed (dependencies are fulfilled), the workflow engine submits a \textit{job script} for the workload manager with the required metadata that will run the Cylc task script.

  \item \textbf{Workload Manager} such as Slurm~\cite{Jette02slurm:simple} is responsible to allocate compute resources to a batch job and perform the job scheduling.
  In our case, it now queues the job that represent the Cylc task and plans its execution considering the scheduling policy of the data centre.
  Once the job is scheduled to be dispatched, i.e., resources are available, and the job priority is the highest, it is started on the supercomputer.

  \item \textbf{Job} provides the environment with the resources it runs the user-provided script on one of the nodes allocated for the job.
  The job sets environment variables containing information about the environment of the batch job, e.g., the compute nodes allocated, and then enacts the Cylc provided script on one node.

  \item \textbf{Script} executes the shell commands which is one or multiple (potentially parallel) applications to run sequentially.
  During the creation of the script, Cylc has included variables that describe the task in the workflow.
  The user script may use commands to create a filename considering the cycle of Cylc.
  The information is typically fed into the application(s) representing the task, and so defines the storage location.

  \item \textbf{Application} are executed taking the generated filenames set by the script.
\end{enumerate}

\subsection{I/O Stack of a Parallel Application}

\fig{width=0.2\textwidth}{pic/layers-xios}{I/O path for an MPI-parallel application}

% In our example, we assume the application is parallelised for performance reasons using MPI.
% Not just for performance reasons, it's about making enough memory too, so removed qualification.

Climate applications may have complex I/O stacks, as can be seen in \cref{pic/layers-xios}.
In this case, we assume the application uses XIOS~\cite{xios} which is providing domain-specific semantics to climate and weather.
It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF4~\cite{netcdf} to store data as a file.

Under the hood, NetCDF4 uses the HDF5 API and file format.
Internally, HDF5 uses MPI and its data types to specify the nature of the data stored.
Finally, data is stored on a parallel file system like Lustre which, on the server-side, stores data in a local file system on block devices such as SSDs and HDDs.

Different applications involved in a workflow may use different I/O stacks to store their outputs.
Naturally, the application which uses previously generated data as its inputs must use a compatible API to read the specific data format.
In \cref{pic/layers-xios}, for example, XIOS may perform parallel I/O via the NetCDF4 API, allowing subsequent processes to read data directly using NetCDF4.
Within the ESiWACE project\footnote{\url{https://www.esiwace.eu/}}, we are developing the Earth System Data Middleware (ESDM)\cite{esdm} to allow applications with this kind of software stack to exploit heterogeneous storage resources in a data centre.
The goal of ESDM is to provide parallel I/O for parallel applications, with advanced features to optimise subsequent read accesses.
Implemented with a standalone API, it also provides NetCDF integration allowing usage in existing applications.

\subsection{Data Center Infrastructure}

Data centres are providing an infrastructure consisting of computing and storage devices with different characteristics making them more efficient for specific tasks and satisfying the needs for different workflows.
Take, for example, the supercomputer Mistral at DKRZ that consists of 3,321 nodes\footnote{\url{https://www.dkrz.de/up/systems/mistral}}.
It offers two types of compute nodes equipped with different CPUs and a range of GPU nodes.
Each node provides an SSD for local storage, and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
Individual users and projects are mapped to one file system explicitly, and users can access it with work or scratch semantics:
While data is kept on the work file system indefinitely, available space is limited by a quota.
The scratch file system allows storing more data, but data is automatically purged after some time.

Future centres are expected to have even more heterogeneity. A variety of accelerators (GPU, TPU, FPGAs), active storage, in-memory, and in-network computing technologies will provide storage and processing capabilities.
\cref{pic/system} shows such a system with a focus on computation and storage.
Some of these technologies might be local to specific compute nodes or globally available.
Depending on the need, the storage characteristics range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and cheap storage for long-term archival (tape).
Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy. % (\url{https://www.ddn.com/products/ime-flash-native-data-cache/})
The tasks within any given workflow could benefit from utilizing different combinations of storage and computing infrastructure.

\fig{width=0.6\columnwidth}{pic/system}{Example of an heterogeneous HPC landscape}

\subsection{Data Management}
\label{sec:datamanagement}

Usually, the scripts representing tasks define how the data is placed on an available storage system.
What happens in many current workflows is that they ignore the benefits of using multiple file systems concurrently and the data locality between tasks to co-locating them.
On top of that, in the current state-of-the-art, scientists optimise the available storage resources intuitively and compile the information about this decision-making process manually.

If a user knows the workflow and the system characteristics, s/he can optimise data placement decisions.
Consider, for instance, the situation where each computing node has access to three file systems: a fast \texttt{scratch} file system on which data may reside only for a week, a slower \texttt{work} file system, and a \texttt{local} file system.
Most current workflows utilize \texttt{work} and \texttt{scratch} systems.
When a task is set to run, the corresponding dataset would be moved from \texttt{work} to \texttt{scratch}, processed, and the resulting dataset transferred back to \texttt{work}.
If the \texttt{scratch} filesystem reached its capacity, the datasets would be moved back to \texttt{work}, and the task would continue running until it is finished, but it might be inefficient. %\texttt{local} is not used.
In this situation, there are many obvious opportunities to utilise data migration to optimize performance, and many possible optimization criteria (e.g., costs).
However, with a multitude of file systems which differ at each data centre, such optimisations would be difficult to achieve manually for users.
Policy-driven systems and burst-buffers perform such optimizations automatically to some extent, however, as they lack information about the workflow, they cannot optimize workflows perfectly.

\section{State-of-the-Art}

Related work can be categorized into:
1) technology that exploits heterogeneous storage environments and supports user-directed policies;
2) solutions for workflow processing.

\textbf{Technology.}
Manual tiering requires the user or application to control the data placement, i.e., storing data (typically in the form of files on a particular storage system) and, usually, moving data between storage by scripts.
One limitation of such an approach is that decisions about how data are mapped and packaged into files are made by the producing application, and cannot be changed without manual intervention by a downstream application.

Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy.
They can be integrated into hardware capabilities such as DDN IME~\cite{BODIAIFSFI19} or pure software solutions.
A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
File systems and data management software such as Spectrum Scale, HPSS, BeeGFS, and Lustre (e.g., using the progressive file layouts feature) provide a hierarchical storage management allowing to store data on different storage technology according to such administrator-provided policies.
However, the semantical information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.

The storage community had also adjusted various higher-level software to support storage tiering on top of several storage systems, for instance, ADIOS provides in-memory staging that had been exploited by applications~\cite{slawinska2013maya}.
Hermes~\cite{kougkas2020acceleration}, provides a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
iRODS~\cite{rajasekar2010irods}, is the rule-oriented data system that allows scientists to organize data into shareable collections and provide metadata such as information about data life cycle.
The system provides several patterns for workflows that consider data locality and data migration/replication.

Finally, there have also been extensions to batch schedulers to perform data staging for utilizing node-local storage, for example, NORNS as an extension to Slurm~\cite{miranda2019norns}.

\smallskip

\textbf{Workflows. }
A good overview of the flavours of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
The article states a SWfMS "should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources".
%"An abstract scientific workflow is defined by the functionality of each activity (or sub-workflow) and data dependencies between activities (or sub-workflows)"
Existing solutions recognize challenges in data variety (formats of the input data), the opportunity to optimize the schedule by moving code to data, and to consider capacity of available data storage.
There have been also solutions that allow to specify the data dependencies for tasks.
The execution engine Dryad~\cite{isard2007dryad}, for example, allows to transfer data between tasks via files or directly using TCP connections and attempts to schedule tasks on the same nodes or racks.
Swift/T is a scripting language for describing dataflow processing allowing to execute ensembles of applications~\cite{ozik2016desktop}.
Recent improvements aim to migrate data to a local cache allowing to exploit locality, e.g., in \cite{dai2018cross}, information about locality is proposed to be stored in extended attributes.
In \cite{TUIBIHWLSC19}, an approach was presented to monitor and analyze IO behavior of HPC workflows.

Various early research covers grid workflows and lately cloud use cases and attempts to maximize data locality in that respect.
Economic factors (including storage costs) for workflow execution are discussed in \cite{alkhanak2016cost}.
In \cite{deelman2019role}, the authors discuss the role of Machine Learning for workflow execution and elaborate an the general potential for resource provisioning such as optimization of runtime parameters, data movements, and hierarchical storage.
An ML model is introduced in \cite{subedi2019leveraging} that stages data for in-situ analysis by exploiting the access patterns.

%TRIO: Burst Buffer Based I/O Orchestration. In: 2015 IEEE International conference on cluster computing, Chicago, IL, USA, 8–11 September 2015, pp. 194–203. Washington, DC: IEEE.
%Chervenak, A.L., Smith, D.E., Chen, W., Deelman, E.:
%Integrating policy with scientific workflow management
%for data-intensive applications. In: Supercomputing (SC)
%Companion: High Performance Computing, Networking
%Storage and Analysis, pp. 140–149 (2012)

% https://www.sciencedirect.com/science/article/pii/S0167739X14002519
%Workflow engine for cloud \cite{pandey2011workflow}
%staging data to an available location just somehow.
%"storage service provides a temporary repository for application files that is, input files that are required for task execution, and output files that are he result of execution.""

Workflow systems can also be utilized specifically for the purpose of reproducing scientific results, i.e., recomputing the results.
Those scalable workflow solutions typically utilize a container solution to allow execution in an arbitrary software environment.
Popper~\cite{jimenez2017popper}, Snakemake~\cite{bts480}, and Nextflow~\cite{Nextflow} provide a language to specify workflows and to execute them.
Snakemake is interesting as it supports to define and infer input and output filenames.

While various aspects of our vision have been addressed individually by related work for different domains, the high level of abstraction that we aim for and the potential it unleashes goes beyond the capabilities of existing approaches.

\section{Vision for I/O-Aware Workflows}
\label{sec:vision}

Nowadays, in order to run a job in an HPC environment efficiently, researchers have to develop profound knowledge, not only about their workflow, which is expected, but also about decisions regarding storage, communication, and computing including considerations regarding cost-efficiency of their operations.
However, applied scientists should not spend much time understanding hardware characteristics and operative knowledge of running a data centre, but using their time to develop their work and just collect and analyse their results.

We aim for achieving an automatic and dynamic mapping of I/O resources to workflows.
Once we have an automated decision about where the job will run and how the storage will be managed, scientists can then run their workflows on any system without further modification and even without previous knowledge about the system architecture.

There are several approaches to implement the technology for the vision proposed in this work, and changes are needed for software components to realise it.
In the next sections, we will discuss a specific design for our transitional roadmap considering climate and weather workflows and tools scientists from this field already use in their routine research.

Our vision for I/O-aware workflows requires two additional pieces of information.
Firstly, the user must augment the workflow description with information about I/O requirements and explicitly annotate dependencies to datasets.
Secondly, the data centre (or expert user) must provide details about the storage architecture.

\subsection{System Information}

While many optimisations are possible once an abstraction is in place, the improvements we discuss here are related to the life cycle of datasets and the placement of such datasets into specific storage according to system performance characteristics and the workflow specification.
To achieve that, the system information shall comprise of all available storage systems, the system topology, and details of each of the required components.
Simplified and complex models of the components can be included to approximate expected performance for specific I/O patterns.
It is expected that the data centre (or expert users) can create such a configuration file, e.g., by using vendor-provided information or by executing benchmarks.
With this information, a scheduler can make data placement, transformation, and migration decisions for individual datasets during their life cycle.
This separation of concerns allows us to abstract from the workflow what is essential and what a system should optimise to ensure smart usage of available resources.

\subsection{Extended Workflow Description}

In general, climate and weather workflows allow specifying tasks and the dependencies among them.
We aim to enhance the current information with characteristics for input, and output, i.e., the datasets.
An example workflow containing input datasets and (intermediate) products is illustrated in \cref{pic/workflow-s3-n}.
Nodes represent tasks or data, and arrows indicate the dependencies.
In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
For each new cycle, the \texttt{checkpoint} from the previous cycle, called here \textbf{Product\,1}, is used as input to starting the next cycle.
Most of the workflow can run automatically, except for the manual quality control of the products and the final data usage of Product\,3.
This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
In the approach proposed in this work, each task is annotated with the required input datasets and the generated products must include metadata such as data life cycle information, the value of data, and how long it should be kept.
The idea here is to embrace the concept that tasks dependencies are really imposed by datasets dependencies.

\fig{width=0.8\columnwidth}{pic/workflow-s3-n}{Example of a high-level workflow with tasks and data dependencies}

%The access to I/O information before the workflow runs can be used to optimise where jobs will be executed and how datasets will be stored during the workflow.
%This information today is static and manual, leaving open opportunities for automating and optimisations.
%Once we insert the datasets explicitly in the workflow, it is easier to see that many optimisations regarding storage can be made.

\subsection{Smarter I/O Scheduling}

The abstraction and automation of the I/O inside a workflow allows a runtime system to improve data placement and apply data reduction on heterogeneous storage systems.
Taking into consideration the architecture and workflow information, a smarter schedule can now be realized by exploiting the additional information.
Value and priority can influence fault-tolerance strategies and imply the quality of service for performance and availability.
Aspects like data reproducibility (can it be recomputed easily), type of the experiment (test, production), and runtime constraints for the overall and potential workflow could allow reducing costs and, hence, increasing scientific output.
Next, we outline the two core strategies and potential the proposed vision can bring to improve current workflows.

% \subsubsection{Data Placement}

\paragraph{Strategy: Data Placement.}

Data placement encompasses all data movement related activities such as transfer, staging, replication, space allocation and de-allocation, registering and unregistering metadata, locating and retrieving data\footnote{https://www.igi-global.com/dictionary/data-aware-distributed-batch-scheduling/6782}.
The general idea is to host a data set on the storage system that is most favorable in terms of performance, cost-effectiveness, availability for the access pattern observed by the workflow.
It requires to optimize data locality, where locality is twofold, spatial and temporal on a variety of characteristics.
For optimising data placement, there are several approaches:%: \textbf{in-memory data}, \textbf{data allocation}, \textbf{data migration}, and \textbf{data replication}.

\begin{description}

\item[Data Allocation] is the assignment of a specific area of an available storage system to particular data.
In current workflows, usually, the user has a script for each task defining the filenames with a prefix that places datasets generated by the same task into a specific storage\footnote{Complicated scripts would have allowed changing the storage type depending on the cycle. Still, it is a significant burden to the user.}.
Because there is one script responsible for generating the configuration, the decision in which directory the dataset will be stored is somewhat fixed.
Such configuration is done manually and with restricted information about the system architecture.
It would be interesting to explore storage options for the datasets and, e.g., having datasets from different cycles placed at different storage systems.
For instance, in \cref{pic/workflow-s3-n}, alternating the storage location for \textbf{Product\,2} into two scratch file systems is something that would be a simple job for an I/O-aware scheduler.
However, currently, that implies having at least two scripts for that task with information about the different storage placement.
%Next, we exemplify three situations in which data placement can be optimised.

\item[Data Migration] is the process of transferring data from one storage system to another.
Data movement involves a large overhead, both in terms of latency and energy-efficient computing, hence needs to be considered carefully.
\cref{fig:lifecycle} introduces three possible life cycles for a specific dataset and explains how migrations can be done to improve datasets accessibility.
In \cref{fig:lifecycle1}, the dataset could be stored on the \texttt{local} storage first to avoid congestion on the \texttt{work} file system, then be migrated to \texttt{work} file system where subsequent tasks of the workflow may read it multiple times.
In the end, this dataset might be an intermediate product that can then be deleted.
Alternatively (see \cref{fig:lifecycle2}), the dataset could be stored on the \texttt{scratch} file system immediately and accessed there.
However, that would require subsequent tasks to be placed on the same node where data is now stored, and the last read access must happen before files on \texttt{scratch} are automatically removed.
The last scenario (see \cref{fig:lifecycle3}) presents the case where the dataset is created on \texttt{work} by a task and it is copied to a \texttt{local} node.
This \texttt{local} node allows reading accesses of subsequent tasks which might be beneficial for small random accesses.
However, it would require that the subsequent tasks are placed on the same node where data is now stored.

\begin{figure}[b]
    \begin{minipage}{.33\linewidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{pic/lifecycle-1}
        \subcaption{Local and work file systems}\label{fig:lifecycle1}
    \end{minipage}
    \begin{minipage}{.33\linewidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{pic/lifecycle-2}
        \subcaption{Scratch file system only}\label{fig:lifecycle2}
    \end{minipage}
    \begin{minipage}{.33\linewidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{pic/lifecycle-3}
        \subcaption{Local and work file systems}\label{fig:lifecycle3}
    \end{minipage}
    \vspace{5pt}
    \caption{Alternative life cycles for mapping a dataset to storage and the operations: \textbf{A}llocation, \textbf{M}igration, \textbf{R}eading, and \textbf{D}eleting}
    \label{fig:lifecycle}
\end{figure}

\item[Data Replication] in computing involves sharing information to ensure consistency between redundant resources, such as software or hardware components, to improve reliability, fault-tolerance, or accessibility.
Data might be replicated by enabling the system to rerun parts of the workflow in case of a data loss.
In addition, the system may combine the replication of data by \textbf{transforming} the data into a different representation allowing to achieve better performance for a variety of access patterns.

\item[Direct-Coupling] replaces I/O by communicating data between subsequent steps of a workflow directly without storing intermediate data products on persistent storage.
As an example, in \cref{pic/workflow-s3-n}, the outcome of Task\,1 may be used directly by Task\,2.
To achieve a decoupling between producer and consumer, it may also be kept in memory and cached.
\end{description}

In the design proposed in this work (\cref{sec:design}), we will focus on the data placement strategy.

\paragraph{Strategy: Data Reduction.}

Data reduction reduces the amount of data stored.
We discuss here two potential optimisations: \textbf{data compression}, and \textbf{data recomputation}.

\begin{description}

\item[Data Compression] is the process of encoding information using fewer bits than the original representation.
Knowing the characteristics of data production and usage makes it simpler for scientists to annotate the required precision of data in those workflows.
The storage system can exploit such information by reducing precision of data and automatically picking an appropriate compression algorithm.

%\lr{Towards Decoupling the Selection of Compression Algorithms from Quality Constraints: An Investigation of Lossy Compression Efficiency~\cite{TDTSOCAFQC17}}

\item[Data Recomputation] Climate/weather scientists are trading recomputation with space usage manually.%, in the so-called field space-time trade-off that has been applied in Cryptanalysis for already forty years~\cite{1056220}.
By knowing how to rerun the workflow behind the data creation, a smarter storage system can automatically trade data availability for potential recomputation opportunities to optimise the cost-efficiency of the system.
Intermediate states could be rerun by utilising virtualisation and container technologies.
Consider \cref{pic/workflow-s3-n} again and that, at every K cycles of the workflow, the generated \textbf{Product\,3} (from Cycle 1 to Cycle K) are used in a validation task, called here \textbf{check}.
From the workflow\footnote{The $P_i C_j$ notation represents the Product\,$i$ generated in the Cycle $j$.}, we know that $P_3 C_1$ will be used to construct $P_3 C_2$ and then \textbf{check}.
This dataset will probably be stored somewhere, and it will not be used until the workflow reaches the K-th cycle.
One alternative is to delete it after it was first used and then recompute it when time is right.
The cost of doing that is storing \texttt{checkpoint} and then use it to reconstruct product $P_3 C_1$.
If, for instance, $P_3 C_1$ is a large dataset, \texttt{checkpoint} is small, and computing time is short.
It is easy to see that deleting and recomputing it may improve the costs for running a workflow.
%We need, then, information about datasets volume and computing time to be able to make an accurate decision.
%The bottom line here is: there are situations in which decisions are being made without considering all the available %information and possible scenarios.
This is just an example, and, currently, scientists perform optimisations manually.
\end{description}

\subsection{Benefit}

The benefits of the proposed vision are:

\begin{description}

\item[Abstraction] The user does not have to know the architecture of the target system on which the workflow will run, removing the specialist from the decision-making process.

\item[Optimisation] The workflow will be optimised specifically for the available system infrastructure and extra information about the data.
In particularly, by exposing the heterogeneous architecture, potentially runtime characteristics can be considered.
By using information about the value of data, policies for data management (storage resilience,  recomputation, replication, etc.) can be decided.

\item[Performance-portability] With both abstraction and optimisation, the user can specify the I/O requirements only once for the tasks of a specific workflow, and the workflow can now run on any system without user intervention.
Even more, if the system characteristics change, e.g., it gets upgraded, an additional storage tier becomes available, or if storage degrades, the I/O-aware workflow could automatically adapt and make use of this new environment. % gain

\end{description}

\section{Design}
\label{sec:design}

This section describes our first approach to incrementally extend workflows for climate and weather that realises parts of our vision.
While individual components such as ESDM and Cylc exist, we have not implemented the described scheduler, yet.
To automatically make scheduling decisions, the software stack needs to:

\begin{enumerate}

\item Deliver information about dataset life cycle together with the workflow, and

\item Adapt the resulting workflow, individual scripts and application executions to consider the potential for data placement and migration.

\end{enumerate}

\subsection{System Information}

ESDM is used as IO middleware in the parallel application (in NetCDF or directly) and orchestrates the IO according to a configuration file.
The information about the system is provided using a simplified ESDM configuration file (\texttt{esdm.conf}) provided by the data centre (or expert users).
This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.

In the example presented in \Cref{lst:esdm.conf}, we have three storage targets: two global accessible file systems (\texttt{lustre01} and \texttt{lustre02}), and one local file system in \texttt{/tmp} that can be accessed via the POSIX backend.
Each of them comes with a lightweight performance model and the maximum size of data fragments.
The metadata section (Line 24) utilises here a POSIX interface to store the information about the ESDM objects. Internally, ESDM creates so-called containers and dataset objects to manage data fragments.

\begin{lstlisting}[caption = Example of an ESDM configuration file (\texttt{esdm.conf}), label=lst:esdm.conf]
"backends": [
    {"type": "POSIX", "id": "work1", "target": "/work/lustre01/projectX/",
        "performance-model" : {"latency" : 0.00001, "throughput" : 500000.0},
        "max-threads-per-node" : 8,
        "max-fragment-size" : 104857600,
        "max-global-threads" : 200,
        "accessibility" : "global"
    },
    {"type": "POSIX", "id": "work2", "target": "/work/lustre02/projectX/",
        "performance-model" : {"latency" : 0.00001, "throughput" : 200000.0},
        "max-threads-per-node" : 8,
        "max-fragment-size" : 104857600,
        "max-global-threads" : 200,
        "accessibility" : "global"
    },
    {"type": "POSIX", "id": "tmp", "target": "/tmp/esdm/",
        "performance-model" : {"latency" : 0.00001, "throughput" : 200.0},
        "max-threads-per-node" : 0,
        "max-fragment-size" : 10485760,
        "max-global-threads" : 0,
        "accessibility" : "local"
    }
],
"metadata": {"type": "POSIX",
        "id": "md",
        "target": "./metadata",
        "accessibility" : "global"
    }
\end{lstlisting}

%\hh{Here I have similar question: we need an extended and new configuration file, which includes characteristics of inputs/outputs dataset. As Julian said above, we can not added these items directly into Cylc configuration because Cylc wouldn't understand it. But you see, it's a burden to have another conf file. It is error-prune, and duplicated information may cause inconsistency. I propose to have these information added to the same Cylc configuration file, but in special comments. These comments will be omitted by Cylc, but they can be parsed by ESDM. This is similar for Doxygen comments in C/Java source file, or the leading line of a shell script, etc.}
%\jk{Good suggestion! That absolutely depends how a Cylc file is parsed and analyzed. Keeping it separately gives us the most flexibility and parsing comments could lead to misinterpreting already existing comments and acting wrongly. Still to consider.}
%\lr{I like this idea a lot. We need to talk to the Cylc team and see what they think.}

%\hh{Other requirements of datasets may be: data block size/fragment size, 1-1 or N-1 access mode, IOPS intensive or throughput-intensive, etc. These will also be considered for data placement and migration.}
%\jk{Fragment size/data blocks are handled by ESDM automatically. The access mode and access pattern is something that we could consider to add. However, difficult for scientists to judge}

ESDM manages a pool of threads that should be created per compute node to achieve good performance and delegates the assignment of optimal block sizes to the storage backend.
The number of threads is defined in the configuration file.
As an example, based on the number of Object Storage Targets (OSTs) available to both Lustre systems at DKRZ, performance tests already developed by ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
%; more degrade it.
To clarify the behaviour, ESDM distributes a single dataset across multiple storage devices depending on their characteristics.
Since ESDM also supports several (non-POSIX) storage backends, an application can utilise all available storage systems without any modifications to the code.

%\bnl{This has me asking all sorts of questions: how does one learn the right information to put in the ESDM configuration, e.g. your 200. What do threads ``per node'' mean? What sort of node? Is the ESDM running as an I/O server, or are you assuming it is being used in an IO server etc}
%\lr{Well noticed. For me, ESDM would be responsible for finding those optimal numbers automatically. But then we had this sentence: ``The number of threads is defined in the configuration file.'', meaning that the user has to know. So I don't have an answer to it.}

The configuration file is inquired by an application that utilises ESDM and steers the distribution of data during I/O.
While the current system information and performance model is based on latency and throughput only, it shows that automatic decision making can be made on behalf of the user.

\subsection{Extended Workflow Description}

The user now has to provide information about datasets required for input and the generated output for each Cylc task in a separate file similarly to Cylc's workflow configuration file.
An example of an I/O-workflow configuration file is shown in \Cref{lst:cylc}. In this file, information about Task\,1 is given by example, and we expect the extra information about all tasks in the same file; ultimately, this could be integrated into the workflow specification file of Cylc.

\begin{lstlisting}[caption = External Cylc I/O-workflow configuration file, label=lst:cylc]
[Task 1]
  [[inputs]]
    topography = "/pool/input/app/config/topography.dat"
    checkpoint = "[Task 1].checkpoint$(CYCLE - 1)"
    init       = "/pool/input/app/config/init.dat"

  [[outputs]]
    [[[varA]]] # This is the name of the variable
      pattern = 1 day
      lifetime = 5 years
      type = product
      datatype = float
      size = 100 GB
      precision.absolute_tolerance = 0.1

    [[[checkpoint]]]
      pattern = $(CYCLE)
      lifetime = 7 days
      type = checkpoint
      datatype = float
      dimension = (100,100,100,50)

    [[[log]]]
      type = logfile
      datatype = text
      size = small
\end{lstlisting}

In this example, the workflow file could define a cycle flexibly to be a month or a year according to the Cylc workflow file.
The notation is similar to the specification of Cylc workflows using a nested INI format.
For each task, inputs and outputs are defined.
In the input section, each entry specifies the virtual name that is used by ESDM as a filename inside NetCDF.
Line~3, for example, defines that the filename \texttt{topography} is mapped to a specific input file.
This dataset does not depend on any previous step of the workflow.
The next line specifies that the input filename \texttt{checkpoint} should be mapped to the output of Task~1 checkpoint dataset from the previous cycle (e.g., the checkpoint generated after completed the last year's output).
For the initial cycle, the checkpoint file will be empty, and the application will load the init data.
In the output section, the datasets are annotated with their characteristics more precisely.
For each variable, a pattern defining how frequently the data is output according to the workflow must be provided.
Most data is input and output in the periodicity of the cycle, except for \texttt{varA}, which is output per day regardless of the cycle.
Next, we formally define the expected annotations in all the fields expected in the I/O-workflow configuration file:

\begin{description}[itemsep=0pt]
  \item[Name] A basic name for the field/data generated (Lines: 8, 16, 23). It is extended by a pattern defined in a variable.
  \item[Pattern] The frequency the data is output (Lines: 9, 17).

  \item[Lifetime] How long the data must be retained on storage (if at all) (Lines: 10, 18).

  \item[Type] The class type of data, i.e., checkpoint, diagnostics, temporary (Lines: 11, 19, 24).

  \item[Datatype] The data type of the data (Lines: 12, 20, 25).

  \item[Size] An estimate of the data size (Lines: 13, 26). This field can be inferred if dimension and datatype are provided.

  \item[Dimension] The data dimension (Line 21).

  \item[Accuracy] This includes various characteristics quantifying the required level of data precision; For example, in Line~14.

\end{description}

Note that the user may not be able to provide all required information.
This can be handled by assuming a default safe behavior.
For instance, in case of a missing data precision, data should be retained in original form.
Knowing the dimension or size a priori might be difficult for scientists, e.g., the log file size is unclear.
In this case, the user may insert relative information like small or big, reflecting that any information is better than no information at all.
In future, we will explore how to automatically infer the output volume from the input or by using monitoring.

\subsection{Smarter I/O Scheduling}

From the list of opportunities, we realise data placement and migration in a heterogeneous (multi-storage) environment.
These goals will be achieved via the proposed I/O-aware scheduler, called here EIOS (ESDM I/O Scheduler).
EIOS will make the schedule considering Cylc workflow, and ESDM provided system characteristics.
Components of it are involved in different steps of the workflow and the I/O path.

While Cylc schedules the workflow, EIOS can provide hints about co-locating tasks which provide the opportunity for keeping data in local storage.
Decisions about data locality will not be made for a whole (and potentially big) workflow.
Instead, the system will make decisions by looking ahead to several steps of the workflow, allowing reacting to the observed dynamics of the execution.
We utilise DDN's IME API to pin data in IME and to trigger migrations between IME and a storage backend explicitly.

Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated by modifying the "filename" which is then used by the ESDM-aware application to determine the I/O scheduler.

\subsection{Modified Workflow Execution}

The steps to execute a workflow enriched with IO information and performing smarter scheduling is depicted in \cref{pic/stages-io}.
The suggested alterations can be seen in boxes pointed by red arrows, and the remaining components are the current state-of-the-art for workflows in climate and weather from \cref{pic/stages}.

In the following, we describe the modifications we propose in this vision paper for each component involved in the software stack.

\fig{width=0.7\columnwidth}{pic/stages-io}{Software stack and stages of execution with our I/O-aware scheduler -- EIOS}

\begin{enumerate}

  \item \textbf{Scientist} The end-user now has to provide an additional file that covers the I/O information for each task and slightly changes (discussed above) have to be made to the current scripts.

  \item \textbf{Cylc} EIOS is invoked by Cylc to identify potential optimisations in the schedule before generating the Slurm script.
  %\hh{Cylc has to be extended to call EIOS? How is this achieved?}
  %\lr{We talked to the Cylc team, and they will deal with it, I suppose.}
  %\jk{Modifying Cylc, Cylc team to decide, not yet decided how that will work, Python code.....}

  \item \textbf{\color{red}{EIOS}} The ESDM I/O Scheduler reads the information about the workflow (original and I/O workflow configuration files) and acts depending on the stage of the execution.
  EIOS consists of several subcomponents:
    \begin{itemize}

      \item The high-level scheduler that interfaces with Cylc.

      \item A tool to generate pseudo-file names used by the ESDM-aware applications.

      \item A data management service (not shown on the figure) that migrate and purge data at the end of the life cycle.
    \end{itemize}
  EIOS components use knowledge about the system by parsing the \texttt{esdm.conf} file.
  EIOS may decide that subsequent jobs shall be placed on the same node, reorder the execution of some jobs, and provide information for conducting data migration.

  \item \textbf{Slurm}
  Cylc may now have added an optimisation identified by EIOS, and that is now handled by a modified Slurm.
  Also, if migrations have to be performed, Slurm will administer them according to the specification in the job script.
  %\lr{Already inserted in the idea of the previous sentence.}
  %\lr{It's clear the idea about what's really going on in each part is still not clear.}

  \item \textbf{Job}
  A job might run on the same nodes than a previous job to utilise local storage.
  %\bnl{Any interesting that needs this performance is running on dozens if not hundreds of nodes}

  \item \textbf{Script}
  Filenames are now generated by a replacement command that calls EIOS to create a pseudo filename.
  This filename will encode additional information for ESDM about how to prioritise data placement according to data access.

  \item \textbf{Application}
  The application may either use XIOS, NetCDF with ESDM support or ESDM directly to access datasets.
  Hence, in \cref{pic/layers-xios}, the HDF5 layer is replaced with ESDM.
  ESDM loads the \texttt{esdm.conf} file that contains the information about the available storage backends and their characteristics.
  ESDM extracts the long-term schedule information from the generated pseudo filenames and employs it during the I/O scheduling to optimise the storage considering data locality between tasks.

\end{enumerate}

\section{Conclusions}
\label{sec:conclusions}

Organising the data placement on storage tiers is typically performed manually (and particularly in the domain of climate and weather) by the users or via policies, leading to suboptimal optimisations.
However, the manual optimization and hard-coding of storage locations is non-portable and an error-prone task.
We believe users must be able to express their workflows in an abstract fashion.
By increasing the abstraction level for scientists, not only tedious manual optimisation could be automatised, but also optimization strategies for data placement and data reduction can be harnessed.
With knowledge about the data pattern, the runtime system could generate (near-)optimal execution plans and monitor their execution.
In this work, we describe the general vision and a specific design for the software stack in the domain in climate and weather that we work on in the ESiWACE project.
The changes proposed increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system in a workflow.

\section*{Acknowledgements}

\small
This project is funded by the European Union's Horizon 2020 research and innovation programme under grant agreement No. 823988.

\openaccess

\bibliography{paper}

\end{document}

Machine Learning has been proven effective in replacing human-beings, and it may also adapt the workflow dynamically in case of unexpected situations (the system is down, nodes are unresponsive, etc.). (ref)

By monitoring I/O accesses, we can automatically draft such I/O descriptions on behalf of the user to simplify the specification.

All we need is to run one cycle of the specific workflow, considering that a workflow cycle is executed many times in climate and weather; this is a cheap approach.

From the recorded I/O accesses, we analyse the I/O patterns and dependencies associated with it to generate the configuration file.

Potentially, EIOS may generate an \texttt{esdm.conf} file for the specific application to adjust the mapping of data to available storage targets.

In particular, a system could store data that is of low value or cheap to reproduce on storage with little redundancy or trade storage space by the need to recompute data products.

This scenario was developed to exemplify some of the potential optimisations we envision about \textbf{data placement}, \textbf{migration}, and \textbf{recomputation} in a heterogeneous storage environment.

The user-specific storage locations are a viable option to run a workflow.

The user might even have analysed various scenarios, such as depicted in Section \ref{sec:datamanagement}.

Assuming the user can provide extra information in advance, ESDM can interact in the process before the user gives the data to the Cylc tool, using the gathered information to construct a tentative optimised workflow.

Here, we focus on utilising a heterogeneous storage environment.

Once we have access to the data pattern, the whole workflow can be improved by trading storage space by the need to recompute data products.

With the right level of abstraction in place, potential data optimisations are unleashed in a heterogeneous storage system according to its characteristics and workflow specification.

Exploiting the value and importance of data could improve the placement decisions on heterogeneous storage systems.

Data systems usually do not respect, or have knowledge of, the importance of their inherent data.

%\hh{I like this idea of machine learning in storage system. A lot of analysis has been done on relation of I/O pattern and storage performance. But how to know the I/O pattern, data pattern in advance? Machine Learning can be used to find pattern. These patterns can then be used to schedule storage resources, data placement, data staging/migration, etc. These pattern can be gradually learned from running cycles of workflows over time, such that policies can evolve to have better performance when more and more cycles of workflow complete. }

### 2020-02-19 paper.tex
Minor

@@
 The information Cylc uses to control a given workflow is the task dependency.
-The scientists make the configuration about parallelism and data location in the script file that is executed for each specific task.
+The developers define, in a script file, for each task the parallelism settings and where the data is to be stored.
 %Decisions about storage must take into consideration the system architecture and the relevance of heterogeneous infrastructure in which the workflow will run.
@@
 We aim for achieving an automatic and dynamic mapping of I/O resources to workflows.
-Once we have an automated decision about where the job will run and how the storage will be managed, scientists can then run their workflows on any system without further modification and even without previous knowledge about the system architecture.
+Once we have an automated decision about where the job will run and how the storage will be managed, scientists can then reuse their workflow specification on any system without further modification and even without previous knowledge about the system architecture.
 
@@
 In future, we will explore how to automatically infer the output volume from the input or by using monitoring.
+By allowing to run using an empty I/O workflow specification and monitoring I/O accesses for one cycle, we can propose an I/O description to the user to simplify the specification.
 
@@
 
-By monitoring I/O accesses, we can automatically draft such I/O descriptions on behalf of the user to simplify the specification.
-
-All we need is to run one cycle of the specific workflow, considering that a workflow cycle is executed many times in climate and weather; this is a cheap approach.
-
 From the recorded I/O accesses, we analyse the I/O patterns and dependencies associated with it to generate the configuration file.

### 2020-02-20 paper.tex
Minor changes (Dave) and review of related work / conclusion

@@
 \usepackage{float}
+\usepackage[utf8]{inputenc}
 \usepackage{listings}
@@
   During the creation of the script, Cylc has included variables that describe the task in the workflow.
-  The user script may use commands to create a filename considering the cycle of Cylc.
+  The user script may use commands to create a filename considering the cycle of Cylc or store data in a workflow-specific shared directory.
   The information is typically fed into the application(s) representing the task, and so defines the storage location.
@@
 \textbf{Technology.}
-Manual tiering requires the user or application to control the data placement, i.e., storing data (typically in the form of files on a particular storage system) and, usually, moving data between storage by scripts.
+Manual tiering requires the user or application to control the data placement, i.e., storing data typically in the form of files on a particular storage system and, usually, moving data between storage by scripts.
 One limitation of such an approach is that decisions about how data are mapped and packaged into files are made by the producing application, and cannot be changed without manual intervention by a downstream application.
@@
 A good overview of the flavours of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
-The article states a SWfMS "should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources".
+The article states a SWfMS \textit{“should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources”}.
 %"An abstract scientific workflow is defined by the functionality of each activity (or sub-workflow) and data dependencies between activities (or sub-workflows)"
-Existing solutions recognize challenges in data variety (formats of the input data), the opportunity to optimize the schedule by moving code to data, and to consider capacity of available data storage.
-There have been also solutions that allow to specify the data dependencies for tasks.
+Existing solutions recognize challenges in data variety (formats of the input data), the opportunity to optimize the schedule by moving code to data, and may consider capacity of available data storage.
+There have been solutions that allow to specify the data dependencies for tasks.
 The execution engine Dryad~\cite{isard2007dryad}, for example, allows to transfer data between tasks via files or directly using TCP connections and attempts to schedule tasks on the same nodes or racks.
@@
 
-Various early research covers grid workflows and lately cloud use cases and attempts to maximize data locality in that respect.
+Various early research in grid workflows and lately cloud use cases attempts to maximize data locality in that respect.
 Economic factors (including storage costs) for workflow execution are discussed in \cite{alkhanak2016cost}.
@@
 
-Organising the data placement on storage tiers is typically performed manually (and particularly in the domain of climate and weather) by the users or via policies, leading to suboptimal optimisations.
-However, the manual optimization and hard-coding of storage locations is non-portable and an error-prone task.
+Organising the data placement on storage tiers is particularly in the domain of climate and weather performed by the users or via policies, leading often to suboptimal decisions.
+Additionally, the manual optimization and hard-coding of storage locations is non-portable and an error-prone task.
 We believe users must be able to express their workflows in an abstract fashion.
-By increasing the abstraction level for scientists, not only tedious manual optimisation could be automatised, but also optimization strategies for data placement and data reduction can be harnessed.
-With knowledge about the data pattern, the runtime system could generate (near-)optimal execution plans and monitor their execution.
+By increasing the abstraction level for scientists, not only tedious manual optimisation could be automatised, but also strategies for data placement and data reduction can be harnessed.
+With knowledge about the data pattern, the runtime system could generate optimized execution plans and monitor their execution.
 In this work, we describe the general vision and a specific design for the software stack in the domain in climate and weather that we work on in the ESiWACE project.
@@
 This project is funded by the European Union's Horizon 2020 research and innovation programme under grant agreement No. 823988.
+We thank our collaborator Bryan Lawrence for their input to this article.
 
@@
 }
+
+CYLC_SUITE_SHARE_DIR
+Typically the data generated by cylc workflows is stored in the “share” directory which is referenced by an environment variable rather than a command:
+https://cylc.github.io/doc/built-sphinx/suite-config.html#suite-share-directories

### 2020-02-20 paper.tex
Stronger abstract

--- 2020-02-19 paper.tex
@@
 The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing. 
-Traditionally, in HPC, the concerns of storage and computing are separated and optimised independently from each other and the needs of the end-to-end user.
-
-As climate and weather workflows become increasingly complex and blended beyond data centres while, at the same time, storage hierarchies become deeper, the community investigates ways to reorganise storage access to utilise such systems fully.
+In HPC, the concerns of storage and computing are traditionally separated and optimised independently from each other and the needs of the end-to-end user, but in complex workflows this is becoming problematic.
+These problems are particularly acute in climate and weather workflows, which as well as become increasingly complex and exploiting deep storage hierarchies, can involve multiple data centres.
 
 The key contributions of this paper are:
-1) we sketch the vision of an integrated data-driven approach and discuss the challenges and implications of this strategy, and 2) we illustrate architecture and roadmap that allows the seamless integration into current climate and weather workflows.
-The tools employed here to achieve an extended workflow are Cylc, XIOS, DDN IME, and ESDM.
+1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated  challenges and implications, and 2) An architecture and roadmap consistent with that vision that would allow seamless integration into current climate and weather workflows as it utilises versions of mostly existing tools (Cylc, XIOS, DDN IME, and ESDM).
 
-We believe workflows composed of data, computing, and communication-intensive tasks should drive the interfaces and hardware configurations to best support the programming models.
-The changes proposed here increase the opportunity of implementations for smarter scheduling of computing and storage in heterogeneous storage systems.
+This vision is built on the belief that workflows composed of data, computing, and communication-intensive tasks should drive  interfaces and hardware configurations to best support the programming models.
+When delivered, the work proposed here will increase the opportunity for smarter scheduling of computing and storage in heterogeneous storage systems.
 

### 2020-02-20 paper.tex
Plural of PB is PB ... and one intro paragraph rewritten as I didn't understand it was it was

@@
 %Historical data spans over 170 years, generating 34Tb for low resolution and 119Tb for high resolution, while %projections will generate approximately half of this.
-In total, at the Met Office, on average 1.5\,PBs and 14\,PBs per day are written and read, respectively, for all climate and weather forecasts across all HPC clusters.
+In total, at the Met Office, on average 1.5\,PB and 14\,PB per day are written and read, respectively, for all climate and weather forecasts across all HPC clusters.
 %The above supports the view of how important data management already is and these numbers will grow with resolution and more ensembles.
@@
 
-Since performance is of crucial importance to large-scale workflows, careful attention must be paid to exploit the system characteristics of the target supercomputer.
-A specific supercomputer typically requires substantial changes to the workflow to tailor it to the particular machine and obtain the best performance.
-For instance, a data-driven workflow may benefit from using a heterogeneous set of computing and storage technology at the same time.
+Since performance is of crucial importance to large-scale workflows, careful attention must be paid to exploit the system characteristics of the target computing centre. 
+For instance, a data-driven workflow may benefit from the explicit and simultaneous use of a locally heterogeneous set of computing and storage technologies. 
+This means that substantial changes may be required to a workflow to tailor it to a particular supercomputer environment in order to obtain the best performance.
 
@@
 This project is funded by the European Union's Horizon 2020 research and innovation programme under grant agreement No. 823988.
-We thank our collaborator Bryan Lawrence for their input to this article.
+We thank our collaborator Bryan Lawrence for his input to this article.
 

### 2020-02-24 paper.tex
Revising and cleaning

@@
 
-\author{Julian M.
-Kunkel\footnote{\label{uread}University of Reading, Reading, United Kingdom}\orcidID{0000-0002-6915-1179} \and Luciana R.
-Pedro\footnoteref{uread}\orcidID{0000-0001-8365-6264} \and \textcolor{red}{Please add yourself...}
+\author{Julian M. Kunkel\footnote{\label{uread}University of Reading, Reading, United Kingdom}\orcidID{0000-0002-6915-1179} \and Luciana R. Pedro\footnoteref{uread}\orcidID{0000-0001-8365-6264}
 }
@@
 The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
-In HPC, the concerns of storage and computing are traditionally separated and optimised independently from each other and the needs of the end-to-end user, but in complex workflows this is becoming problematic.
+In HPC, the concerns of storage and computing are traditionally separated and optimised independently from each other and the needs of the end-to-end user.
+Still, in complex workflows, this is becoming problematic.
 These problems are particularly acute in climate and weather workflows, which as well as become increasingly complex and exploiting deep storage hierarchies, can involve multiple data centres.
@@
 In total, at the Met Office, on average 1.5\,PB and 14\,PB per day are written and read, respectively, for all climate and weather forecasts across all HPC clusters.
-%The above supports the view of how important data management already is and these numbers will grow with resolution and more ensembles.
+%The above supports the view of how vital data management already is, and these numbers will grow with resolution and more ensembles.
 %Existing climate experiments do ensemble experiments sequentially as few institutions have resources for parallel climate ensembles.
@@
 
-In this section, we describe how workflows are executed in a typical software stack and the typical IO stack involved in running a single application.
+In this section, we describe how workflows are executed in a typical software stack and the standard I/O stack involved in running a single application.
 
@@
 %With the current techniques, the evolution of the workflow could be a significant impediment
-With Cylc, tasks from multiple cycles may be able to run concurrently without violating dependencies preventing the issue of delays that cause one cycle to run into another.
+With Cylc, tasks from multiple cycles may be able to run concurrently without violating dependencies, preventing the issue of delays that cause one cycle to run into another.
 Cylc was written in Python and built around a new scheduling algorithm that can manage infinite workflows of cycling tasks without a sequential cycle loop.
@@
 Consider the Cylc workflow for a toy monthly cycling workflow in \cref{pic/cylc1}.
-In this workflow, an atmospheric model (labelled as \textbf{model} in the figure) simulates the physics from a current state to predict the future, for example, a month later.
+In this workflow, an atmospheric model (labelled as \texttt{model} in the figure) simulates the physics from a current state to predict the future, for example, a month later.
 In climate research, this process is repeated in the model to simulate years into the future.
 Once the simulation of any month is computed, this task completes and implicitly the data for this month becomes available and can now be analysed.
-In this workflow, the \textbf{model} task is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks, all specified as a workflow in a Cylc configuration file (\texttt{conf.rc}).
+In this workflow, the \texttt{model} task is followed by postprocessing (\texttt{post}), forecast verification (\texttt{ver}), and product generation (\texttt{prod}) tasks, all specified as a workflow in a Cylc configuration file (\texttt{flow.cylc}).
 
@@
 The information Cylc uses to control a given workflow is the task dependency.
-The developers define, in a script file, for each task the parallelism settings and where the data is to be stored.
+In a script file, the developers define, for each task, the parallelism settings and where the data is to be stored.
 %Decisions about storage must take into consideration the system architecture and the relevance of heterogeneous infrastructure in which the workflow will run.
@@
 
-While Cylc is directing the execution of workflows, several components are involved in the execution.
-The software stack involved in executing a general workflow is depicted in \cref{pic/stages}.
+While Cylc is directing the execution of workflows, several components are involved in the implementation.
+The software stack involved a general workflow is depicted in \cref{pic/stages}.
 %As a preliminary, the user has to specify a configuration representing the Cylc workflow and provide scripts for the individual tasks/applications to be executed.
@@
 
-  \item \textbf{Workload Manager} such as Slurm~\cite{Jette02slurm:simple} is responsible to allocate compute resources to a batch job and perform the job scheduling.
-  In our case, it now queues the job that represent the Cylc task and plans its execution considering the scheduling policy of the data centre.
+  \item \textbf{Workload Manager} such as Slurm~\cite{Jette02slurm:simple} is responsible for allocating compute resources to a batch job and perform the job scheduling.
+  In our case, it now queues the job that represents the Cylc task and plans its execution, considering the scheduling policy of the data centre.
   Once the job is scheduled to be dispatched, i.e., resources are available, and the job priority is the highest, it is started on the supercomputer.
@@
 
-  \item \textbf{Application} are executed taking the generated filenames set by the script.
+  \item \textbf{Application} is executed taking the generated filenames set by the script.
 \end{enumerate}
@@
 Climate applications may have complex I/O stacks, as can be seen in \cref{pic/layers-xios}.
-In this case, we assume the application uses XIOS~\cite{xios} which is providing domain-specific semantics to climate and weather.
+In this case, we assume the application uses XIOS~\cite{xios}, which is providing domain-specific semantics to climate and weather.
 It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF4~\cite{netcdf} to store data as a file.
@@
 Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy. % (\url{https://www.ddn.com/products/ime-flash-native-data-cache/})
-The tasks within any given workflow could benefit from utilizing different combinations of storage and computing infrastructure.
+The tasks within any given workflow could benefit from utilising different combinations of storage and computing infrastructure.
 
@@
 Consider, for instance, the situation where each computing node has access to three file systems: a fast \texttt{scratch} file system on which data may reside only for a week, a slower \texttt{work} file system, and a \texttt{local} file system.
-Most current workflows utilize \texttt{work} and \texttt{scratch} systems.
+Most current workflows utilise \texttt{work} and \texttt{scratch} systems.
 When a task is set to run, the corresponding dataset would be moved from \texttt{work} to \texttt{scratch}, processed, and the resulting dataset transferred back to \texttt{work}.
@@
 %\texttt{local} is not used.
-In this situation, there are many obvious opportunities to utilise data migration to optimize performance, and many possible optimization criteria (e.g., costs).
+In this situation, there are many obvious opportunities to utilise data migration to optimise performance, and many possible optimisation criteria (e.g., costs).
 However, with a multitude of file systems which differ at each data centre, such optimisations would be difficult to achieve manually for users.
-Policy-driven systems and burst-buffers perform such optimizations automatically to some extent, however, as they lack information about the workflow, they cannot optimize workflows perfectly.
+Policy-driven systems and burst-buffers perform such optimisations automatically to some extent. However, as they lack information about the workflow, they cannot optimise workflows perfectly.
 
@@
 
-Related work can be categorized into:
+Related work can be categorised into:
 1) technology that exploits heterogeneous storage environments and supports user-directed policies;
@@
 Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy.
-They can be integrated into hardware capabilities such as DDN IME~\cite{BODIAIFSFI19} or pure software solutions.
+They can be integrated into hardware capabilities such as DDN IME~\cite{BODIAIFSFI19} or simple software solutions.
 A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
-File systems and data management software such as Spectrum Scale, HPSS, BeeGFS, and Lustre (e.g., using the progressive file layouts feature) provide a hierarchical storage management allowing to store data on different storage technology according to such administrator-provided policies.
+File systems and data management software such as Spectrum Scale, HPSS, BeeGFS, and Lustre (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to such administrator-provided policies.
 However, the semantical information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.
 
-The storage community had also adjusted various higher-level software to support storage tiering on top of several storage systems, for instance, ADIOS provides in-memory staging that had been exploited by applications~\cite{slawinska2013maya}.
+The storage community had also adjusted various higher-level software to support storage tiering on top of several storage systems.
+For instance, ADIOS provides in-memory staging that had been exploited by applications~\cite{slawinska2013maya}.
 Hermes~\cite{kougkas2020acceleration}, provides a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
-iRODS~\cite{rajasekar2010irods}, is the rule-oriented data system that allows scientists to organize data into shareable collections and provide metadata such as information about data life cycle.
+iRODS~\cite{rajasekar2010irods}, is the rule-oriented data system that allows scientists to organise data into shareable collections and provide metadata such as information about the data life cycle.
 The system provides several patterns for workflows that consider data locality and data migration/replication.
 
-Finally, there have also been extensions to batch schedulers to perform data staging for utilizing node-local storage, for example, NORNS as an extension to Slurm~\cite{miranda2019norns}.
+Finally, there have also been extensions to batch schedulers to perform data staging for utilising node-local storage, for example, NORNS as an extension to Slurm~\cite{miranda2019norns}.
 
@@
 A good overview of the flavours of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
-The article states a SWfMS \textit{“should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources”}.
+The article states an SWfMS \textit{“should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources”}.
 %"An abstract scientific workflow is defined by the functionality of each activity (or sub-workflow) and data dependencies between activities (or sub-workflows)"
-Existing solutions recognize challenges in data variety (formats of the input data), the opportunity to optimize the schedule by moving code to data, and may consider capacity of available data storage.
-There have been solutions that allow to specify the data dependencies for tasks.
-The execution engine Dryad~\cite{isard2007dryad}, for example, allows to transfer data between tasks via files or directly using TCP connections and attempts to schedule tasks on the same nodes or racks.
+Existing solutions recognise challenges in data variety (formats of the input data), the opportunity to optimise the schedule by moving code to data, and may consider the capacity of available data storage.
+There have been solutions that allow specifying the data dependencies for tasks.
+The execution engine Dryad~\cite{isard2007dryad}, for example, allows transferring data between tasks via files or directly using TCP connections and attempts to schedule tasks on the same nodes or racks.
 Swift/T is a scripting language for describing dataflow processing allowing to execute ensembles of applications~\cite{ozik2016desktop}.
 Recent improvements aim to migrate data to a local cache allowing to exploit locality, e.g., in \cite{dai2018cross}, information about locality is proposed to be stored in extended attributes.
-In \cite{TUIBIHWLSC19}, an approach was presented to monitor and analyze IO behavior of HPC workflows.
+In \cite{TUIBIHWLSC19}, an approach was presented to monitor and analyse I/O behaviour of HPC workflows.
 
-Various early research in grid workflows and lately cloud use cases attempts to maximize data locality in that respect.
+Various early research in grid workflows and lately cloud use cases attempts to maximise data locality in that respect.
 Economic factors (including storage costs) for workflow execution are discussed in \cite{alkhanak2016cost}.
-In \cite{deelman2019role}, the authors discuss the role of Machine Learning for workflow execution and elaborate an the general potential for resource provisioning such as optimization of runtime parameters, data movements, and hierarchical storage.
+In \cite{deelman2019role}, the authors discuss the role of Machine Learning for workflow execution and elaborate a general potential for resource provisionings such as optimisation of runtime parameters, data movements, and hierarchical storage.
 An ML model is introduced in \cite{subedi2019leveraging} that stages data for in-situ analysis by exploiting the access patterns.
@@
 %for data-intensive applications. In: Supercomputing (SC)
-%Companion: High Performance Computing, Networking
+%Companion: High-Performance Computing, Networking
 %Storage and Analysis, pp. 140–149 (2012)
@@
 %staging data to an available location just somehow.
-%"storage service provides a temporary repository for application files that is, input files that are required for task execution, and output files that are he result of execution.""
+%"storage service provides a temporary repository for application files that is, input files that are required for task execution, and output files that are the result of execution.""
 
-Workflow systems can also be utilized specifically for the purpose of reproducing scientific results, i.e., recomputing the results.
-Those scalable workflow solutions typically utilize a container solution to allow execution in an arbitrary software environment.
+Workflow systems can also be utilised specifically to reproduce scientific results, i.e., recompute the results.
+Those scalable workflow solutions typically utilise a container solution to allow execution in an arbitrary software environment.
 Popper~\cite{jimenez2017popper}, Snakemake~\cite{bts480}, and Nextflow~\cite{Nextflow} provide a language to specify workflows and to execute them.
@@
 In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
-For each new cycle, the \texttt{checkpoint} from the previous cycle, called here \textbf{Product\,1}, is used as input to starting the next cycle.
+For each new cycle, the \texttt{checkpoint} from the previous cycle, called here Product\,1, is used as input to starting the next cycle.
 Most of the workflow can run automatically, except for the manual quality control of the products and the final data usage of Product\,3.
@@
 
-The abstraction and automation of the I/O inside a workflow allows a runtime system to improve data placement and apply data reduction on heterogeneous storage systems.
-Taking into consideration the architecture and workflow information, a smarter schedule can now be realized by exploiting the additional information.
+The abstraction and automation of the I/O inside a workflow allow a runtime system to improve data placement and apply data reduction on heterogeneous storage systems.
+Taking into consideration the architecture and workflow information, a smarter schedule can now be realised by exploiting the additional information.
 Value and priority can influence fault-tolerance strategies and imply the quality of service for performance and availability.
 Aspects like data reproducibility (can it be recomputed easily), type of the experiment (test, production), and runtime constraints for the overall and potential workflow could allow reducing costs and, hence, increasing scientific output.
-Next, we outline the two core strategies and potential the proposed vision can bring to improve current workflows.
+Next, we outline the two core strategies, and potential the proposed vision can bring to improve current workflows.
 
@@
 
-Data placement encompasses all data movement related activities such as transfer, staging, replication, space allocation and de-allocation, registering and unregistering metadata, locating and retrieving data\footnote{https://www.igi-global.com/dictionary/data-aware-distributed-batch-scheduling/6782}.
-The general idea is to host a data set on the storage system that is most favorable in terms of performance, cost-effectiveness, availability for the access pattern observed by the workflow.
-It requires to optimize data locality, where locality is twofold, spatial and temporal on a variety of characteristics.
+Data placement encompasses all data movement-related activities such as transfer, staging, replication, space allocation and de-allocation, registering and unregistering metadata, locating and retrieving data\footnote{https://www.igi-global.com/dictionary/data-aware-distributed-batch-scheduling/6782}.
+The general idea is to host a data set on the storage system that is most favourable in terms of performance, cost-effectiveness, availability for the access pattern observed by the workflow.
+It requires to optimise data locality, where locality is twofold, spatial and temporal on a variety of characteristics.
 For optimising data placement, there are several approaches:%: \textbf{in-memory data}, \textbf{data allocation}, \textbf{data migration}, and \textbf{data replication}.
@@
 It would be interesting to explore storage options for the datasets and, e.g., having datasets from different cycles placed at different storage systems.
-For instance, in \cref{pic/workflow-s3-n}, alternating the storage location for \textbf{Product\,2} into two scratch file systems is something that would be a simple job for an I/O-aware scheduler.
+For instance, in \cref{pic/workflow-s3-n}, alternating the storage location for Product\,2 into two scratch file systems is something that would be a simple job for an I/O-aware scheduler.
 However, currently, that implies having at least two scripts for that task with information about the different storage placement.
@@
 \item[Data Migration] is the process of transferring data from one storage system to another.
-Data movement involves a large overhead, both in terms of latency and energy-efficient computing, hence needs to be considered carefully.
+Data movement involves a significant overhead, both in terms of latency and energy-efficient computing, hence needs to be considered carefully.
 \cref{fig:lifecycle} introduces three possible life cycles for a specific dataset and explains how migrations can be done to improve datasets accessibility.
@@
 As an example, in \cref{pic/workflow-s3-n}, the outcome of Task\,1 may be used directly by Task\,2.
-To achieve a decoupling between producer and consumer, it may also be kept in memory and cached.
+To achieve decoupling between producer and consumer, it may also be kept in memory and cached.
 \end{description}
@@
 Knowing the characteristics of data production and usage makes it simpler for scientists to annotate the required precision of data in those workflows.
-The storage system can exploit such information by reducing precision of data and automatically picking an appropriate compression algorithm.
+The storage system can exploit such information by reducing the precision of data and automatically picking an appropriate compression algorithm.
 
@@
 Intermediate states could be rerun by utilising virtualisation and container technologies.
-Consider \cref{pic/workflow-s3-n} again and that, at every K cycles of the workflow, the generated \textbf{Product\,3} (from Cycle 1 to Cycle K) are used in a validation task, called here \textbf{check}.
-From the workflow\footnote{The $P_i C_j$ notation represents the Product\,$i$ generated in the Cycle $j$.}, we know that $P_3 C_1$ will be used to construct $P_3 C_2$ and then \textbf{check}.
+Consider \cref{pic/workflow-s3-n} again and that, at every K cycles of the workflow, the generated Product\,3 (from Cycle 1 to Cycle K) are used in a validation task, called here \texttt{check}.
+From the workflow\footnote{The $P_i C_j$ notation represents the Product\,$i$ generated in the Cycle $j$.}, we know that $P_3 C_1$ will be used to construct $P_3 C_2$ and then \texttt{check}.
 This dataset will probably be stored somewhere, and it will not be used until the workflow reaches the K-th cycle.
 One alternative is to delete it after it was first used and then recompute it when time is right.
-The cost of doing that is storing \texttt{checkpoint} and then use it to reconstruct product $P_3 C_1$.
+The cost of doing that is storing \texttt{checkpoint} and then use it to reconstruct $P_3 C_1$.
 If, for instance, $P_3 C_1$ is a large dataset, \texttt{checkpoint} is small, and computing time is short.
-It is easy to see that deleting and recomputing it may improve the costs for running a workflow.
+It is easy to see that deleting and recomputing it may improve the costs of running a workflow.
 %We need, then, information about datasets volume and computing time to be able to make an accurate decision.
@@
 \item[Optimisation] The workflow will be optimised specifically for the available system infrastructure and extra information about the data.
-In particularly, by exposing the heterogeneous architecture, potentially runtime characteristics can be considered.
+In particular, by exposing the heterogeneous architecture, potentially runtime characteristics can be considered.
 By using information about the value of data, policies for data management (storage resilience,  recomputation, replication, etc.) can be decided.
@@
 
-ESDM is used as IO middleware in the parallel application (in NetCDF or directly) and orchestrates the IO according to a configuration file.
+ESDM is used as I/O middleware in the parallel application (in NetCDF or directly) and orchestrates the I/O according to a configuration file.
 The information about the system is provided using a simplified ESDM configuration file (\texttt{esdm.conf}) provided by the data centre (or expert users).
@@
 %\jk{Good suggestion!
-That absolutely depends how a Cylc file is parsed and analyzed.
+That absolutely depends how a Cylc file is parsed and analysed.
 Keeping it separately gives us the most flexibility and parsing comments could lead to misinterpreting already existing comments and acting wrongly.
@@
 What sort of node?
-Is the ESDM running as an I/O server, or are you assuming it is being used in an IO server etc}
+Is the ESDM running as an I/O server, or are you assuming it is being used in an I/O server etc}
 %\lr{Well noticed.
@@
 An example of an I/O-workflow configuration file is shown in \Cref{lst:cylc}.
-In this file, information about Task\,1 is given by example, and we expect the extra information about all tasks in the same file; ultimately, this could be integrated into the workflow specification file of Cylc.
+In this file, information about Task\,1 is given by example, and we expect the extra information about all tasks in the same file.
+Ultimately, this could be integrated into the workflow specification file of Cylc.
 
@@
 \begin{description}[itemsep=0pt]
-  \item[Name] A basic name for the field/data generated (Lines: 8, 16, 23). It is extended by a pattern defined in a variable.
+  \item[Name] A basic name for the field/data generated. It is extended by a pattern defined in a variable (Lines: 8, 16, 23).
+
   \item[Pattern] The frequency the data is output (Lines: 9, 17).
@@
 
-  \item[Size] An estimate of the data size (Lines: 13, 26). This field can be inferred if dimension and datatype are provided.
+  \item[Size] An estimate of the data size\footnote{This field can be inferred if dimension and datatype are provided.} (Lines: 13, 26).
 
-  \item[Dimension] The data dimension (Line 21).
+  \item[Dimension] The data dimension (Line: 21).
 
-  \item[Accuracy] This includes various characteristics quantifying the required level of data precision; For example, in Line~14.
+  \item[Accuracy] Characteristics quantifying the required level of data precision (Line: 14).
 
@@
 Note that the user may not be able to provide all required information.
-This can be handled by assuming a default safe behavior.
-For instance, in case of a missing data precision, data should be retained in original form.
+This can be handled by assuming a default safe behaviour.
+For instance, in the case of missing data precision, data should be retained in original form.
 Knowing the dimension or size a priori might be difficult for scientists, e.g., the log file size is unclear.
-In this case, the user may insert relative information like small or big, reflecting that any information is better than no information at all.
+In this case, the user may insert relevant information like small or big, reflecting that any information is better than no information at all.
 In future, we will explore how to automatically infer the output volume from the input or by using monitoring.
@@
 
-The steps to execute a workflow enriched with IO information and performing smarter scheduling is depicted in \cref{pic/stages-io}.
+The steps to execute a workflow enriched with I/O information and performing smarter scheduling is depicted in \cref{pic/stages-io}.
 The suggested alterations can be seen in boxes pointed by red arrows, and the remaining components are the current state-of-the-art for workflows in climate and weather from \cref{pic/stages}.
@@
 
-Organising the data placement on storage tiers is particularly in the domain of climate and weather performed by the users or via policies, leading often to suboptimal decisions.
-Additionally, the manual optimization and hard-coding of storage locations is non-portable and an error-prone task.
-We believe users must be able to express their workflows in an abstract fashion.
+In the domain of climate and weather, organising the data placement on storage tiers is performed by the users or via policies, often leading to suboptimal decisions.
+Additionally, the manual optimisation and hard-coding of storage locations is non-portable and an error-prone task.
+We believe users must be able to express their workflows abstractly.
 By increasing the abstraction level for scientists, not only tedious manual optimisation could be automatised, but also strategies for data placement and data reduction can be harnessed.
-With knowledge about the data pattern, the runtime system could generate optimized execution plans and monitor their execution.
+With knowledge about the data pattern, the runtime system could generate optimised execution plans and monitor their execution.
 In this work, we describe the general vision and a specific design for the software stack in the domain in climate and weather that we work on in the ESiWACE project.
-The changes proposed increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system in a workflow.
+The proposed changes increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system in a workflow.
 
@@
 This project is funded by the European Union's Horizon 2020 research and innovation programme under grant agreement No. 823988.
-We thank our collaborator Bryan Lawrence for his input to this article.
+We thank our collaborators Bryan Lawrence, Glenn Greed, and David Matthews for their input to this article.
 

### 2020-02-24 pic/svgs/convertSVGtoPDF.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-mkdir "$PWD"/pdf
-for file in $PWD/*.svg
-    do
-        filename=$(basename "$file")
-        inkscape "$file" -d 1200 -A "$PWD"/pdf/"${filename%.svg}.pdf"
-    done

### 2020-02-25 paper.tex
Revision - Intro

@@
 
+\lr{JT, Hua ?}
+
 \title{Potential of I/O-Aware Workflows in Climate and Weather}
@@
 In HPC, the concerns of storage and computing are traditionally separated and optimised independently from each other and the needs of the end-to-end user.
-Still, in complex workflows, this is becoming problematic.
+However, in complex workflows, this is becoming problematic.
 These problems are particularly acute in climate and weather workflows, which as well as become increasingly complex and exploiting deep storage hierarchies, can involve multiple data centres.
@@
 The key contributions of this paper are:
-1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with that vision that would allow seamless integration into current climate and weather workflows as it utilises versions of mostly existing tools (Cylc, XIOS, DDN IME, and ESDM).
+1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of mostly existing tools (Cylc, XIOS, DDN IME, and ESDM).
 
-This vision is built on the belief that workflows composed of data, computing, and communication-intensive tasks should drive interfaces and hardware configurations to best support the programming models.
-When delivered, the work proposed here will increase the opportunity for smarter scheduling of computing and storage in heterogeneous storage systems.
+\lr{mostly??}
+
+The vision proposed here is built on the belief that workflows composed of data, computing, and communication-intensive tasks should drive interfaces and hardware configurations to best support the programming models.
+When delivered, this work will increase the opportunity for smarter scheduling of computing and storage in heterogeneous storage systems.
+
+\lr{computing and storage?? We talked about it before. Not only storage? I think the Cylc team spotted this problem too.}
 
@@
 In the domain of climate and weather, there is a considerable demand for the orchestration of ensembles of simulation models and the generation of data products.
-A weather forecast service such as the operational weather forecast workflow in Met Office writes around 200TB every day and reads around 600\,TB.
+A service such as the operational weather forecast workflow in Met Office writes around 200\,TB and reads around 600\,TB every day.
 %For a typical Met Office Unified Model (UM) from the Coupled Model Intercomparison Project (CMIP) run, some %rough estimates of data volumes are:
@@
 %Historical data spans over 170 years, generating 34Tb for low resolution and 119Tb for high resolution, while %projections will generate approximately half of this.
-In total, at the Met Office, on average 1.5\,PB and 14\,PB per day are written and read, respectively, for all climate and weather forecasts across all HPC clusters.
+In total, at the Met Office, on average 1.5\,PB and 14\,PB are written and read per day, respectively, for all climate and weather forecasts across all HPC clusters.
 %The above supports the view of how vital data management already is, and these numbers will grow with resolution and more ensembles.
@@
 
-Based on their needs, the HPC community has developed a software ecosystem that supports climate and weather scientists to execute their large-scale workflows.
+\lr{I don't quite understand the numbers.}
+
+Based on the needs of climate and weather researchers, the HPC community has developed a software ecosystem that supports scientists to execute their large-scale workflows.
 While the current advances correspond to a big leap forward, many processes still require experts.
@@
 
+\lr{For instance, a data-driven workflow may benefit from the explicit and simultaneous use of a locally heterogeneous set of computing and storage technologies.
+For example, porting a workflow from one system to another still requires adjusting runtime parameters of applications and deciding on how data is managed.}
+
+\lr{What the relation between those two examples? They should be together, no?}
+
+\lr{computing and storage technologies again.}
+
 Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
-As the complexity of systems expands and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise resource usage in workflows.
-In many cases, modifications are not performed because: 1) They are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) Users are not aware of the potential of the complex system; and, 3) because it is a one-time explorative workflow.
+As the complexity of systems expands and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise the usage of resources in workflows.
+In many cases, modifications are not performed because: 1) They are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) It is a one-time explorative workflow and, 3) Users are not aware of the potential of the complex system.
+
 In this paper, we illustrate how knowing the Input/Output (I/O) characteristics of workflow tasks and overall experimental design helps to optimise the execution of climate and weather workflows.
@@
 
-This paper is structured as follows:
+\lr{computing and storage technologies again.}
+
+This paper is structured as follows.
 First, we describe the software stack involved in executing workflows in climate and weather in \Cref{sec:workflows}.
-Next, the vision for including knowledge about data requirements and characteristics is sketched in  \Cref{sec:vision} outlining the potential benefit the automatic exploitation may bring.
-Our design based on existing components in climate and weather is described in \Cref{sec:design}.
+Related work in heterogeneous storage environments and solutions for workflow processing are presented in \Cref{sec:art}.
+Next, the vision for including knowledge about data requirements and characteristics is sketched in \Cref{sec:vision} outlining the potential benefit the automatic exploitation might bring.
+Our design, based on existing components in climate and weather, is described in \Cref{sec:design}.
 The paper is concluded in \Cref{sec:conclusions}.
+
+\lr{data requirements and characteristics == data management?}
 
@@
 \section{State-of-the-Art}
+\label{sec:art}
 

### 2020-02-25 paper.tex
Chapter 1

@@
 For instance, a data-driven workflow may benefit from the explicit and simultaneous use of a locally heterogeneous set of computing and storage technologies.
+\lr{computing and storage technologies again.}
 This means that substantial changes may be required to a workflow to tailor it to a particular supercomputer environment in order to obtain the best performance.
@@
 
-\lr{What the relation between those two examples? They should be together, no?}
-
-\lr{computing and storage technologies again.}
+\lr{What's the relation between those two examples? They should be together, no?}
 
 Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
-As the complexity of systems expands and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise the usage of resources in workflows.
+As the complexity of systems expands and alternative storage and computing technologies \lr{computing and storage technologies again.} provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise the usage of resources in workflows.
 In many cases, modifications are not performed because: 1) They are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) It is a one-time explorative workflow and, 3) Users are not aware of the potential of the complex system.
@@
 
-\lr{computing and storage technologies again.}
-
 This paper is structured as follows.
@@
 Related work in heterogeneous storage environments and solutions for workflow processing are presented in \Cref{sec:art}.
-Next, the vision for including knowledge about data requirements and characteristics is sketched in \Cref{sec:vision} outlining the potential benefit the automatic exploitation might bring.
+Next, the vision for including knowledge about data requirements and characteristics \lr{data requirements and characteristics == data management?} is sketched in \Cref{sec:vision} outlining the potential benefit the automatic exploitation might bring.
 Our design, based on existing components in climate and weather, is described in \Cref{sec:design}.
 The paper is concluded in \Cref{sec:conclusions}.
-
-\lr{data requirements and characteristics == data management?}
 
@@
 
-In this section, we describe how workflows are executed in a typical software stack and the standard I/O stack involved in running a single application.
+In this section, we describe how workflows are executed in a typical software stack \lr{software stack == climate and weather application?
+The section is not only about the software stack.} and the standard I/O stack involved in running a single application.
 
@@
 At any point during workflow execution, only the dependence between the individual tasks matters, regardless of their particular cycle points.
+The information Cylc uses to control a given workflow is the task dependency.
+In a script file, the developers define, for each task, the parallelism settings and where the data is to be stored.
+%Decisions about storage must take into consideration the system architecture and the relevance of heterogeneous infrastructure in which the workflow will run.
 
@@
 In climate research, this process is repeated in the model to simulate years into the future.
-Once the simulation of any month is computed, this task completes and implicitly the data for this month becomes available and can now be analysed.
-In this workflow, the \texttt{model} task is followed by postprocessing (\texttt{post}), forecast verification (\texttt{ver}), and product generation (\texttt{prod}) tasks, all specified as a workflow in a Cylc configuration file (\texttt{flow.cylc}).
+% Once the simulation of any month is computed, this task completes and implicitly the data for this month becomes available and can now be analysed.
+Once the simulation of any month is computed, the data for this month becomes available and can now be analysed.
+In this workflow, the task \texttt{model} is followed by tasks postprocessing (\texttt{post}), forecast verification (\texttt{ver}), and product generation (\texttt{prod}), all specified as a workflow in a Cylc configuration file (\texttt{flow.cylc}).
 
-\fig{width=0.6\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file}
-
-The information Cylc uses to control a given workflow is the task dependency.
-In a script file, the developers define, for each task, the parallelism settings and where the data is to be stored.
-%Decisions about storage must take into consideration the system architecture and the relevance of heterogeneous infrastructure in which the workflow will run.
+\fig{width=0.8\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file}
 
@@
 
-While Cylc is directing the execution of workflows, several components are involved in the implementation.
-The software stack involved a general workflow is depicted in \cref{pic/stages}.
+While Cylc is directing the execution of workflows, several components are presented in the implementation.
+The software stack involved in a general workflow is depicted in \cref{pic/stages}.
 %As a preliminary, the user has to specify a configuration representing the Cylc workflow and provide scripts for the individual tasks/applications to be executed.
@@
 
-  \item \textbf{Workload Manager} such as Slurm~\cite{Jette02slurm:simple} is responsible for allocating compute resources to a batch job and perform the job scheduling.
-  In our case, it now queues the job that represents the Cylc task and plans its execution, considering the scheduling policy of the data centre.
+  \item \textbf{Workload Manager} such as Slurm~\cite{Jette02slurm:simple} is responsible for allocating compute resources to a batch job and performing the job scheduling.
+%  In our case, it now queues the job that represents the Cylc task and plans its execution, considering the scheduling policy of the data centre.
+  The selected tool queues the job that represents the Cylc task and plans its execution, considering the scheduling policy of the data centre.
   Once the job is scheduled to be dispatched, i.e., resources are available, and the job priority is the highest, it is started on the supercomputer.
 
-  \item \textbf{Job} provides the environment with the resources it runs the user-provided script on one of the nodes allocated for the job.
-  The job sets environment variables containing information about the environment of the batch job, e.g., the compute nodes allocated, and then enacts the Cylc provided script on one node.
+%  \item \textbf{Job} provides the environment with the resources it runs the user-provided script on one of the nodes allocated for the job.
+%  The job sets environment variables containing information about the environment of the batch job, e.g., the compute nodes allocated, and then enacts the Cylc provided script on one node.
 
-  \item \textbf{Script} executes the shell commands which is one or multiple (potentially parallel) applications to run sequentially.
+  \lr{I rewrite this part.}
+
+  \item \textbf{Job} provides the environment with the resources and it runs the user-provided script on one of the nodes allocated for it.
+  Local variables containing information about the environment of the batch job, e.g., the compute nodes allocated, enact the Cylc provided script on that node.
+
+  \item \textbf{Script} executes the shell commands with one or multiple (potentially parallel) applications to run sequentially.
   During the creation of the script, Cylc has included variables that describe the task in the workflow.
-  The user script may use commands to create a filename considering the cycle of Cylc or store data in a workflow-specific shared directory.
   The information is typically fed into the application(s) representing the task, and so defines the storage location.
+  The user script may use\lr{May use?
+"Application" talks about generated filenames.} commands to create a filename considering the cycle or \lr{or??} store data in a workflow-specific shared directory.
 
@@
 It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF4~\cite{netcdf} to store data as a file.
-
-Under the hood, NetCDF4 uses the HDF5 API and file format.
+Under the hood, NetCDF4 uses the HDF5 API and file format. \lr{and file format? What does that mean?}
 Internally, HDF5 uses MPI and its data types to specify the nature of the data stored.
 Finally, data is stored on a parallel file system like Lustre which, on the server-side, stores data in a local file system on block devices such as SSDs and HDDs.
+
+\lr{NetCDF4 or NetCDF? We should be consistent in the whole paper.}
 
@@
 
+\lr{We shoud talk about how ESDM would replace the components in Fig 3. That was in the text before and it's important.}
+
 \subsection{Data Center Infrastructure}
 
-Data centres are providing an infrastructure consisting of computing and storage devices with different characteristics making them more efficient for specific tasks and satisfying the needs for different workflows.
-Take, for example, the supercomputer Mistral at DKRZ that consists of 3,321 nodes\footnote{\url{https://www.dkrz.de/up/systems/mistral}}.
-It offers two types of compute nodes equipped with different CPUs and a range of GPU nodes.
-Each node provides an SSD for local storage, and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
-Individual users and projects are mapped to one file system explicitly, and users can access it with work or scratch semantics:
+Data centres are providing an infrastructure consisting of computing and storage devices with different characteristics, making them more efficient for specific tasks and satisfying the needs of different workflows.
+Take, for example, the supercomputer Mistral at DKRZ, that consists of 3,321 nodes\footnote{\url{https://www.dkrz.de/up/systems/mistral}} and offers two types of compute nodes equipped with different CPUs and GPU nodes.
+Each node has an SSD for local storage, and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
+Individual users and projects are mapped to one file system explicitly, and users can access it with work or scratch semantics.
 While data is kept on the work file system indefinitely, available space is limited by a quota.
@@
 Depending on the need, the storage characteristics range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and cheap storage for long-term archival (tape).
-Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy. % (\url{https://www.ddn.com/products/ime-flash-native-data-cache/})
+% Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy. % (\url{https://www.ddn.com/products/ime-flash-native-data-cache/})
 The tasks within any given workflow could benefit from utilising different combinations of storage and computing infrastructure.
 
+\lr{Are SSD and HDD online storage?}
+
 \fig{width=0.6\columnwidth}{pic/system}{Example of an heterogeneous HPC landscape}
+
+\lr{Consider put figs 3 and 4 together to better use the space.}
 
@@
 
-Usually, the scripts representing tasks define how the data is placed on an available storage system.
-What happens in many current workflows is that they ignore the benefits of using multiple file systems concurrently and the data locality between tasks to co-locating them.
+Usually, the scripts representing tasks define how the data is placed on the available storage system.
+What happens in many current workflows is that they ignore the benefits of using multiple file systems concurrently and the data locality between tasks to colocating them.
 On top of that, in the current state-of-the-art, scientists optimise the available storage resources intuitively and compile the information about this decision-making process manually.
 
-If a user knows the workflow and the system characteristics, s/he can optimise data placement decisions.
+If a user \lr{it's really the user here?} knows the workflow and the system characteristics, s/he can optimise data placement decisions.
 Consider, for instance, the situation where each computing node has access to three file systems: a fast \texttt{scratch} file system on which data may reside only for a week, a slower \texttt{work} file system, and a \texttt{local} file system.
 Most current workflows utilise \texttt{work} and \texttt{scratch} systems.
-When a task is set to run, the corresponding dataset would be moved from \texttt{work} to \texttt{scratch}, processed, and the resulting dataset transferred back to \texttt{work}.
-If the \texttt{scratch} filesystem reached its capacity, the datasets would be moved back to \texttt{work}, and the task would continue running until it is finished, but it might be inefficient.
+When a task is set to run, the corresponding dataset would be moved from \texttt{work} to \texttt{scratch}, processed, and the resulting dataset would be transferred back to \texttt{work}.
+If the \texttt{scratch} filesystem reaches its capacity, the dataset would be moved back to \texttt{work}, and the task would continue running until it is finished, which might be inefficient.
 %\texttt{local} is not used.
-In this situation, there are many obvious opportunities to utilise data migration to optimise performance, and many possible optimisation criteria (e.g., costs).
-However, with a multitude of file systems which differ at each data centre, such optimisations would be difficult to achieve manually for users.
+In this situation, there are many obvious opportunities to utilise data migration to optimise performance, and many possible optimisation criteria (e.g., costs). \lr{Another criterion?}
+However, with a multitude of file systems that differ at each data centre, such optimisations would be difficult to achieve manually by users.
 Policy-driven systems and burst-buffers perform such optimisations automatically to some extent. However, as they lack information about the workflow, they cannot optimise workflows perfectly.
@@
 
-While Cylc schedules the workflow, EIOS can provide hints about co-locating tasks which provide the opportunity for keeping data in local storage.
+While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
 Decisions about data locality will not be made for a whole (and potentially big) workflow.

### 2020-02-25 paper.tex
Chapter 2

@@
 Related work can be categorised into:
-1) technology that exploits heterogeneous storage environments and supports user-directed policies;
-2) solutions for workflow processing.
+1) Technology that exploits heterogeneous storage environments and supports user-directed policies and
+2) Solutions for workflow processing.
 
@@
 A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
-File systems and data management software such as Spectrum Scale, HPSS, BeeGFS, and Lustre (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to such administrator-provided policies.
-However, the semantical information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.
+File systems and data management software such as Spectrum Scale, HPSS, BeeGFS, \lr{ref about those topics?
+webpage as footnote, maybe?} and Lustre (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to administrator-provided policies.
+However, the semantic information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.
 
@@
 For instance, ADIOS provides in-memory staging that had been exploited by applications~\cite{slawinska2013maya}.
-Hermes~\cite{kougkas2020acceleration}, provides a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
-iRODS~\cite{rajasekar2010irods}, is the rule-oriented data system that allows scientists to organise data into shareable collections and provide metadata such as information about the data life cycle.
-The system provides several patterns for workflows that consider data locality and data migration/replication.
+\lr{What kind of applications?
+Medical, climate, industrial, etc.}
+Hermes~\cite{kougkas2020acceleration} provides a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
+iRODS~\cite{rajasekar2010irods} is a rule-oriented data system that allows scientists to organise data into shareable collections and provide metadata such as information about the data life cycle.
+\lr{such as information about the data life cycle?
+I don't get it.}
+The system provides several patterns for workflows that consider data locality and data migration/replication. \lr{Which system? Is still talking about iRODS?}
 
-Finally, there have also been extensions to batch schedulers to perform data staging for utilising node-local storage, for example, NORNS as an extension to Slurm~\cite{miranda2019norns}.
+Finally, there have also been extensions to batch schedulers to perform data staging for utilising node-local storage, for example, NORNS~\cite{miranda2019norns} as an extension to Slurm.
 
@@
 \textbf{Workflows. }
-A good overview of the flavours of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
-The article states an SWfMS \textit{“should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources”}.
+A good overview of the flavours \lr{flavours?} of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
+The article states that SWfMS should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources.
 %"An abstract scientific workflow is defined by the functionality of each activity (or sub-workflow) and data dependencies between activities (or sub-workflows)"
-Existing solutions recognise challenges in data variety (formats of the input data), the opportunity to optimise the schedule by moving code to data, and may consider the capacity of available data storage.
-There have been solutions that allow specifying the data dependencies for tasks.
+Existing solutions recognise challenges in data variety (formats of the input data), opportunities to optimise the schedule by moving code to data, allow specifying the data dependencies for tasks, and they even may consider the capacity of the available data storage.
 The execution engine Dryad~\cite{isard2007dryad}, for example, allows transferring data between tasks via files or directly using TCP connections and attempts to schedule tasks on the same nodes or racks.
 Swift/T is a scripting language for describing dataflow processing allowing to execute ensembles of applications~\cite{ozik2016desktop}.
-Recent improvements aim to migrate data to a local cache allowing to exploit locality, e.g., in \cite{dai2018cross}, information about locality is proposed to be stored in extended attributes.
+Recent improvements aim to migrate data to a local cache allowing to exploit locality.
+For instance, in \cite{dai2018cross}, information about locality is proposed to be stored in extended attributes.
 In \cite{TUIBIHWLSC19}, an approach was presented to monitor and analyse I/O behaviour of HPC workflows.
@@
 Economic factors (including storage costs) for workflow execution are discussed in \cite{alkhanak2016cost}.
-In \cite{deelman2019role}, the authors discuss the role of Machine Learning for workflow execution and elaborate a general potential for resource provisionings such as optimisation of runtime parameters, data movements, and hierarchical storage.
-An ML model is introduced in \cite{subedi2019leveraging} that stages data for in-situ analysis by exploiting the access patterns.
+In \cite{deelman2019role}, the authors discuss the role of Machine Learning (ML) for workflow execution and elaborate a general potential for resource provisionings such as optimisation of runtime parameters, data movements, and hierarchical storage.
+In \cite{subedi2019leveraging}, an ML model that stages data for in-situ analysis by exploiting the access patterns is introduced.
 
@@
 Those scalable workflow solutions typically utilise a container solution to allow execution in an arbitrary software environment.
-Popper~\cite{jimenez2017popper}, Snakemake~\cite{bts480}, and Nextflow~\cite{Nextflow} provide a language to specify workflows and to execute them.
+Popper~\cite{jimenez2017popper}, Snakemake~\cite{bts480}, and Nextflow~\cite{Nextflow} provide a language \lr{Is it a DSL?
+I think it's important to emphasize this here.} to specify workflows and to execute them.
 Snakemake is interesting as it supports to define and infer input and output filenames.

### 2020-02-25 paper.tex
Chapter 3

@@
 
-If a user \lr{it's really the user here?} knows the workflow and the system characteristics, s/he can optimise data placement decisions.
+If a user \lr{It's really the user here?} knows the workflow and the system characteristics, s/he can optimise data placement decisions.
 Consider, for instance, the situation where each computing node has access to three file systems: a fast \texttt{scratch} file system on which data may reside only for a week, a slower \texttt{work} file system, and a \texttt{local} file system.
@@
 
-Nowadays, in order to run a job in an HPC environment efficiently, researchers have to develop profound knowledge, not only about their workflow, which is expected, but also about decisions regarding storage, communication, and computing including considerations regarding cost-efficiency of their operations.
-However, applied scientists should not spend much time understanding hardware characteristics and operative knowledge of running a data centre, but using their time to develop their work and just collect and analyse their results.
+Nowadays, in order to run a job in an HPC environment efficiently, researchers have to develop profound knowledge, not only about their workflow, which is expected, but also about decisions regarding storage, communication, computing, and considerations regarding cost-efficiency of those operations.
+However, applied scientists should not spend much time understanding hardware characteristics and operative knowledge of running a data centre, but using their knowledge to develop their work and just collect and analyse the results.
 
@@
 
-There are several approaches to implement the technology for the vision proposed in this work, and changes are needed for software components to realise it.
+There are several approaches to implement the technology for the vision proposed in this work, and changes are needed in the software components to realise it.
 In the next sections, we will discuss a specific design for our transitional roadmap considering climate and weather workflows and tools scientists from this field already use in their routine research.
+
+\lr{"a specific design" here? Next section, no?}
 
@@
 Firstly, the user must augment the workflow description with information about I/O requirements and explicitly annotate dependencies to datasets.
-Secondly, the data centre (or expert user) must provide details about the storage architecture.
+Secondly, details about the storage architecture must be available.
 
@@
 It is expected that the data centre (or expert users) can create such a configuration file, e.g., by using vendor-provided information or by executing benchmarks.
-With this information, a scheduler can make data placement, transformation, and migration decisions for individual datasets during their life cycle.
-This separation of concerns allows us to abstract from the workflow what is essential and what a system should optimise to ensure smart usage of available resources.
+With this information, a scheduler can make data placement, transformation, and migration decisions for individual datasets during their life cycle.\lr{The scheduler making data transformation?
+And it's migration already classified as data placement?}
+This separation of concerns allows us to abstract from the workflow what is essential and what a system should optimise to ensure smart usage of the available resources.
 
@@
 In general, climate and weather workflows allow specifying tasks and the dependencies among them.
-We aim to enhance the current information with characteristics for input, and output, i.e., the datasets.
+We aim to enhance the current information with characteristics for input and output, i.e., the datasets.
 An example workflow containing input datasets and (intermediate) products is illustrated in \cref{pic/workflow-s3-n}.
-Nodes represent tasks or data, and arrows indicate the dependencies.
-In the example, Task\,1 needs two datasets to perform its work, it directly communicates with Task\,2 and produces Product\,1.
-For each new cycle, the \texttt{checkpoint} from the previous cycle, called here Product\,1, is used as input to starting the next cycle.
+Round nodes represent tasks, squared nodes represent data, and arrows indicate dependencies.
+In the example, Task\,1 needs two datasets to perform its work, produces Product\,1, and it directly communicates with Task\,2.
+For each new cycle, the \texttt{checkpoint} from the previous cycle (Product\,1) is used as input to starting the next cycle.
 Most of the workflow can run automatically, except for the manual quality control of the products and the final data usage of Product\,3.
 This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
-In the approach proposed in this work, each task is annotated with the required input datasets and the generated products must include metadata such as data life cycle information, the value of data, and how long it should be kept.
+In the approach proposed in this work, each task is annotated with the required input datasets and the generated products must include metadata such as data life cycle information, value of data, and how long it should be kept.
 The idea here is to embrace the concept that tasks dependencies are really imposed by datasets dependencies.
+
+\lr{Talk a little bit more about the figure and the N cycles.}
 
@@
 Aspects like data reproducibility (can it be recomputed easily), type of the experiment (test, production), and runtime constraints for the overall and potential workflow could allow reducing costs and, hence, increasing scientific output.
-Next, we outline the two core strategies, and potential the proposed vision can bring to improve current workflows.
+Next, we outline the two core strategies and potential the proposed vision can bring to improve current workflows.
+In the design proposed in this work (\cref{sec:design}), we will focus on the data placement strategy.
 
@@
 Data placement encompasses all data movement-related activities such as transfer, staging, replication, space allocation and de-allocation, registering and unregistering metadata, locating and retrieving data\footnote{https://www.igi-global.com/dictionary/data-aware-distributed-batch-scheduling/6782}.
-The general idea is to host a data set on the storage system that is most favourable in terms of performance, cost-effectiveness, availability for the access pattern observed by the workflow.
-It requires to optimise data locality, where locality is twofold, spatial and temporal on a variety of characteristics.
+The general idea is to host a dataset on the storage system that is most favourable in terms of performance, cost-effectiveness, availability for the access pattern observed in the workflow.
+Here we are considering the optimisation of data locality, where locality is twofold, spatial and temporal on a variety of characteristics.
 For optimising data placement, there are several approaches:%: \textbf{in-memory data}, \textbf{data allocation}, \textbf{data migration}, and \textbf{data replication}.
@@
 \item[Data Allocation] is the assignment of a specific area of an available storage system to particular data.
-In current workflows, usually, the user has a script for each task defining the filenames with a prefix that places datasets generated by the same task into a specific storage\footnote{Complicated scripts would have allowed changing the storage type depending on the cycle.
+In current workflows, the user usually has a script for each task defining the filenames with a prefix that places datasets generated by the same task into a specific storage\footnote{Complicated scripts would have allowed changing the storage type depending on the cycle.
 Still, it is a significant burden to the user.}.
@@
 
+\lr{Weak justification. We are better than that.}
+
 \item[Data Migration] is the process of transferring data from one storage system to another.
@@
 \cref{fig:lifecycle} introduces three possible life cycles for a specific dataset and explains how migrations can be done to improve datasets accessibility.
-In \cref{fig:lifecycle1}, the dataset could be stored on the \texttt{local} storage first to avoid congestion on the \texttt{work} file system, then be migrated to \texttt{work} file system where subsequent tasks of the workflow may read it multiple times.
-In the end, this dataset might be an intermediate product that can then be deleted.
-Alternatively (see \cref{fig:lifecycle2}), the dataset could be stored on the \texttt{scratch} file system immediately and accessed there.
-However, that would require subsequent tasks to be placed on the same node where data is now stored, and the last read access must happen before files on \texttt{scratch} are automatically removed.
-The last scenario (see \cref{fig:lifecycle3}) presents the case where the dataset is created on \texttt{work} by a task and it is copied to a \texttt{local} node.
+In \cref{fig:lifecycle1}, the dataset could be first stored on the \texttt{local} storage to avoid congestion on the \texttt{work} file system, then be migrated to \texttt{work} file system where subsequent tasks of the workflow may read it multiple times.
+In the end, this dataset may be an intermediate product that can then be deleted.
+In \cref{fig:lifecycle2}, the dataset is stored on the \texttt{scratch} file system immediately and accessed there.
+%However, that would require subsequent tasks to be placed on the same node where data is now stored, and the last read access must happen before files on \texttt{scratch} are automatically removed.
+However, the last read access must happen before files on \texttt{scratch} are automatically removed.
+Alternatively, \cref{fig:lifecycle3} presents the case where the dataset is created on \texttt{work} by a task and it is copied to a \texttt{local} node.
 This \texttt{local} node allows reading accesses of subsequent tasks which might be beneficial for small random accesses.
-However, it would require that the subsequent tasks are placed on the same node where data is now stored.
+For the last two scenarios, subsequent tasks would have to be placed on the same node where previous data was stored.
+%However, it would require that the subsequent tasks are placed on the same node where data is now stored.
 
@@
 
+\lr{If we are really migrating, why do we have to deletes in (a) and (c)? Maybe the delete should appear at the same time as the migration in the timeline.}
+
 \item[Data Replication] in computing involves sharing information to ensure consistency between redundant resources, such as software or hardware components, to improve reliability, fault-tolerance, or accessibility.
@@
 In addition, the system may combine the replication of data by \textbf{transforming} the data into a different representation allowing to achieve better performance for a variety of access patterns.
+\lr{Talk about ESDM!}
 
@@
 As an example, in \cref{pic/workflow-s3-n}, the outcome of Task\,1 may be used directly by Task\,2.
-To achieve decoupling between producer and consumer, it may also be kept in memory and cached.
+To achieve decoupling\lr{Decoupling in the direct-coupling session? Confusing to me.} between producer and consumer, data may also be kept in memory and cached.
 \end{description}
 
-In the design proposed in this work (\cref{sec:design}), we will focus on the data placement strategy.
+\lr{We talked about balancing the text for each item. I think there was no time for it, right?}
 
@@
 
-\item[Data Recomputation] Climate/weather scientists are trading recomputation with space usage manually.%, in the so-called field space-time trade-off that has been applied in Cryptanalysis for already forty years~\cite{1056220}.
+\item[Data Recomputation] Climate/weather scientists are trading recomputation with space usage manually.
+%, in the so-called field space-time trade-off that has been applied in Cryptanalysis for already forty years~\cite{1056220}.
 By knowing how to rerun the workflow behind the data creation, a smarter storage system can automatically trade data availability for potential recomputation opportunities to optimise the cost-efficiency of the system.
@@
 Consider \cref{pic/workflow-s3-n} again and that, at every K cycles of the workflow, the generated Product\,3 (from Cycle 1 to Cycle K) are used in a validation task, called here \texttt{check}.
-From the workflow\footnote{The $P_i C_j$ notation represents the Product\,$i$ generated in the Cycle $j$.}, we know that $P_3 C_1$ will be used to construct $P_3 C_2$ and then \texttt{check}.
-This dataset will probably be stored somewhere, and it will not be used until the workflow reaches the K-th cycle.
+From the workflow, we know that $P_3 C_1$\footnote{The $P_i C_j$ notation represents the Product\,$i$ generated in the Cycle $j$.} will be used to construct $P_3 C_2$ and then \texttt{check}.
+This dataset would probably be stored somewhere, and it will not be used until the workflow reaches the K-th cycle.
 One alternative is to delete it after it was first used and then recompute it when time is right.
 The cost of doing that is storing \texttt{checkpoint} and then use it to reconstruct $P_3 C_1$.
-If, for instance, $P_3 C_1$ is a large dataset, \texttt{checkpoint} is small, and computing time is short.
-It is easy to see that deleting and recomputing it may improve the costs of running a workflow.
+If, for instance, $P_3 C_1$ is a large dataset, \texttt{checkpoint} is small, and computing time is short, it is easy to see that deleting and recomputing it may improve the costs of running the workflow.
 %We need, then, information about datasets volume and computing time to be able to make an accurate decision.
 %The bottom line here is: there are situations in which decisions are being made without considering all the available %information and possible scenarios.
-This is just an example, and, currently, scientists perform optimisations manually.
+This is just an example, and, currently, scientists perform those optimisations manually.
 \end{description}
@@
 
-\item[Performance-portability] With both abstraction and optimisation, the user can specify the I/O requirements only once for the tasks of a specific workflow, and the workflow can now run on any system without user intervention.
+\item[Performance-portability] With both abstraction and optimisation, the user can specify the I/O requirements only once for the tasks of a specific workflow, and the I/O-aware workflow can now run with optimised data storage on any system without user intervention.
 Even more, if the system characteristics change, e.g., it gets upgraded, an additional storage tier becomes available, or if storage degrades, the I/O-aware workflow could automatically adapt and make use of this new environment.

### 2020-02-25 paper.tex
Chapters 4 and 5.

@@
 Implemented with a standalone API, it also provides NetCDF integration allowing usage in existing applications.
+Hence, in \cref{pic/layers-xios}, the HDF5 layer can be replaced with ESDM.
 
-\lr{We shoud talk about how ESDM would replace the components in Fig 3. That was in the text before and it's important.}
+\lr{I added the info of ESDM here, but also keep it in the original place (line 712).}
 
@@
 
-\item Adapt the resulting workflow, individual scripts and application executions to consider the potential for data placement and migration.
+\item Adapt the resulting workflow, individual scripts and application executions to consider the potential for data placement and migration.\lr{data placement and migration again.}
 
@@
 
-ESDM is used as I/O middleware in the parallel application (in NetCDF or directly) and orchestrates the I/O according to a configuration file.
-The information about the system is provided using a simplified ESDM configuration file (\texttt{esdm.conf}) provided by the data centre (or expert users).
+ESDM is used as I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
+% The information about the system is provided using a simplified ESDM configuration file (\texttt{esdm.conf}) provided by the data centre (or expert users).
 This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.
-
 In the example presented in \Cref{lst:esdm.conf}, we have three storage targets: two global accessible file systems (\texttt{lustre01} and \texttt{lustre02}), and one local file system in \texttt{/tmp} that can be accessed via the POSIX backend.
@@
 The number of threads is defined in the configuration file.
-As an example, based on the number of Object Storage Targets (OSTs) available to both Lustre systems at DKRZ, performance tests already developed by ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
+As an example, based on the number of Object Storage Targets (OSTs) available to\lr{on?} both Lustre systems at DKRZ, performance tests already developed using ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
 %; more degrade it.
@@
 The configuration file is inquired by an application that utilises ESDM and steers the distribution of data during I/O.
-While the current system information and performance model is based on latency and throughput only, it shows that automatic decision making can be made on behalf of the user.
+While the current system information and performance model are based on latency and throughput only, it shows that automatic decision making can be made on behalf of the user.
 
@@
 
-The user now has to provide information about datasets required for input and the generated output for each Cylc task in a separate file similarly to Cylc's workflow configuration file.
-An example of an I/O-workflow configuration file is shown in \Cref{lst:cylc}.
+The user now has to provide information about datasets required for input and the generated output for each workflow task in a separate file similarly to Cylc's workflow configuration file.
+An example of an I/O-workflow configuration file (\texttt{io.cylc}) is shown in \Cref{lst:cylc}.
 In this file, information about Task\,1 is given by example, and we expect the extra information about all tasks in the same file.
@@
 
-\begin{lstlisting}[caption = External Cylc I/O-workflow configuration file, label=lst:cylc]
+\begin{lstlisting}[caption = External Cylc I/O-workflow configuration file (\texttt{io.cylc}), label=lst:cylc]
 [Task 1]
@@
 
-In this example, the workflow file could define a cycle flexibly to be a month or a year according to the Cylc workflow file.
+%In this example, the workflow file could define a cycle flexibly to be a month or a year according to the Cylc workflow file.
+In this example, the \texttt{io.cylc} file could define a cycle flexibly to be a month or a year according to the \texttt{flow.cylc} file.\lr{Again, where is this info?
+File flow.cylc?
+io.cylc?
+I'm rewriting it in the way I'm understanding.}
 The notation is similar to the specification of Cylc workflows using a nested INI format.
@@
 Line~3, for example, defines that the filename \texttt{topography} is mapped to a specific input file.
-This dataset does not depend on any previous step of the workflow.
+This dataset \lr{Is dataset the right term here?} does not depend on any previous step of the workflow.
 The next line specifies that the input filename \texttt{checkpoint} should be mapped to the output of Task~1 checkpoint dataset from the previous cycle (e.g., the checkpoint generated after completed the last year's output).
@@
 
-Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated by modifying the "filename" which is then used by the ESDM-aware application to determine the I/O scheduler.
+% Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated by modifying the filename, which is then used by the ESDM-aware application to determine the I/O scheduler.
+
+\lr{Check it, please.}
+
+Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated from EIOS through a modified filename, which is then used by the ESDM-aware application to determine the final workflow schedule.
 
@@
 
-The steps to execute a workflow enriched with I/O information and performing smarter scheduling is depicted in \cref{pic/stages-io}.
+The steps to execute a workflow enriched with I/O information and perform smarter scheduling is depicted in \cref{pic/stages-io}.
 The suggested alterations can be seen in boxes pointed by red arrows, and the remaining components are the current state-of-the-art for workflows in climate and weather from \cref{pic/stages}.
@@
 
-\fig{width=0.7\columnwidth}{pic/stages-io}{Software stack and stages of execution with our I/O-aware scheduler -- EIOS}
+\fig{width=0.7\columnwidth}{pic/stages-io}{Software stack and stages of execution with the I/O-aware scheduler (EIOS)}
 
@@
 
-  \item \textbf{Scientist} The end-user now has to provide an additional file that covers the I/O information for each task and slightly changes (discussed above) have to be made to the current scripts.
+  \item \textbf{Scientist} The end-user \lr{Is the end-user the scientist?} now has to provide an additional file that covers the I/O information for each task and slightly changes have to be made to the current scripts (discussed above).\lr{Where?
+I'm lost again.
+Do we change anything in the original scripts that the user has to provide?
+He has to provide an additional file and then everything will be under our control, no?
+The scripts will change, but not by the user, by EIOS.}
 
@@
 
-  \item \textbf{\color{red}{EIOS}} The ESDM I/O Scheduler reads the information about the workflow (original and I/O workflow configuration files) and acts depending on the stage of the execution.
+  \item \textbf{\color{red}{EIOS}} The ESDM I/O Scheduler reads the information about the workflow (\texttt{flow.cylc} and \texttt{io.cylc} configuration files) and acts depending on the stage of the execution.
   EIOS consists of several subcomponents:
@@
 
-      \item A tool to generate pseudo-file names used by the ESDM-aware applications.
+      \item A tool to generate pseudo filenames used by the ESDM-aware applications.
 
-      \item A data management service (not shown on the figure) that migrate and purge data at the end of the life cycle.
+      \item A data management service (not shown in the figure) that migrate and purge data at the end of the life cycle.
     \end{itemize}
@@
   \item \textbf{Slurm}
-  Cylc may now have added an optimisation identified by EIOS, and that is now handled by a modified Slurm.
+  Cylc may now have added an optimisation identified by EIOS which is now handled by a modified Slurm.
   Also, if migrations have to be performed, Slurm will administer them according to the specification in the job script.
@@
   \item \textbf{Job}
-  A job might run on the same nodes than a previous job to utilise local storage.
+  A job might run on the same node than a previous job to utilise local storage.
   %\bnl{Any interesting that needs this performance is running on dozens if not hundreds of nodes}
+  \lr{Only that might happen?}
 
@@
   ESDM loads the \texttt{esdm.conf} file that contains the information about the available storage backends and their characteristics.
-  ESDM extracts the long-term schedule information from the generated pseudo filenames and employs it during the I/O scheduling to optimise the storage considering data locality between tasks.
+  ESDM extracts the long-term schedule information from the generated pseudo filenames and employs it during the I/O scheduling to optimise the storage considering data locality between tasks.\lr{Is that really the only thing we are exploring?}
+
+  \lr{In this happening in the application level? Not before slurm? I'm confused now.}
 
@@
 In this work, we describe the general vision and a specific design for the software stack in the domain in climate and weather that we work on in the ESiWACE project.
-The proposed changes increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system in a workflow.
+The proposed changes increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system \lr{architecture, maybe?} in a workflow.
 

### 2020-02-25 paper.tex
Dealt with some comments.

@@
 
-\lr{JT, Hua ?}
-
 \title{Potential of I/O-Aware Workflows in Climate and Weather}
@@
 The key contributions of this paper are:
-1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of mostly existing tools (Cylc, XIOS, DDN IME, and ESDM).
-
-\lr{mostly??}
+1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of existing tools (Cylc, XIOS, DDN IME, and ESDM).
 
 The vision proposed here is built on the belief that workflows composed of data, computing, and communication-intensive tasks should drive interfaces and hardware configurations to best support the programming models.
-When delivered, this work will increase the opportunity for smarter scheduling of computing and storage in heterogeneous storage systems.
-
-\lr{computing and storage?? We talked about it before. Not only storage? I think the Cylc team spotted this problem too.}
+When delivered, this work will increase the opportunity for smarter scheduling of computing by considering storage in heterogeneous storage systems.
 
@@
 
-%\bnl{I don't believe people port workflows, they port simulations, and they write and move data ... we want to be able to port workflows ... }
-%\jk{I believe they do port aspects of workflows as scripts must be adjusted (where is data and such). Cylc team to reply}
-%\lr{It's not clear to me the difference between workflows and simulations. I can imagine that a simulation is a subset of a workflow, right? So, jobs, in terms of computer science?}
-
 Since performance is of crucial importance to large-scale workflows, careful attention must be paid to exploit the system characteristics of the target computing centre.
 For instance, a data-driven workflow may benefit from the explicit and simultaneous use of a locally heterogeneous set of computing and storage technologies.
-\lr{computing and storage technologies again.}
 This means that substantial changes may be required to a workflow to tailor it to a particular supercomputer environment in order to obtain the best performance.
 
-\lr{For instance, a data-driven workflow may benefit from the explicit and simultaneous use of a locally heterogeneous set of computing and storage technologies.
-For example, porting a workflow from one system to another still requires adjusting runtime parameters of applications and deciding on how data is managed.}
-
-\lr{What's the relation between those two examples? They should be together, no?}
+%For example, porting a workflow from one system to another still requires adjusting runtime parameters of applications and deciding on how data is managed.}
 
 Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
-As the complexity of systems expands and alternative storage and computing technologies \lr{computing and storage technologies again.} provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise the usage of resources in workflows.
+As the complexity of systems expands and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise the usage of resources in workflows.
 In many cases, modifications are not performed because: 1) They are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) It is a one-time explorative workflow and, 3) Users are not aware of the potential of the complex system.
@@
 
-In this section, we describe how workflows are executed in a typical software stack \lr{software stack == climate and weather application?
-The section is not only about the software stack.} and the standard I/O stack involved in running a single application.
+In this section, we describe how workflows are executed in a typical software stack and the typical hardware and software environment involved in running an application.
 
@@
 
-  \item \textbf{Scientist} specifies the workflow and provides scripts for each task.
+  \item \textbf{Scientist} specifies the workflow and provides a command or a script for each task.
+  As part of the Cylc configuration, the command(s) to be run, any environment variables used by these application(s), and any workload manager directives.
   After that, the user enacts Cylc to start the workflow.
@@
 
-  \lr{I rewrite this part.}
+  \item \textbf{Job} provides the environment with the resources and it runs the user-provided program or script on one of the nodes allocated for it.
+  Local variables containing information about the environment of the batch job, e.g., the compute nodes allocated, and then enacts the Cylc provided script on one node.
 
-  \item \textbf{Job} provides the environment with the resources and it runs the user-provided script on one of the nodes allocated for it.
-  Local variables containing information about the environment of the batch job, e.g., the compute nodes allocated, enact the Cylc provided script on that node.
-
-  \item \textbf{Script} executes the shell commands with one or multiple (potentially parallel) applications to run sequentially.
+  \item \textbf{Script} executes the commands with one or multiple (potentially parallel) applications to run sequentially.
   During the creation of the script, Cylc has included variables that describe the task in the workflow.
   The information is typically fed into the application(s) representing the task, and so defines the storage location.
-  The user script may use\lr{May use?
-"Application" talks about generated filenames.} commands to create a filename considering the cycle or \lr{or??} store data in a workflow-specific shared directory.
+  The user script may use commands to generate filenames considering the cycle and may store data in a workflow-specific shared directory.
 
@@
 It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF4~\cite{netcdf} to store data as a file.
-Under the hood, NetCDF4 uses the HDF5 API and file format. \lr{and file format? What does that mean?}
+Under the hood, NetCDF4 uses the HDF5 API with its file format.
 Internally, HDF5 uses MPI and its data types to specify the nature of the data stored.
 Finally, data is stored on a parallel file system like Lustre which, on the server-side, stores data in a local file system on block devices such as SSDs and HDDs.
-
-\lr{NetCDF4 or NetCDF? We should be consistent in the whole paper.}
 
@@
 Hence, in \cref{pic/layers-xios}, the HDF5 layer can be replaced with ESDM.
-
-\lr{I added the info of ESDM here, but also keep it in the original place (line 712).}
 
@@
 
-\lr{Are SSD and HDD online storage?}
-
 \fig{width=0.6\columnwidth}{pic/system}{Example of an heterogeneous HPC landscape}
@@
 
-If a user \lr{It's really the user here?} knows the workflow and the system characteristics, s/he can optimise data placement decisions.
+If a user knows the workflow and the system characteristics, s/he can optimise data placement decisions.
 Consider, for instance, the situation where each computing node has access to three file systems: a fast \texttt{scratch} file system on which data may reside only for a week, a slower \texttt{work} file system, and a \texttt{local} file system.
@@
 %\texttt{local} is not used.
-In this situation, there are many obvious opportunities to utilise data migration to optimise performance, and many possible optimisation criteria (e.g., costs). \lr{Another criterion?}
+In this situation, there are many obvious opportunities to utilise data migration to optimise performance, and other criteria (e.g., costs).
 However, with a multitude of file systems that differ at each data centre, such optimisations would be difficult to achieve manually by users.
@@
 The storage community had also adjusted various higher-level software to support storage tiering on top of several storage systems.
-For instance, ADIOS provides in-memory staging that had been exploited by applications~\cite{slawinska2013maya}.
-\lr{What kind of applications?
-Medical, climate, industrial, etc.}
+For instance, ADIOS provides in-memory staging that had been exploited by a variety of applications~\cite{slawinska2013maya}.
 Hermes~\cite{kougkas2020acceleration} provides a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
-iRODS~\cite{rajasekar2010irods} is a rule-oriented data system that allows scientists to organise data into shareable collections and provide metadata such as information about the data life cycle.
-\lr{such as information about the data life cycle?
-I don't get it.}
-The system provides several patterns for workflows that consider data locality and data migration/replication. \lr{Which system? Is still talking about iRODS?}
+iRODS~\cite{rajasekar2010irods} is a rule-oriented data system that allows scientists to organise data into shareable collections and provide metadata such as the data life cycle.
+\lr{such as information about the data life cycle?}
+iRODS provides several patterns for workflows that consider data locality and data migration/replication.
 
@@
 \textbf{Workflows. }
-A good overview of the flavours \lr{flavours?} of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
+A good overview of the flavours of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
 The article states that SWfMS should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources.
@@
 Those scalable workflow solutions typically utilise a container solution to allow execution in an arbitrary software environment.
-Popper~\cite{jimenez2017popper}, Snakemake~\cite{bts480}, and Nextflow~\cite{Nextflow} provide a language \lr{Is it a DSL?
-I think it's important to emphasize this here.} to specify workflows and to execute them.
+Popper~\cite{jimenez2017popper}, Snakemake~\cite{bts480}, and Nextflow~\cite{Nextflow} provide a language to specify workflows and to execute them.
 Snakemake is interesting as it supports to define and infer input and output filenames.
@@
 
-%\lr{Towards Decoupling the Selection of Compression Algorithms from Quality Constraints: An Investigation of Lossy Compression Efficiency~\cite{TDTSOCAFQC17}}
-
 \item[Data Recomputation] Climate/weather scientists are trading recomputation with space usage manually.
@@
 Still to consider.}
-%\lr{I like this idea a lot. We need to talk to the Cylc team and see what they think.}
 
@@
 The number of threads is defined in the configuration file.
-As an example, based on the number of Object Storage Targets (OSTs) available to\lr{on?} both Lustre systems at DKRZ, performance tests already developed using ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
+As an example, based on the number of Object Storage Targets (OSTs) provided by both Lustre systems at DKRZ, performance tests already developed using ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
 %; more degrade it.
@@
 Since ESDM also supports several (non-POSIX) storage backends, an application can utilise all available storage systems without any modifications to the code.
-
-%\bnl{This has me asking all sorts of questions: how does one learn the right information to put in the ESDM configuration, e.g.
-your 200.
-What do threads ``per node'' mean?
-What sort of node?
-Is the ESDM running as an I/O server, or are you assuming it is being used in an I/O server etc}
-%\lr{Well noticed.
-For me, ESDM would be responsible for finding those optimal numbers automatically.
-But then we had this sentence: ``The number of threads is defined in the configuration file.'', meaning that the user has to know.
-So I don't have an answer to it.}
 
@@
 %In this example, the workflow file could define a cycle flexibly to be a month or a year according to the Cylc workflow file.
-In this example, the \texttt{io.cylc} file could define a cycle flexibly to be a month or a year according to the \texttt{flow.cylc} file.\lr{Again, where is this info?
-File flow.cylc?
-io.cylc?
-I'm rewriting it in the way I'm understanding.}
+In this example, the \texttt{io.cylc} file could define a cycle flexibly to be a month or a year according to the \texttt{flow.cylc} file.
 The notation is similar to the specification of Cylc workflows using a nested INI format.
@@
 Line~3, for example, defines that the filename \texttt{topography} is mapped to a specific input file.
-This dataset \lr{Is dataset the right term here?} does not depend on any previous step of the workflow.
+This dataset does not depend on any previous step of the workflow.
 The next line specifies that the input filename \texttt{checkpoint} should be mapped to the output of Task~1 checkpoint dataset from the previous cycle (e.g., the checkpoint generated after completed the last year's output).
@@
 
-% Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated by modifying the filename, which is then used by the ESDM-aware application to determine the I/O scheduler.
-
-\lr{Check it, please.}
-
-Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated from EIOS through a modified filename, which is then used by the ESDM-aware application to determine the final workflow schedule.
+Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated from EIOS through a modified filename, which is then used by the ESDM-aware application to determine the data placement.
 
@@
   \item \textbf{Cylc} EIOS is invoked by Cylc to identify potential optimisations in the schedule before generating the Slurm script.
-  %\hh{Cylc has to be extended to call EIOS? How is this achieved?}
-  %\lr{We talked to the Cylc team, and they will deal with it, I suppose.}
-  %\jk{Modifying Cylc, Cylc team to decide, not yet decided how that will work, Python code.....}
 
@@
   Also, if migrations have to be performed, Slurm will administer them according to the specification in the job script.
-  %\lr{Already inserted in the idea of the previous sentence.}
-  %\lr{It's clear the idea about what's really going on in each part is still not clear.}
 
@@
 In this work, we describe the general vision and a specific design for the software stack in the domain in climate and weather that we work on in the ESiWACE project.
-The proposed changes increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system \lr{architecture, maybe?} in a workflow.
+The proposed changes increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system architecture in a workflow.
 
@@
 This project is funded by the European Union's Horizon 2020 research and innovation programme under grant agreement No. 823988.
-We thank our collaborators Bryan Lawrence, Glenn Greed, and David Matthews for their input to this article.
+We thank our collaborators Bryan Lawrence, Glenn Greed, David Matthews, and Hua Huang for their input to this article.
 

### 2020-02-25 paper.tex
Plus.

@@
 %That's an exascale use case requiring not only a big machine but the ability to schedule big jobs.
-
-\lr{I don't quite understand the numbers.}
+%\lr{I don't quite understand the numbers.}
 
@@
 
+\fig{width=0.8\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file}
+
 Consider the Cylc workflow for a toy monthly cycling workflow in \cref{pic/cylc1}.
@@
 In this workflow, the task \texttt{model} is followed by tasks postprocessing (\texttt{post}), forecast verification (\texttt{ver}), and product generation (\texttt{prod}), all specified as a workflow in a Cylc configuration file (\texttt{flow.cylc}).
-
-\fig{width=0.8\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file}
 
@@
 
-\fig{width=0.2\textwidth}{pic/layers-xios}{I/O path for an MPI-parallel application}
+% \fig{width=0.2\textwidth}{pic/layers-xios}{I/O path for an MPI-parallel application}
+
+\lr{Figs 3 and 4 together as an option.}
+
+\begin{figure}[b]
+    \begin{minipage}{.42\linewidth}
+        \centering
+        \includegraphics[width=0.42\textwidth]{pic/layers-xios}
+        \subcaption{I/O path for an MPI-parallel application}\label{pic/layers-xios}
+    \end{minipage}
+    \begin{minipage}{.58\linewidth}
+        \centering
+        \includegraphics[width=1\textwidth]{pic/system}
+        \subcaption{Example of an heterogeneous HPC landscape}\label{pic/system}
+    \end{minipage}
+    \vspace{5pt}
+    \caption{I/O stack and data centre infrastructure}
+    \label{fig:34}
+\end{figure}
 
@@
 
-\fig{width=0.6\columnwidth}{pic/system}{Example of an heterogeneous HPC landscape}
-
-\lr{Consider put figs 3 and 4 together to better use the space.}
+% \fig{width=0.6\columnwidth}{pic/system}{Example of an heterogeneous HPC landscape}
 
@@
 A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
-File systems and data management software such as Spectrum Scale, HPSS, BeeGFS, \lr{ref about those topics?
-webpage as footnote, maybe?} and Lustre (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to administrator-provided policies.
+File systems and data management software such as IBM Spectrum Scale~\cite{SchmuckH02}, HPSS~\cite{528214}, BeeGFS~\cite{beegfs}, and Lustre~\cite{abs-1903-01955} (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to administrator-provided policies.
 However, the semantic information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.
@@
 Hermes~\cite{kougkas2020acceleration} provides a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
-iRODS~\cite{rajasekar2010irods} is a rule-oriented data system that allows scientists to organise data into shareable collections and provide metadata such as the data life cycle.
-\lr{such as information about the data life cycle?}
-iRODS provides several patterns for workflows that consider data locality and data migration/replication.
-
+iRODS~\cite{rajasekar2010irods} is a rule-oriented data system that allows scientists to organise data into shareable collections and provides several patterns for workflows considering data locality and data migration/replication.
+% and provide metadata such as the data life cycle.
 Finally, there have also been extensions to batch schedulers to perform data staging for utilising node-local storage, for example, NORNS~\cite{miranda2019norns} as an extension to Slurm.
@@
 We aim to enhance the current information with characteristics for input and output, i.e., the datasets.
-An example workflow containing input datasets and (intermediate) products is illustrated in \cref{pic/workflow-s3-n}.
+An example workflow with \texttt{N} cycles containing input datasets and (intermediate) products is illustrated in \cref{pic/workflow-s3-n}.
 Round nodes represent tasks, squared nodes represent data, and arrows indicate dependencies.
@@
 The idea here is to embrace the concept that tasks dependencies are really imposed by datasets dependencies.
-
-\lr{Talk a little bit more about the figure and the N cycles.}
 
@@
 Intermediate states could be rerun by utilising virtualisation and container technologies.
-Consider \cref{pic/workflow-s3-n} again and that, at every K cycles of the workflow, the generated Product\,3 (from Cycle 1 to Cycle K) are used in a validation task, called here \texttt{check}.
+Consider \cref{pic/workflow-s3-n} again and that, at every \texttt{K} cycles of the workflow, the generated Product\,3 (from Cycle 1 to Cycle \texttt{K}) are used in a validation task, called here \texttt{check}.
 From the workflow, we know that $P_3 C_1$\footnote{The $P_i C_j$ notation represents the Product\,$i$ generated in the Cycle $j$.} will be used to construct $P_3 C_2$ and then \texttt{check}.
-This dataset would probably be stored somewhere, and it will not be used until the workflow reaches the K-th cycle.
+This dataset would probably be stored somewhere, and it will not be used until the workflow reaches the \texttt{K}-th cycle.
 One alternative is to delete it after it was first used and then recompute it when time is right.
@@
 This project is funded by the European Union's Horizon 2020 research and innovation programme under grant agreement No. 823988.
-We thank our collaborators Bryan Lawrence, Glenn Greed, David Matthews, and Hua Huang for their input to this article.
+We thank our collaborators Bryan Lawrence, Glenn Greed, David Matthews, and Hua Huang for their input to this paper.
 

### 2020-02-26 paper.tex
Further comments.

@@
 Related work in heterogeneous storage environments and solutions for workflow processing are presented in \Cref{sec:art}.
-Next, the vision for including knowledge about data requirements and characteristics \lr{data requirements and characteristics == data management?} is sketched in \Cref{sec:vision} outlining the potential benefit the automatic exploitation might bring.
+Next, the vision for including knowledge about data requirements and characteristics is sketched in \Cref{sec:vision} outlining the potential benefit the automatic exploitation might bring.
 Our design, based on existing components in climate and weather, is described in \Cref{sec:design}.
@@
 
-  \item \textbf{Application} is executed taking the generated filenames set by the script.
+  \item \textbf{Application} is executed taking the filenames set by the script.
 \end{enumerate}
@@
 % \fig{width=0.2\textwidth}{pic/layers-xios}{I/O path for an MPI-parallel application}
-
-\lr{Figs 3 and 4 together as an option.}
 
@@
     \vspace{5pt}
-    \caption{I/O stack and data centre infrastructure}
+    \caption{Typical hardware and software environment for applications}
     \label{fig:34}
@@
 There are several approaches to implement the technology for the vision proposed in this work, and changes are needed in the software components to realise it.
-In the next sections, we will discuss a specific design for our transitional roadmap considering climate and weather workflows and tools scientists from this field already use in their routine research.
-
-\lr{"a specific design" here? Next section, no?}
+In \Cref{sec:design}, we will discuss a specific design for our transitional roadmap considering climate and weather workflows and tools scientists from this field already use in their routine research.
 
@@
 It is expected that the data centre (or expert users) can create such a configuration file, e.g., by using vendor-provided information or by executing benchmarks.
-With this information, a scheduler can make data placement, transformation, and migration decisions for individual datasets during their life cycle.\lr{The scheduler making data transformation?
-And it's migration already classified as data placement?}
+With this information, a scheduler can make the initial data placement, transformation, and migration decisions for individual datasets during their life cycle.
 This separation of concerns allows us to abstract from the workflow what is essential and what a system should optimise to ensure smart usage of the available resources.
@@
 For instance, in \cref{pic/workflow-s3-n}, alternating the storage location for Product\,2 into two scratch file systems is something that would be a simple job for an I/O-aware scheduler.
-However, currently, that implies having at least two scripts for that task with information about the different storage placement.
+However, currently, that implies providing scripts for that task and all tasks depending on the data with information about the different storage placement.
 %Next, we exemplify three situations in which data placement can be optimised.
 
-\lr{Weak justification. We are better than that.}
-
-\item[Data Migration] is the process of transferring data from one storage system to another.
-Data movement involves a significant overhead, both in terms of latency and energy-efficient computing, hence needs to be considered carefully.
+\item[Data Migration] is the process of transferring data from one storage system to another; typically this involves to delete the data but this decision can be delayed to provide read access on multiple storage systems.
+Data movement involves a significant overhead as data must be read on one storage and written to another, both in terms of latency and energy-efficient computing, hence needs to be considered carefully.
 \cref{fig:lifecycle} introduces three possible life cycles for a specific dataset and explains how migrations can be done to improve datasets accessibility.
@@
 
-\lr{If we are really migrating, why do we have to deletes in (a) and (c)? Maybe the delete should appear at the same time as the migration in the timeline.}
-
 \item[Data Replication] in computing involves sharing information to ensure consistency between redundant resources, such as software or hardware components, to improve reliability, fault-tolerance, or accessibility.
@@
 In addition, the system may combine the replication of data by \textbf{transforming} the data into a different representation allowing to achieve better performance for a variety of access patterns.
-\lr{Talk about ESDM!}
 
@@
 As an example, in \cref{pic/workflow-s3-n}, the outcome of Task\,1 may be used directly by Task\,2.
-To achieve decoupling\lr{Decoupling in the direct-coupling session? Confusing to me.} between producer and consumer, data may also be kept in memory and cached.
+To achieve some level of independence between producer and consumer, data may also be kept in memory and cached.
 \end{description}
-
-\lr{We talked about balancing the text for each item. I think there was no time for it, right?}
 
@@
 
-\item Adapt the resulting workflow, individual scripts and application executions to consider the potential for data placement and migration.\lr{data placement and migration again.}
-
+\item Adapt the resulting workflow, individual scripts and application executions to consider the potential for data placement.
 \end{enumerate}
@@
 
-  \item \textbf{Scientist} The end-user \lr{Is the end-user the scientist?} now has to provide an additional file that covers the I/O information for each task and slightly changes have to be made to the current scripts (discussed above).\lr{Where?
-I'm lost again.
-Do we change anything in the original scripts that the user has to provide?
-He has to provide an additional file and then everything will be under our control, no?
-The scripts will change, but not by the user, by EIOS.}
+  \item \textbf{Scientist} The user now has to provide an additional file that covers the I/O information for each task and slightly changes have to be made to the current scripts.
 
@@
   %\bnl{Any interesting that needs this performance is running on dozens if not hundreds of nodes}
-  \lr{Only that might happen?}
 
@@
   ESDM loads the \texttt{esdm.conf} file that contains the information about the available storage backends and their characteristics.
-  ESDM extracts the long-term schedule information from the generated pseudo filenames and employs it during the I/O scheduling to optimise the storage considering data locality between tasks.\lr{Is that really the only thing we are exploring?}
-
-  \lr{In this happening in the application level? Not before slurm? I'm confused now.}
-
+  ESDM extracts the long-term schedule information from the generated pseudo filenames and employs it during the I/O scheduling to optimise the storage considering data locality between tasks.
+  Basically, ESDM can now change the priorities for data placement of the different storage locations that would normally be encoded in its configuration file.
 \end{enumerate}

### 2020-02-26 paper.tex
Minor

@@
 This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
-In the approach proposed in this work, each task is annotated with the required input datasets and the generated products must include metadata such as data life cycle information, value of data, and how long it should be kept.
+In the approach proposed in this work, each task is annotated with the required input datasets and the generated products must include metadata such as data life cycle information, the value of data, and how long it should be kept.
 The idea here is to embrace the concept that tasks dependencies are really imposed by datasets dependencies.
@@
 
-\item[Data Migration] is the process of transferring data from one storage system to another; typically this involves to delete the data but this decision can be delayed to provide read access on multiple storage systems.
+\item[Data Migration] is the process of transferring data from one storage system to another.
+Typically, it involves to delete the data, but this decision can be delayed to provide read access to multiple storage systems.
 Data movement involves a significant overhead as data must be read on one storage and written to another, both in terms of latency and energy-efficient computing, hence needs to be considered carefully.
@@
 The suggested alterations can be seen in boxes pointed by red arrows, and the remaining components are the current state-of-the-art for workflows in climate and weather from \cref{pic/stages}.
-
 In the following, we describe the modifications we propose in this vision paper for each component involved in the software stack.
@@
 
-  \item \textbf{Scientist} The user now has to provide an additional file that covers the I/O information for each task and slightly changes have to be made to the current scripts.
+  \item \textbf{Scientist} The user now has to provide an additional file that covers the I/O information for each task and slight changes have to be made to the current scripts.
 
@@
   \item \textbf{Job}
-  A job might run on the same node than a previous job to utilise local storage.
+  A job might run on the same node as a previous job to utilise local storage.
   %\bnl{Any interesting that needs this performance is running on dozens if not hundreds of nodes}

### 2020-02-26 paper.tex
Clarify Cylc environment variables.

@@
   The information is typically fed into the application(s) representing the task, and so defines the storage location.
-  The user script may use commands to generate filenames considering the cycle and may store data in a workflow-specific shared directory.
+  The script uses commands to generate filenames considering the cycle and may store data in a workflow-specific shared directory.
+  Either these commands are set in the Cylc workflow and then injected as environment variables or directly utilised as part of a user-provided script.
 

### 2020-02-26 paper.tex
Fin

@@
 
-\item[Abstraction] The user does not have to know the architecture of the target system on which the workflow will run, removing the specialist from the decision-making process.
+\item[Abstraction] Providing the abstraction that enables a separation of concerns.
+Once the IO characteristics for a workflow is defined, the user does not have to know the architecture of the target system on which the workflow will run.
+Thus, removing the specialist from the decision-making process of individual workflows.
 
@@
 This project is funded by the European Union's Horizon 2020 research and innovation programme under grant agreement No. 823988.
-We thank our collaborators Bryan Lawrence, Glenn Greed, David Matthews, and Hua Huang for their input to this paper.
+We thank our collaborators Bryan Lawrence, Glenn Greed, David Matthews, and Hua Huang for their input to this paper, and the NGI initiative for contributions to the vision.
 

### 2020-05-02 paper.tex
Minor fixes

@@
   \item \textbf{Job} provides the environment with the resources and it runs the user-provided program or script on one of the nodes allocated for it.
-  Local variables containing information about the environment of the batch job, e.g., the compute nodes allocated, and then enacts the Cylc provided script on one node.
+  Local variables containing information about the environment of the batch job, e.g., the compute nodes allocated, enacts the Cylc provided script on the node.
 
@@
 1) Technology that exploits heterogeneous storage environments and supports user-directed policies and
-2) Solutions for workflow processing.
+2) Solutions for workflow processing.\\
 
@@
 % and provide metadata such as the data life cycle.
-Finally, there have also been extensions to batch schedulers to perform data staging for utilising node-local storage, for example, NORNS~\cite{miranda2019norns} as an extension to Slurm.
+Finally, there have also been extensions to batch schedulers to perform data staging for utilising node-local storage, for example, NORNS~\cite{miranda2019norns} as an extension to Slurm.\\
 
-\smallskip
+%\smallskip
 
@@
 
-\begin{lstlisting}[caption = Example of an ESDM configuration file (\texttt{esdm.conf}), label=lst:esdm.conf]
+\begin{figure}[!ht]
+
+\begin{lstlisting}
 "backends": [
@@
 \end{lstlisting}
+
+\caption{Example of an ESDM configuration file (\texttt{esdm.conf})}
+\label{lst:esdm.conf}
+\end{figure}
 
@@
 
-\begin{lstlisting}[caption = External Cylc I/O-workflow configuration file (\texttt{io.cylc}), label=lst:cylc]
+\begin{figure}[!ht]
+
+\begin{lstlisting}
 [Task 1]
@@
 \end{lstlisting}
+
+\caption{External Cylc I/O-workflow configuration file (\texttt{io.cylc})}
+\label{lst:cylc}
+\end{figure}
 
@@
 
-  \item \textbf{\color{red}{EIOS}} The ESDM I/O Scheduler reads the information about the workflow (\texttt{flow.cylc} and \texttt{io.cylc} configuration files) and acts depending on the stage of the execution.
+  \item \textbf{EIOS} The ESDM I/O Scheduler reads the information about the workflow (\texttt{flow.cylc} and \texttt{io.cylc} configuration files) and acts depending on the stage of the execution.
   EIOS consists of several subcomponents:
@@
 
-\section{Conclusions}
+\section*{Conclusions}
 \label{sec:conclusions}

### 2020-05-04 paper.tex
Minor

@@
   \item \textbf{Job} provides the environment with the resources and it runs the user-provided program or script on one of the nodes allocated for it.
-  Local variables containing information about the environment of the batch job, e.g., the compute nodes allocated, enacts the Cylc provided script on the node.
+  Local variables contain information about the environment of the batch job, e.g., the compute nodes allocated, and they enact the Cylc provided script on the node.
 

### 2020-05-22 paper.tex
Cleaning

@@
 
-\newcommand{\jk}[1]{\todo[inline]{JK: #1}}
-\newcommand{\lr}[1]{\textcolor{cyan}{LR: #1}}
-\newcommand{\hh}[1]{\textcolor{blue}{HuangHua: #1}}
-\newcommand{\bnl}[1]{\textcolor{green}{BNL: #1}}
 \pagestyle{plain}
@@
 A service such as the operational weather forecast workflow in Met Office writes around 200\,TB and reads around 600\,TB every day.
-%For a typical Met Office Unified Model (UM) from the Coupled Model Intercomparison Project (CMIP) run, some %rough estimates of data volumes are:
-%\begin{itemize}
-%\item N96-ORCA1 -- 200Gb per model year
-%\item N216-ORCA025 -- 700Gb per model year
-%\end{itemize}
-%Historical data spans over 170 years, generating 34Tb for low resolution and 119Tb for high resolution, while %projections will generate approximately half of this.
 In total, at the Met Office, on average 1.5\,PB and 14\,PB are written and read per day, respectively, for all climate and weather forecasts across all HPC clusters.
-%The above supports the view of how vital data management already is, and these numbers will grow with resolution and more ensembles.
-%Existing climate experiments do ensemble experiments sequentially as few institutions have resources for parallel climate ensembles.
-%That's an exascale use case requiring not only a big machine but the ability to schedule big jobs.
-%\lr{I don't quite understand the numbers.}
 
@@
 This means that substantial changes may be required to a workflow to tailor it to a particular supercomputer environment in order to obtain the best performance.
-
-%For example, porting a workflow from one system to another still requires adjusting runtime parameters of applications and deciding on how data is managed.}
 
@@
 Cylc~\cite{8675433} is in charge of executing and monitoring cyclic workflows in which each step is submitted to the batch scheduler of a data centre.
-%Cylc development began in 2008 in response to deficiencies of existing NWP workflow schedulers related to the creation of a single never-ending workflow rather than a never-ending sequence of single-cycle workflows.
-%With the current techniques, the evolution of the workflow could be a significant impediment
 With Cylc, tasks from multiple cycles may be able to run concurrently without violating dependencies, preventing the issue of delays that cause one cycle to run into another.
@@
 In a script file, the developers define, for each task, the parallelism settings and where the data is to be stored.
-%Decisions about storage must take into consideration the system architecture and the relevance of heterogeneous infrastructure in which the workflow will run.
 
@@
 In climate research, this process is repeated in the model to simulate years into the future.
-% Once the simulation of any month is computed, this task completes and implicitly the data for this month becomes available and can now be analysed.
 Once the simulation of any month is computed, the data for this month becomes available and can now be analysed.
@@
 The software stack involved in a general workflow is depicted in \cref{pic/stages}.
-%As a preliminary, the user has to specify a configuration representing the Cylc workflow and provide scripts for the individual tasks/applications to be executed.
 In the following, each stage of the execution is further described.
@@
   \item \textbf{Workload Manager} such as Slurm~\cite{Jette02slurm:simple} is responsible for allocating compute resources to a batch job and performing the job scheduling.
-%  In our case, it now queues the job that represents the Cylc task and plans its execution, considering the scheduling policy of the data centre.
   The selected tool queues the job that represents the Cylc task and plans its execution, considering the scheduling policy of the data centre.
   Once the job is scheduled to be dispatched, i.e., resources are available, and the job priority is the highest, it is started on the supercomputer.
-
-%  \item \textbf{Job} provides the environment with the resources it runs the user-provided script on one of the nodes allocated for the job.
-%  The job sets environment variables containing information about the environment of the batch job, e.g., the compute nodes allocated, and then enacts the Cylc provided script on one node.
 
@@
   \item \textbf{Application} is executed taking the filenames set by the script.
+
 \end{enumerate}
@@
 \subsection{I/O Stack of a Parallel Application}
-
-% \fig{width=0.2\textwidth}{pic/layers-xios}{I/O path for an MPI-parallel application}
 
@@
 \end{figure}
-
-% In our example, we assume the application is parallelised for performance reasons using MPI.
-% Not just for performance reasons, it's about making enough memory too, so removed qualification.
 
@@
 Depending on the need, the storage characteristics range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and cheap storage for long-term archival (tape).
-% Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy. % (\url{https://www.ddn.com/products/ime-flash-native-data-cache/})
 The tasks within any given workflow could benefit from utilising different combinations of storage and computing infrastructure.
-
-% \fig{width=0.6\columnwidth}{pic/system}{Example of an heterogeneous HPC landscape}
 
@@
 If the \texttt{scratch} filesystem reaches its capacity, the dataset would be moved back to \texttt{work}, and the task would continue running until it is finished, which might be inefficient.
-%\texttt{local} is not used.
 In this situation, there are many obvious opportunities to utilise data migration to optimise performance, and other criteria (e.g., costs).
@@
 iRODS~\cite{rajasekar2010irods} is a rule-oriented data system that allows scientists to organise data into shareable collections and provides several patterns for workflows considering data locality and data migration/replication.
-% and provide metadata such as the data life cycle.
 Finally, there have also been extensions to batch schedulers to perform data staging for utilising node-local storage, for example, NORNS~\cite{miranda2019norns} as an extension to Slurm.\\
-
-%\smallskip
 
@@
 The article states that SWfMS should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources.
-%"An abstract scientific workflow is defined by the functionality of each activity (or sub-workflow) and data dependencies between activities (or sub-workflows)"
 Existing solutions recognise challenges in data variety (formats of the input data), opportunities to optimise the schedule by moving code to data, allow specifying the data dependencies for tasks, and they even may consider the capacity of the available data storage.
@@
 In \cite{subedi2019leveraging}, an ML model that stages data for in-situ analysis by exploiting the access patterns is introduced.
-
-%TRIO: Burst Buffer Based I/O Orchestration. In: 2015 IEEE International conference on cluster computing, Chicago, IL, USA, 8–11 September 2015, pp. 194–203. Washington, DC: IEEE.
-%Chervenak, A.L., Smith, D.E., Chen, W., Deelman, E.:
-%Integrating policy with scientific workflow management
-%for data-intensive applications. In: Supercomputing (SC)
-%Companion: High-Performance Computing, Networking
-%Storage and Analysis, pp. 140–149 (2012)
-
-% https://www.sciencedirect.com/science/article/pii/S0167739X14002519
-%Workflow engine for cloud \cite{pandey2011workflow}
-%staging data to an available location just somehow.
-%"storage service provides a temporary repository for application files that is, input files that are required for task execution, and output files that are the result of execution.""
 
@@
 
-%The access to I/O information before the workflow runs can be used to optimise where jobs will be executed and how datasets will be stored during the workflow.
-%This information today is static and manual, leaving open opportunities for automating and optimisations.
-%Once we insert the datasets explicitly in the workflow, it is easier to see that many optimisations regarding storage can be made.
-
 \subsection{Smarter I/O Scheduling}
@@
 
-% \subsubsection{Data Placement}
-
 \paragraph{Strategy: Data Placement.}
@@
 Here we are considering the optimisation of data locality, where locality is twofold, spatial and temporal on a variety of characteristics.
-For optimising data placement, there are several approaches:%: \textbf{in-memory data}, \textbf{data allocation}, \textbf{data migration}, and \textbf{data replication}.
+For optimising data placement, there are several approaches:
 
@@
 However, currently, that implies providing scripts for that task and all tasks depending on the data with information about the different storage placement.
-%Next, we exemplify three situations in which data placement can be optimised.
 
@@
 In \cref{fig:lifecycle2}, the dataset is stored on the \texttt{scratch} file system immediately and accessed there.
-%However, that would require subsequent tasks to be placed on the same node where data is now stored, and the last read access must happen before files on \texttt{scratch} are automatically removed.
 However, the last read access must happen before files on \texttt{scratch} are automatically removed.
@@
 For the last two scenarios, subsequent tasks would have to be placed on the same node where previous data was stored.
-%However, it would require that the subsequent tasks are placed on the same node where data is now stored.
 
@@
 \item[Data Recomputation] Climate/weather scientists are trading recomputation with space usage manually.
-%, in the so-called field space-time trade-off that has been applied in Cryptanalysis for already forty years~\cite{1056220}.
 By knowing how to rerun the workflow behind the data creation, a smarter storage system can automatically trade data availability for potential recomputation opportunities to optimise the cost-efficiency of the system.
@@
 If, for instance, $P_3 C_1$ is a large dataset, \texttt{checkpoint} is small, and computing time is short, it is easy to see that deleting and recomputing it may improve the costs of running the workflow.
-%We need, then, information about datasets volume and computing time to be able to make an accurate decision.
-%The bottom line here is: there are situations in which decisions are being made without considering all the available %information and possible scenarios.
 This is just an example, and, currently, scientists perform those optimisations manually.
@@
 Even more, if the system characteristics change, e.g., it gets upgraded, an additional storage tier becomes available, or if storage degrades, the I/O-aware workflow could automatically adapt and make use of this new environment.
-% gain
 
@@
 ESDM is used as I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
-% The information about the system is provided using a simplified ESDM configuration file (\texttt{esdm.conf}) provided by the data centre (or expert users).
 This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.
@@
 
-%\hh{Here I have similar question: we need an extended and new configuration file, which includes characteristics of inputs/outputs dataset.
-As Julian said above, we can not added these items directly into Cylc configuration because Cylc wouldn't understand it.
-But you see, it's a burden to have another conf file.
-It is error-prune, and duplicated information may cause inconsistency.
-I propose to have these information added to the same Cylc configuration file, but in special comments.
-These comments will be omitted by Cylc, but they can be parsed by ESDM.
-This is similar for Doxygen comments in C/Java source file, or the leading line of a shell script, etc.}
-%\jk{Good suggestion!
-That absolutely depends how a Cylc file is parsed and analysed.
-Keeping it separately gives us the most flexibility and parsing comments could lead to misinterpreting already existing comments and acting wrongly.
-Still to consider.}
-
-%\hh{Other requirements of datasets may be: data block size/fragment size, 1-1 or N-1 access mode, IOPS intensive or throughput-intensive, etc.
-These will also be considered for data placement and migration.}
-%\jk{Fragment size/data blocks are handled by ESDM automatically. The access mode and access pattern is something that we could consider to add. However, difficult for scientists to judge}
-
 ESDM manages a pool of threads that should be created per compute node to achieve good performance and delegates the assignment of optimal block sizes to the storage backend.
@@
 As an example, based on the number of Object Storage Targets (OSTs) provided by both Lustre systems at DKRZ, performance tests already developed using ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
-%; more degrade it.
 To clarify the behaviour, ESDM distributes a single dataset across multiple storage devices depending on their characteristics.
@@
 
-%In this example, the workflow file could define a cycle flexibly to be a month or a year according to the Cylc workflow file.
 In this example, the \texttt{io.cylc} file could define a cycle flexibly to be a month or a year according to the \texttt{flow.cylc} file.
@@
   A job might run on the same node as a previous job to utilise local storage.
-  %\bnl{Any interesting that needs this performance is running on dozens if not hundreds of nodes}
 
@@
 \end{document}
-
-Machine Learning has been proven effective in replacing human-beings, and it may also adapt the workflow dynamically in case of unexpected situations (the system is down, nodes are unresponsive, etc.).
-(ref)
-
-From the recorded I/O accesses, we analyse the I/O patterns and dependencies associated with it to generate the configuration file.
-
-Potentially, EIOS may generate an \texttt{esdm.conf} file for the specific application to adjust the mapping of data to available storage targets.
-
-In particular, a system could store data that is of low value or cheap to reproduce on storage with little redundancy or trade storage space by the need to recompute data products.
-
-This scenario was developed to exemplify some of the potential optimisations we envision about \textbf{data placement}, \textbf{migration}, and \textbf{recomputation} in a heterogeneous storage environment.
-
-The user-specific storage locations are a viable option to run a workflow.
-
-The user might even have analysed various scenarios, such as depicted in Section \ref{sec:datamanagement}.
-
-Assuming the user can provide extra information in advance, ESDM can interact in the process before the user gives the data to the Cylc tool, using the gathered information to construct a tentative optimised workflow.
-
-Here, we focus on utilising a heterogeneous storage environment.
-
-Once we have access to the data pattern, the whole workflow can be improved by trading storage space by the need to recompute data products.
-
-With the right level of abstraction in place, potential data optimisations are unleashed in a heterogeneous storage system according to its characteristics and workflow specification.
-
-Exploiting the value and importance of data could improve the placement decisions on heterogeneous storage systems.
-
-Data systems usually do not respect, or have knowledge of, the importance of their inherent data.
-
-%\hh{I like this idea of machine learning in storage system.
-A lot of analysis has been done on relation of I/O pattern and storage performance.
-But how to know the I/O pattern, data pattern in advance?
-Machine Learning can be used to find pattern.
-These patterns can then be used to schedule storage resources, data placement, data staging/migration, etc.
-These pattern can be gradually learned from running cycles of workflows over time, such that policies can evolve to have better performance when more and more cycles of workflow complete.
-}
-
-CYLC_SUITE_SHARE_DIR
-Typically the data generated by cylc workflows is stored in the “share” directory which is referenced by an environment variable rather than a command:
-https://cylc.github.io/doc/built-sphinx/suite-config.html#suite-share-directories

### 2020-05-22 paper.tex
New additions about Cylc

@@
 The key contributions of this paper are:
-1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of existing tools (Cylc, XIOS, DDN IME, and ESDM).
+1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of existing tools (Cylc, XIOS, DDN's IME, and ESDM).
 
@@
 Burst buffer solutions provide a tiered storage system that aims to exploit a storage hierarchy.
-They can be integrated into hardware capabilities such as DDN IME~\cite{BODIAIFSFI19} or simple software solutions.
+They can be integrated into hardware capabilities such as DDN's Infinite Memory Engine (IME)~\cite{BODIAIFSFI19} or simple software solutions.
 A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
@@
 \begin{description}[itemsep=0pt]
+
   \item[Name] A basic name for the field/data generated. It is extended by a pattern defined in a variable (Lines: 8, 16, 23).
@@
 Components of it are involved in different steps of the workflow and the I/O path.
+While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
+There are two main requirements from Cylc, both covered by Cylc normal functionalities:
 
-While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
+\begin{description}
+
+\item[The ability to dynamically set the job (Slurm) directives for a task]
+
+\ \\ \vspace{-0.5cm}
+
+This will be achieved by calling an external command (run on the Cylc scheduler host) which defines directives to be used by the job.
+This external command will determine some attributes of previous tasks through simple SQL queries to the Cylc database.
+We plan on using Cylc broadcast to alter the directives used by a task by running an extra local task prior to any task where we need to be able to alter the directives.
+
+\item[The ability to dynamically set storage locations]
+
+\ \\ \vspace{-0.5cm}
+
+This will be achieved by defining environment variables in the job script which are set to the output of another external command.
+This external command (run on the job host) will have access to all the normal Cylc environment variables providing details about the current task.
+
+\end{description}
+
+We plan on utilising DDN's IME API to pin data in IME and to trigger migrations between IME and a storage backend explicitly.
 Decisions about data locality will not be made for a whole (and potentially big) workflow.
 Instead, the system will make decisions by looking ahead to several steps of the workflow, allowing reacting to the observed dynamics of the execution.
-We utilise DDN's IME API to pin data in IME and to trigger migrations between IME and a storage backend explicitly.
-
 Ultimately, when a user-script runs, the information about the intended I/O schedule is communicated from EIOS through a modified filename, which is then used by the ESDM-aware application to determine the data placement.

### 2020-05-22 paper.tex
Changed title, added one sentence about performance model.

@@
 
-\title{Potential of I/O-Aware Workflows in Climate and Weather}
+\title{Potential of I/O Aware Workflows in Climate and Weather}
 
@@
 When delivered, this work will increase the opportunity for smarter scheduling of computing by considering storage in heterogeneous storage systems.
+We illustrate the performance-impact on an example workload using a model built on measured performance data at DKRZ.
 
-\keywords{workflow, climate, weather, heterogeneous storage, data-driven}
+\keywords{workflow, heterogeneous storage, data-driven, climate/weather}
 \end{abstract}

### 2020-05-22 paper.tex
Figure changes / grayscaling for "irrelevant" elements.

@@
     keywordstyle=\color{pblue},
-    stringstyle=\color{pred},
+    stringstyle=\bfseries,
     moredelim=[il][\textcolor{pgrey}]{\[\]},
@@
 
-\fig{width=0.8\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file}
+\fig{width=0.8\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file \cite{8675433}}
 
@@
         \includegraphics[width=0.42\textwidth]{pic/layers-xios}
-        \subcaption{I/O path for an MPI-parallel application}\label{pic/layers-xios}
+        \subcaption{I/O path for an MPI application}\label{pic/layers-xios}
     \end{minipage}

### 2020-05-22 paper.tex
Review of Lucy's text.

@@
 While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
-There are two main requirements from Cylc, both covered by Cylc normal functionalities:
+Our design imposes only minor changes to Cylc as the core requirements are covered by normal functionalities:
 
@@
 
-This will be achieved by calling an external command (run on the Cylc scheduler host) which defines directives to be used by the job.
-This external command will determine some attributes of previous tasks through simple SQL queries to the Cylc database.
-We plan on using Cylc broadcast to alter the directives used by a task by running an extra local task prior to any task where we need to be able to alter the directives.
+This will be achieved by calling an external command (run on the Cylc scheduler host) which adds additional directives to be used by the job.
+This command provided by EIOS will determine some attributes of previous tasks through simple SQL queries to the Cylc database.
+We plan on using the Cylc broadcast functionality to change the directives used by a task by running an external program prior to any task where we need to be able to alter the directives.
 
@@
 This will be achieved by defining environment variables in the job script which are set to the output of another external command.
-This external command (run on the job host) will have access to all the normal Cylc environment variables providing details about the current task.
-
+This command provided by EIOS and run on the job host will have access to all the normal Cylc environment variables providing details about the current task.
 \end{description}

### 2020-05-22 plot.R
Data

@@ -1,0 +1,12 @@
+#!/usr/bin/env Rscript
+
+library(ggplot2)
+library(dplyr)
+
+d = read.csv("data.csv")
+# nodes, PPN, type, size, read, write
+d$tpGiB = d$tpMiBs / 1024
+d$nodes = as.factor(d$nodes)
+
+ggplot(d, aes(nodes, tpGiB, color=config, group=config)) + geom_line() + facet_grid(. ~ op, switch = 'both') + xlab("Nodes") + ylab("Performance in GiB/s") + theme(legend.position="bottom") + ylim(0, 250)
+ggsave("data.png", width=6, height=4)

### 2020-05-22 paper.tex
Minor

@@
 While individual components such as ESDM and Cylc exist, we have not implemented the described scheduler, yet.
-To automatically make scheduling decisions, the software stack needs to:
-
+To automatically make scheduling decisions, the software stack needs to: \vspace{0.3cm}
 \begin{enumerate}
-
 \item Deliver information about dataset life cycle together with the workflow, and
-
 \item Adapt the resulting workflow, individual scripts and application executions to consider the potential for data placement.
@@
 This will be achieved by calling an external command (run on the Cylc scheduler host) which adds additional directives to be used by the job.
-This command provided by EIOS will determine some attributes of previous tasks through simple SQL queries to the Cylc database.
+This command, provided by EIOS, will determine some attributes of previous tasks through simple SQL queries to the Cylc database.
 We plan on using the Cylc broadcast functionality to change the directives used by a task by running an external program prior to any task where we need to be able to alter the directives.
@@
 
-This will be achieved by defining environment variables in the job script which are set to the output of another external command.
-This command provided by EIOS and run on the job host will have access to all the normal Cylc environment variables providing details about the current task.
+This will be achieved by defining environment variables in the job script which are set to the output of another external command (run on the job host).
+This command, provided by EIOS, will have access to all the normal Cylc environment variables providing details about the current task.
 \end{description}

### 2020-05-23 paper.tex
Some fixing

@@
 When delivered, this work will increase the opportunity for smarter scheduling of computing by considering storage in heterogeneous storage systems.
-We illustrate the performance-impact on an example workload using a model built on measured performance data at DKRZ.
+We illustrate the performance-impact on an example workload using a model built on measured performance data using ESDM at DKRZ.
 
@@
 
+Climate applications may have complex I/O stacks, as can be seen in \cref{pic/layers-xios}.
+In this case, we assume the application uses XIOS~\cite{xios}, which is providing domain-specific semantics to climate and weather.
+It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF4~\cite{netcdf} to store data as a file.
+Under the hood, NetCDF4 uses the HDF5 API with its file format.
+Internally, HDF5 uses MPI and its data types to specify the nature of the data stored.
+Finally, data is stored on a parallel file system like Lustre which, on the server-side, stores data in a local file system on block devices such as SSDs and HDDs.
+
 \begin{figure}[b]
@@
 
-Climate applications may have complex I/O stacks, as can be seen in \cref{pic/layers-xios}.
-In this case, we assume the application uses XIOS~\cite{xios}, which is providing domain-specific semantics to climate and weather.
-It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF4~\cite{netcdf} to store data as a file.
-Under the hood, NetCDF4 uses the HDF5 API with its file format.
-Internally, HDF5 uses MPI and its data types to specify the nature of the data stored.
-Finally, data is stored on a parallel file system like Lustre which, on the server-side, stores data in a local file system on block devices such as SSDs and HDDs.
-
 Different applications involved in a workflow may use different I/O stacks to store their outputs.
@@
 
-\subsection{Data Center Infrastructure}
+\subsection{Data Centre Infrastructure}
 
@@
 Next, we outline the two core strategies and potential the proposed vision can bring to improve current workflows.
-In the design proposed in this work (\cref{sec:design}), we will focus on the data placement strategy.
+In the design proposed in this work (\Cref{sec:design}), we will focus on the data placement strategy.
 
-\paragraph{Strategy: Data Placement.}
+\paragraph{Strategy: Data Placement}
 
@@
 
-\paragraph{Strategy: Data Reduction.}
+\paragraph{Strategy: Data Reduction}
 
@@
 This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.
-In the example presented in \Cref{lst:esdm.conf}, we have three storage targets: two global accessible file systems (\texttt{lustre01} and \texttt{lustre02}), and one local file system in \texttt{/tmp} that can be accessed via the POSIX backend.
+In the example presented in \cref{lst:esdm.conf}, we have three storage targets: two global accessible file systems (\texttt{lustre01} and \texttt{lustre02}), and one local file system in \texttt{/tmp} that can be accessed via the POSIX backend.
 Each of them comes with a lightweight performance model and the maximum size of data fragments.
@@
 The user now has to provide information about datasets required for input and the generated output for each workflow task in a separate file similarly to Cylc's workflow configuration file.
-An example of an I/O-workflow configuration file (\texttt{io.cylc}) is shown in \Cref{lst:cylc}.
+An example of an I/O-workflow configuration file (\texttt{io.cylc}) is shown in \cref{lst:cylc}.
 In this file, information about Task\,1 is given by example, and we expect the extra information about all tasks in the same file.
@@
 These goals will be achieved via the proposed I/O-aware scheduler, called here EIOS (ESDM I/O Scheduler).
-EIOS will make the schedule considering Cylc workflow, and ESDM provided system characteristics.
-Components of it are involved in different steps of the workflow and the I/O path.
+EIOS will make the schedule considering Cylc workflow and ESDM provided system characteristics.
+We are working together with Cylc Team in developing how EIOS interfaces with Cylc.
+% Components of it are involved in different steps of the workflow and the I/O path.
 While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
@@
 This will be achieved by defining environment variables in the job script which are set to the output of another external command (run on the job host).
-This command, provided by EIOS, will have access to all the normal Cylc environment variables providing details about the current task.
+This command, also provided by EIOS, will have access to all the normal Cylc environment variables with details about the current task.
 \end{description}
@@
 The steps to execute a workflow enriched with I/O information and perform smarter scheduling is depicted in \cref{pic/stages-io}.
+Components of EIOS are involved in different steps of the workflow and the I/O path.
 The suggested alterations can be seen in boxes pointed by red arrows, and the remaining components are the current state-of-the-art for workflows in climate and weather from \cref{pic/stages}.

### 2020-05-24 paper.tex
Ported the text into the paper.

@@
 Our design, based on existing components in climate and weather, is described in \Cref{sec:design}.
+An example use case demonstrates the impact on running a workload at the Mistral supercomputer in \Cref{sec:evaluation}.
 The paper is concluded in \Cref{sec:conclusions}.
@@
 
-\fig{width=0.7\columnwidth}{pic/stages-io}{Software stack and stages of execution with the I/O-aware scheduler (EIOS)}
+\fig{width=0.7\columnwidth}{pic/stages-io}{Software stack and stages of execution with the I/O-aware scheduler (EIOS).
+The red arrows and boxes indicate additions to the workflow compared to \cref{pic/stages}.}
 
@@
 
+\section{Potential Benefit}
+\label{sec:evaluation}
+
+In this section, we discuss the potential benefit the scheduler may have considering DKRZ Mistral supercomputer , ESDM current version and a hypothetical workload related to \cref{pic/cylc1}.
+
+\fig{width=0.6\columnwidth}{pic/data}{Lustre performance for 100, 200, and 500 nodes}
+
+\cref{pic/data} shows the read/write performance when using ESDM to store a time series of 10 steps of a variable with 200k*200k dimension (about 1km global resolution of the model, equivalent to 3 TB of data in total).
+It is stored on either Lustre02 or both Lustre file systems -- ESDM will split the data of a single variable internally and distribute it across file systems.
+While the mapping is not yet optimal, it shows that write performance benefits from this approach.
+Note that each Lustre file system has 128 OSTs (250 total) which means that performance will not improve beyond 500 nodes.
+
+In our scenario, we compare the usage of the node-local file system with a globally shared Lustre file system to store intermediate data.
+When the subsequent job requires the whole data to generate the data product, it could be run on the same nodes.
+
+Each local SSD of Mistral is a Micron M600 MTFD (256 GB) which has a nominal sequential read/write performance slightly above 500 MB/s.
+Hence, with 500 nodes, we could achieve 250 GB/s which surpasses the Lustre performance observed in the figure.
+Another benefit of using local storage is that the interference with IO activities of other jobs would be minimized.
+Since DKRZ has about 3,000 nodes, using the local SSD would sum up to 1.5 TB/s.
+For model runs with 1km resolution, such large runs would be plausible.
+
+Even more, for 500 nodes, our 3 TB of data would be about 6 GB per node, which could be cached in memory and overlap with the computation phases.
+Actually, for all reasonable sizes of the experimental data with 500+ nodes, the observed performance of node-local storage would be higher.
+It might be suitable to couple the model with a parallel analysis process directly, in which case an in-memory file system tmpfs could be used.
+In this case, the performance per node can be assumed to be 3 GiB/s making it a viable option for smaller runs.
+
+Users could make the decision to use the node-local storage manually and create specific run-scripts and create a similar behavior utilizing the local storage.
+However, this would be tedious and error-prone.
+The purpose of our proposal is to establish an abstraction layer to allow for semi-automatic decision making and remove the manual intervention.
+
 \section*{Conclusions}

### 2020-05-24 paper.tex
Fin.

@@
 
-In this section, we discuss the potential benefit the scheduler may have considering DKRZ Mistral supercomputer , ESDM current version and a hypothetical workload related to \cref{pic/cylc1}.
+In this section, we discuss the potential benefit the scheduler may have considering DKRZ Mistral supercomputer, ESDM current version and a hypothetical workload related to \cref{pic/cylc1}.
+We focus on the model execution and the subsequent verification and postprocessing steps.
 
@@
 \cref{pic/data} shows the read/write performance when using ESDM to store a time series of 10 steps of a variable with 200k*200k dimension (about 1km global resolution of the model, equivalent to 3 TB of data in total).
+Note while we only consider the volume data for one variable with one level and timstep, it could be multiplied with a sensible number of levels and timesteps of the model.
 It is stored on either Lustre02 or both Lustre file systems -- ESDM will split the data of a single variable internally and distribute it across file systems.
@@
 In our scenario, we compare the usage of the node-local file system with a globally shared Lustre file system to store intermediate data.
-When the subsequent job requires the whole data to generate the data product, it could be run on the same nodes.
+When the subsequent (verification and postprocessing) job requires the whole data to generate the data product, it could be run on the same nodes.
 
@@
 Since DKRZ has about 3,000 nodes, using the local SSD would sum up to 1.5 TB/s.
-For model runs with 1km resolution, such large runs would be plausible.
+For model runs with 1km resolution, such big configurations would be plausible.
 
 Even more, for 500 nodes, our 3 TB of data would be about 6 GB per node, which could be cached in memory and overlap with the computation phases.
-Actually, for all reasonable sizes of the experimental data with 500+ nodes, the observed performance of node-local storage would be higher.
-It might be suitable to couple the model with a parallel analysis process directly, in which case an in-memory file system tmpfs could be used.
+Actually, for all reasonable sizes of the experimental data with 500+ nodes, the observed performance of node-local storage would be higher, and thus, improving the execution time of such a model.
+It might also be suitable to couple the model with a parallel analysis process directly, in which case an in-memory file system tmpfs could be used.
 In this case, the performance per node can be assumed to be 3 GiB/s making it a viable option for smaller runs.

### 2020-05-25 paper.tex
Final revision

--- 2020-05-23 paper.tex
@@
 \lstset{
-    basicstyle=\ttfamily\footnotesize,
+    basicstyle=\ttfamily\scriptsize,
     numberstyle=\ttfamily\tiny,
@@
 \begin{abstract}
-The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are key for productivity in scientific computing.
+The efficient, convenient, and robust execution of data-driven workflows and enhanced data management are essential for productivity in scientific computing.
 In HPC, the concerns of storage and computing are traditionally separated and optimised independently from each other and the needs of the end-to-end user.
 However, in complex workflows, this is becoming problematic.
-These problems are particularly acute in climate and weather workflows, which as well as become increasingly complex and exploiting deep storage hierarchies, can involve multiple data centres.
+These problems are particularly acute in climate and weather workflows, which as well as becoming increasingly complex and exploiting deep storage hierarchies, can involve multiple data centres.
 
 The key contributions of this paper are:
-1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow seamless integration into current climate and weather workflows as it utilises versions of existing tools (Cylc, XIOS, DDN's IME, and ESDM).
+1) A sketch of a vision for an integrated data-driven approach, with a discussion of the associated challenges and implications, and 2) An architecture and roadmap consistent with this vision that would allow a seamless integration into current climate and weather workflows as it utilises versions of existing tools (ESDM, Cylc, XIOS, and DDN's IME).
 
-The vision proposed here is built on the belief that workflows composed of data, computing, and communication-intensive tasks should drive interfaces and hardware configurations to best support the programming models.
+The vision proposed here is built on the belief that workflows composed of data, computing, and communication-intensive tasks should drive interfaces and hardware configurations to better support the programming models.
 When delivered, this work will increase the opportunity for smarter scheduling of computing by considering storage in heterogeneous storage systems.
@@
 While the current advances correspond to a big leap forward, many processes still require experts.
-For example, porting a workflow from one system to another still requires adjusting runtime parameters of applications and deciding on how data is managed.
+For example, porting a workflow from one system to another requires adjusting runtime parameters of applications and deciding on how data is managed.
 
@@
 For instance, a data-driven workflow may benefit from the explicit and simultaneous use of a locally heterogeneous set of computing and storage technologies.
-This means that substantial changes may be required to a workflow to tailor it to a particular supercomputer environment in order to obtain the best performance.
+This aspect means that substantial changes may be required to a workflow to tailor it to a particular supercomputer environment in order to obtain the best performance.
 
 Knowing the capabilities, interfaces, and performance characteristics of individual components are mandatory to make the best use of them.
-As the complexity of systems expands and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult even for experts to manually optimise the usage of resources in workflows.
-In many cases, modifications are not performed because: 1) They are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) It is a one-time explorative workflow and, 3) Users are not aware of the potential of the complex system.
+As the complexity of systems expands and alternative storage and computing technologies provide unique characteristics, it becomes increasingly difficult, even for experts, to manually optimise the usage of resources in workflows.
+In many cases, modifications are not performed because: 1) They are labour intense: any change to the workflow requires careful validation which may not pay off for small scale runs; 2) It is a one-time explorative workflow and; 3) Users are not aware of the potential of the complex system.
 
@@
 First, we describe the software stack involved in executing workflows in climate and weather in \Cref{sec:workflows}.
-Related work in heterogeneous storage environments and solutions for workflow processing are presented in \Cref{sec:art}.
+Related work in heterogeneous storage environments and solutions for workflow processing is presented in \Cref{sec:art}.
 Next, the vision for including knowledge about data requirements and characteristics is sketched in \Cref{sec:vision} outlining the potential benefit the automatic exploitation might bring.
 Our design, based on existing components in climate and weather, is described in \Cref{sec:design}.
+An example use case demonstrates the impact on running a workload at the Mistral supercomputer in \Cref{sec:evaluation}.
 The paper is concluded in \Cref{sec:conclusions}.
@@
 
-In this section, we describe how workflows are executed in a typical software stack and the typical hardware and software environment involved in running an application.
+In this section, we describe how workflows are executed in a conventional software stack and the typical hardware and software environment involved in running a climate and weather application.
 
@@
 
-Cylc~\cite{8675433} is in charge of executing and monitoring cyclic workflows in which each step is submitted to the batch scheduler of a data centre.
-With Cylc, tasks from multiple cycles may be able to run concurrently without violating dependencies, preventing the issue of delays that cause one cycle to run into another.
+Cylc~\cite{8675433} is a general-purpose workflow engine in charge of executing and monitoring cyclic workflows in which each step is submitted to the batch scheduler of a data centre.
+With Cylc, tasks from multiple cycles may be able to run concurrently without violating dependencies and preventing the issue of delays that cause one cycle to run into another.
 Cylc was written in Python and built around a new scheduling algorithm that can manage infinite workflows of cycling tasks without a sequential cycle loop.
@@
 The information Cylc uses to control a given workflow is the task dependency.
-In a script file, the developers define, for each task, the parallelism settings and where the data is to be stored.
+In a script file, the developers define, for each task, the parallelism settings and where data is to be stored.
 
@@
 In climate research, this process is repeated in the model to simulate years into the future.
-Once the simulation of any month is computed, the data for this month becomes available and can now be analysed.
+Once the simulation of any month is computed, data for this month becomes available and can now be analysed.
 In this workflow, the task \texttt{model} is followed by tasks postprocessing (\texttt{post}), forecast verification (\texttt{ver}), and product generation (\texttt{prod}), all specified as a workflow in a Cylc configuration file (\texttt{flow.cylc}).
@@
 The software stack involved in a general workflow is depicted in \cref{pic/stages}.
-In the following, each stage of the execution is further described.
+Next, each stage of the execution is further described.
 
@@
 
-  \item \textbf{Cylc} parses the workflow configuration file, generates tasks dependencies, defines a schedule for the execution, and monitors the progress of the workflow.
+  \item \textbf{Cylc} parses the workflow configuration file (\texttt{flow.cylc}), generates tasks dependencies, defines a schedule for the execution, and monitors the progress of the workflow.
   Once a task can be executed (dependencies are fulfilled), the workflow engine submits a \textit{job script} for the workload manager with the required metadata that will run the Cylc task script.
@@
 
-  \item \textbf{Job} provides the environment with the resources and it runs the user-provided program or script on one of the nodes allocated for it.
-  Local variables contain information about the environment of the batch job, e.g., the compute nodes allocated, and they enact the Cylc provided script on the node.
+  \item \textbf{Job} provides the environment with the resources and runs the user-provided program or script on one of the nodes allocated for it.
+  Local variables containing information about the environment of the batch job, e.g., the compute nodes allocated, enact the Cylc provided script on the node.
 
@@
   During the creation of the script, Cylc has included variables that describe the task in the workflow.
-  The information is typically fed into the application(s) representing the task, and so defines the storage location.
+  The information is typically fed into the application(s) representing the task and defining the storage location.
   The script uses commands to generate filenames considering the cycle and may store data in a workflow-specific shared directory.
-  Either these commands are set in the Cylc workflow and then injected as environment variables or directly utilised as part of a user-provided script.
+  Either these commands are set in the Cylc workflow and then injected as environment variables or directly utilised as a part of the user-provided script.
 
@@
 In this case, we assume the application uses XIOS~\cite{xios}, which is providing domain-specific semantics to climate and weather.
-It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF4~\cite{netcdf} to store data as a file.
-Under the hood, NetCDF4 uses the HDF5 API with its file format.
-Internally, HDF5 uses MPI and its data types to specify the nature of the data stored.
+It may gather data from individual fields distributed across the machine (exploiting MPI for parallelism) and then uses NetCDF~\cite{netcdf} to store data as a file.
+Under the hood, NetCDF uses the HDF5 API with its file format.
+Internally, HDF5 uses MPI and its data types to specify the nature of data stored.
 Finally, data is stored on a parallel file system like Lustre which, on the server-side, stores data in a local file system on block devices such as SSDs and HDDs.
@@
 Naturally, the application which uses previously generated data as its inputs must use a compatible API to read the specific data format.
-In \cref{pic/layers-xios}, for example, XIOS may perform parallel I/O via the NetCDF4 API, allowing subsequent processes to read data directly using NetCDF4.
-Within the ESiWACE project\footnote{\url{https://www.esiwace.eu/}}, we are developing the Earth System Data Middleware (ESDM)\cite{esdm} to allow applications with this kind of software stack to exploit heterogeneous storage resources in a data centre.
-The goal of ESDM is to provide parallel I/O for parallel applications, with advanced features to optimise subsequent read accesses.
-Implemented with a standalone API, it also provides NetCDF integration allowing usage in existing applications.
+In \cref{pic/layers-xios}, for example, XIOS may perform parallel I/O via the NetCDF API, allowing subsequent processes to read data directly using NetCDF.
+Within the ESiWACE project\footnote{\url{https://www.esiwace.eu/}}, we are developing the Earth System Data Middleware (ESDM)\cite{esdm} to allow applications with this kind of software stack to exploit heterogeneous storage resources in data centres.
+The goal of ESDM is to provide parallel I/O for parallel applications with advanced features to optimise subsequent read accesses.
+Implemented as a standalone API, it also provides NetCDF integration allowing its usage in existing applications.
 Hence, in \cref{pic/layers-xios}, the HDF5 layer can be replaced with ESDM.
@@
 
-Data centres are providing an infrastructure consisting of computing and storage devices with different characteristics, making them more efficient for specific tasks and satisfying the needs of different workflows.
+At present-day, data centres provide an infrastructure consisting of computing and storage devices with different characteristics, making them more efficient for specific tasks and satisfying the needs of different workflows.
 Take, for example, the supercomputer Mistral at DKRZ, that consists of 3,321 nodes\footnote{\url{https://www.dkrz.de/up/systems/mistral}} and offers two types of compute nodes equipped with different CPUs and GPU nodes.
 Each node has an SSD for local storage, and DKRZ has additionally two shared Lustre file systems with different performance characteristics.
-Individual users and projects are mapped to one file system explicitly, and users can access it with work or scratch semantics.
-While data is kept on the work file system indefinitely, available space is limited by a quota.
-The scratch file system allows storing more data, but data is automatically purged after some time.
+Individual users and projects are mapped to one file system explicitly, and users can access it with \texttt{work} or \texttt{scratch} semantics.
+While data is kept on the \texttt{work} file system indefinitely, available space is limited by a quota.
+The \texttt{scratch} file system allows storing additional data, but data is automatically purged after some time.
 
 Future centres are expected to have even more heterogeneity.
-A variety of accelerators (GPU, TPU, FPGAs), active storage, in-memory, and in-network computing technologies will provide storage and processing capabilities.
+A variety of accelerators (GPU, TPU, FPGAs), active storage, in-memory, and in-network computing technologies will provide further storage and processing capabilities.
 \cref{pic/system} shows such a system with a focus on computation and storage.
-Some of these technologies might be local to specific compute nodes or globally available.
-Depending on the need, the storage characteristics range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and cheap storage for long-term archival (tape).
+Some of these technologies might be locally (specific compute nodes) or globally available.
+Depending on the need, the storage characteristics range from predictable low-latency (in-memory storage, NVMe) to online storage (SSD, HDD), and also cheap storage for long-term archival (tape).
 The tasks within any given workflow could benefit from utilising different combinations of storage and computing infrastructure.
@@
 
-Usually, the scripts representing tasks define how the data is placed on the available storage system.
-What happens in many current workflows is that they ignore the benefits of using multiple file systems concurrently and the data locality between tasks to colocating them.
-On top of that, in the current state-of-the-art, scientists optimise the available storage resources intuitively and compile the information about this decision-making process manually.
+Usually, the scripts representing tasks define how data is placed on the available storage system.
+What happens in many current workflows is that they ignore the benefits of using multiple file systems concurrently and data locality between tasks to colocating them.
+On top of that, in the current state-of-the-art scientists optimise the available storage resources intuitively and compile the information about this decision-making process manually.
 
-If a user knows the workflow and the system characteristics, s/he can optimise data placement decisions.
-Consider, for instance, the situation where each computing node has access to three file systems: a fast \texttt{scratch} file system on which data may reside only for a week, a slower \texttt{work} file system, and a \texttt{local} file system.
+If a user knows the workflow and the system characteristics, data placement decisions can be optimised.
+Consider, for instance, the situation where each computing node has access to three file systems: a fast \texttt{scratch} file system on which data may reside only for a week, a slower \texttt{work} file system, and a local file system.
 Most current workflows utilise \texttt{work} and \texttt{scratch} systems.
 When a task is set to run, the corresponding dataset would be moved from \texttt{work} to \texttt{scratch}, processed, and the resulting dataset would be transferred back to \texttt{work}.
-If the \texttt{scratch} filesystem reaches its capacity, the dataset would be moved back to \texttt{work}, and the task would continue running until it is finished, which might be inefficient.
-In this situation, there are many obvious opportunities to utilise data migration to optimise performance, and other criteria (e.g., costs).
+If the \texttt{scratch} file system reaches its capacity, the dataset would be moved back to \texttt{work}, and the task would continue running until it is finished, which might be inefficient.
+In this situation, there are many straightforward opportunities to utilise data migration to optimise performance, and also other criteria (e.g., costs).
 However, with a multitude of file systems that differ at each data centre, such optimisations would be difficult to achieve manually by users.
-Policy-driven systems and burst-buffers perform such optimisations automatically to some extent. However, as they lack information about the workflow, they cannot optimise workflows perfectly.
+Policy-driven systems and burst buffers perform such optimisations automatically to some extent. However, as they lack information about the workflow, they cannot optimise workflows altogether.
 
@@
 
-Related work can be categorised into:
+Related work to the proposed approach can be categorised into:
 1) Technology that exploits heterogeneous storage environments and supports user-directed policies and
@@
 \textbf{Technology.}
-Manual tiering requires the user or application to control the data placement, i.e., storing data typically in the form of files on a particular storage system and, usually, moving data between storage by scripts.
+Manual tiering requires the user or application to control data placement, i.e., storing data typically in the form of files on a particular storage system and, usually, moving data between storage by scripts.
 One limitation of such an approach is that decisions about how data are mapped and packaged into files are made by the producing application, and cannot be changed without manual intervention by a downstream application.
@@
 They can be integrated into hardware capabilities such as DDN's Infinite Memory Engine (IME)~\cite{BODIAIFSFI19} or simple software solutions.
-A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify the data movement for the user, but typically migrates objects in the coarse granularity of files.
+A policy system, e.g., deployed on a burst buffer~\cite{RomanusRP15}, aims to simplify data movement for the user, but typically migrates objects in the coarse granularity of files.
 File systems and data management software such as IBM Spectrum Scale~\cite{SchmuckH02}, HPSS~\cite{528214}, BeeGFS~\cite{beegfs}, and Lustre~\cite{abs-1903-01955} (e.g., using the progressive file layouts feature) provide hierarchical storage management allowing to store data on different storage technology according to administrator-provided policies.
-However, the semantic information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, age of the file, etc.
+However, the semantic information that can be used by this type of system to make decisions is limited, e.g., data location, file extension, file age, etc.
 
@@
 For instance, ADIOS provides in-memory staging that had been exploited by a variety of applications~\cite{slawinska2013maya}.
-Hermes~\cite{kougkas2020acceleration} provides a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
+Hermes~\cite{kougkas2020acceleration} introduces a multi-tiered I/O buffering system with pre-fetcher that provides several data placement policies.
 iRODS~\cite{rajasekar2010irods} is a rule-oriented data system that allows scientists to organise data into shareable collections and provides several patterns for workflows considering data locality and data migration/replication.
@@
 
-\textbf{Workflows. }
+\textbf{Workflows.}
 A good overview of the flavours of Scientific Workflow Management Systems (SWfMS) and their application to data-intensive workflows is given in \cite{liu2015survey}.
-The article states that SWfMS should enable the parallel execution of data-intensive scientific workflows and exploit large amounts of distributed resources.
-Existing solutions recognise challenges in data variety (formats of the input data), opportunities to optimise the schedule by moving code to data, allow specifying the data dependencies for tasks, and they even may consider the capacity of the available data storage.
+The article states that SWfMS should enable the parallel execution of data-intensive scientific workflows and exploit vast amounts of distributed resources.
+Existing solutions recognise challenges in data variety (formats of the input data), opportunities to optimise the schedule by moving code to data, specification of the data dependencies for tasks, and they even may consider the capacity of the available data storage.
 The execution engine Dryad~\cite{isard2007dryad}, for example, allows transferring data between tasks via files or directly using TCP connections and attempts to schedule tasks on the same nodes or racks.
-Swift/T is a scripting language for describing dataflow processing allowing to execute ensembles of applications~\cite{ozik2016desktop}.
+%Swift/T is a scripting language for describing dataflow processing enabling the execution of ensembles of applications~\cite{ozik2016desktop}.
+%Recent improvements aim to migrate data to a local cache allowing to exploit locality.
+%For instance, in \cite{dai2018cross}, information about locality is proposed to be stored in extended attributes.
+%In \cite{TUIBIHWLSC19}, an approach was presented to monitor and analyse I/O behaviour of HPC workflows.
+In \cite{TUIBIHWLSC19}, an approach was presented to monitor and analyse I/O behaviour of HPC workflows.
+Swift/T~\cite{6546066}, a scripting language for describing dataflow processing enabling the execution of ensembles of applications, is now openly used as a prototype platform~\cite{ozik2016desktop}.
 Recent improvements aim to migrate data to a local cache allowing to exploit locality.
 For instance, in \cite{dai2018cross}, information about locality is proposed to be stored in extended attributes.
-In \cite{TUIBIHWLSC19}, an approach was presented to monitor and analyse I/O behaviour of HPC workflows.
 
-Various early research in grid workflows and lately cloud use cases attempts to maximise data locality in that respect.
+Several early research in grid workflows and, lately, cloud workflows, use cases of interest to maximise data locality.
 Economic factors (including storage costs) for workflow execution are discussed in \cite{alkhanak2016cost}.
@@
 
-Workflow systems can also be utilised specifically to reproduce scientific results, i.e., recompute the results.
+Workflow systems can also be specifically utilised to reproduce scientific results, i.e., recompute the results.
 Those scalable workflow solutions typically utilise a container solution to allow execution in an arbitrary software environment.
 Popper~\cite{jimenez2017popper}, Snakemake~\cite{bts480}, and Nextflow~\cite{Nextflow} provide a language to specify workflows and to execute them.
-Snakemake is interesting as it supports to define and infer input and output filenames.
+Snakemake is interesting as it supports definition and inference of input and output filenames.
 
@@
 Nowadays, in order to run a job in an HPC environment efficiently, researchers have to develop profound knowledge, not only about their workflow, which is expected, but also about decisions regarding storage, communication, computing, and considerations regarding cost-efficiency of those operations.
-However, applied scientists should not spend much time understanding hardware characteristics and operative knowledge of running a data centre, but using their knowledge to develop their work and just collect and analyse the results.
+However, applied scientists should not spend much time understanding hardware characteristics and operational knowledge of running a data centre, but using their expertise to develop their work and just collect and analyse the results of their experiments.
 
@@
 Our vision for I/O-aware workflows requires two additional pieces of information.
-Firstly, the user must augment the workflow description with information about I/O requirements and explicitly annotate dependencies to datasets.
+Firstly, the user must augment the workflow description with information about I/O specifications and explicitly annotate dependencies to datasets.
 Secondly, details about the storage architecture must be available.
@@
 
-While many optimisations are possible once an abstraction is in place, the improvements we discuss here are related to the life cycle of datasets and the placement of such datasets into specific storage according to system performance characteristics and the workflow specification.
-To achieve that, the system information shall comprise of all available storage systems, the system topology, and details of each of the required components.
+While many optimisations are possible once an abstraction is in place, the improvements we discuss here are related to the life cycle and placement of datasets into specific storage according to system performance characteristics and workflow specification.
+To achieve that, the system information shall comprise of all available storage systems, the system topology, and details of each available component.
 Simplified and complex models of the components can be included to approximate expected performance for specific I/O patterns.
-It is expected that the data centre (or expert users) can create such a configuration file, e.g., by using vendor-provided information or by executing benchmarks.
+It is expected that the data centre (or expert user) can create such a configuration file, e.g., by using vendor-provided information or by executing benchmarks.
 With this information, a scheduler can make the initial data placement, transformation, and migration decisions for individual datasets during their life cycle.
@@
 
-In general, climate and weather workflows allow specifying tasks and the dependencies among them.
+In general, climate and weather workflows allow specifying tasks and dependencies among them.
 We aim to enhance the current information with characteristics for input and output, i.e., the datasets.
@@
 Round nodes represent tasks, squared nodes represent data, and arrows indicate dependencies.
-In the example, Task\,1 needs two datasets to perform its work, produces Product\,1, and it directly communicates with Task\,2.
+In the example, Task\,1 needs two datasets to perform its work, it produces Product\,1, and directly communicates with Task\,2.
 For each new cycle, the \texttt{checkpoint} from the previous cycle (Product\,1) is used as input to starting the next cycle.
@@
 This last step represents the typical uncertainty of data reuse, i.e., it is unclear how Product\,3 will be used further.
-In the approach proposed in this work, each task is annotated with the required input datasets and the generated products must include metadata such as data life cycle information, the value of data, and how long it should be kept.
+In the approach proposed in this work, each task is annotated with the required input datasets and the generated products must include metadata such as data life cycle, the value of data, and how long it should be kept.
 The idea here is to embrace the concept that tasks dependencies are really imposed by datasets dependencies.
@@
 Value and priority can influence fault-tolerance strategies and imply the quality of service for performance and availability.
-Aspects like data reproducibility (can it be recomputed easily), type of the experiment (test, production), and runtime constraints for the overall and potential workflow could allow reducing costs and, hence, increasing scientific output.
-Next, we outline the two core strategies and potential the proposed vision can bring to improve current workflows.
+Aspects like data reproducibility (can it be recomputed easily), type of the experiment (test, production), and runtime constraints for the overall and potential workflow could allow reducing costs and, hence, increase scientific output.
+Next, we outline two core strategies and the potential the proposed vision can bring to the improvement of current workflows.
 In the design proposed in this work (\Cref{sec:design}), we will focus on the data placement strategy.
@@
 Data placement encompasses all data movement-related activities such as transfer, staging, replication, space allocation and de-allocation, registering and unregistering metadata, locating and retrieving data\footnote{https://www.igi-global.com/dictionary/data-aware-distributed-batch-scheduling/6782}.
-The general idea is to host a dataset on the storage system that is most favourable in terms of performance, cost-effectiveness, availability for the access pattern observed in the workflow.
-Here we are considering the optimisation of data locality, where locality is twofold, spatial and temporal on a variety of characteristics.
-For optimising data placement, there are several approaches:
+The general idea is to host a dataset on the storage system that is most favourable in terms of performance, cost-effectiveness, and availability for the access pattern observed in the workflow.
+Here we are considering the optimisation of data locality, where locality is twofold, spatial and temporal, on a variety of characteristics.
+For optimising data placement, we introduce four approaches: data allocation, data migration, data replication and direct-coupling.
 
@@
 Such configuration is done manually and with restricted information about the system architecture.
-It would be interesting to explore storage options for the datasets and, e.g., having datasets from different cycles placed at different storage systems.
-For instance, in \cref{pic/workflow-s3-n}, alternating the storage location for Product\,2 into two scratch file systems is something that would be a simple job for an I/O-aware scheduler.
-However, currently, that implies providing scripts for that task and all tasks depending on the data with information about the different storage placement.
+It would be interesting to explore storage options for the datasets and, e.g., to have datasets from different cycles placed at different storage systems.
+For instance, in \cref{pic/workflow-s3-n}, alternating the storage location for Product\,2 into two \texttt{scratch} file systems is something that would be a simple job for an I/O-aware scheduler.
+However, currently, it implies providing scripts for that task and all tasks depending on it with information about the different storage placement.
 
 \item[Data Migration] is the process of transferring data from one storage system to another.
-Typically, it involves to delete the data, but this decision can be delayed to provide read access to multiple storage systems.
-Data movement involves a significant overhead as data must be read on one storage and written to another, both in terms of latency and energy-efficient computing, hence needs to be considered carefully.
+Typically, it involves to delete data, but this decision can be delayed to provide read access to multiple storage systems.
+Data movement involves a significant overhead, both in terms of latency and energy-efficient computing, as data must be read on one storage and written to another.
+Hence, it needs to be considered carefully.
 \cref{fig:lifecycle} introduces three possible life cycles for a specific dataset and explains how migrations can be done to improve datasets accessibility.
-In \cref{fig:lifecycle1}, the dataset could be first stored on the \texttt{local} storage to avoid congestion on the \texttt{work} file system, then be migrated to \texttt{work} file system where subsequent tasks of the workflow may read it multiple times.
+In \cref{fig:lifecycle1}, the dataset could be first stored on the local storage to avoid congestion on the \texttt{work} file system, then it is migrated to \texttt{work} file system where subsequent tasks of the workflow may read it multiple times.
 In the end, this dataset may be an intermediate product that can then be deleted.
@@
 However, the last read access must happen before files on \texttt{scratch} are automatically removed.
-Alternatively, \cref{fig:lifecycle3} presents the case where the dataset is created on \texttt{work} by a task and it is copied to a \texttt{local} node.
-This \texttt{local} node allows reading accesses of subsequent tasks which might be beneficial for small random accesses.
+Alternatively, \cref{fig:lifecycle3} presents the case where the dataset is created on \texttt{work} and it is copied to a local node.
+This local node allows reading accesses of subsequent tasks which might be beneficial for small random accesses.
 For the last two scenarios, subsequent tasks would have to be placed on the same node where previous data was stored.
@@
         \includegraphics[width=0.9\columnwidth]{pic/lifecycle-1}
-        \subcaption{Local and work file systems}\label{fig:lifecycle1}
+        \subcaption{Local and \texttt{work} file systems}\label{fig:lifecycle1}
     \end{minipage}
@@
         \includegraphics[width=0.9\columnwidth]{pic/lifecycle-2}
-        \subcaption{Scratch file system only}\label{fig:lifecycle2}
+        \subcaption{\texttt{Scratch} file system only}\label{fig:lifecycle2}
     \end{minipage}
@@
         \includegraphics[width=0.9\columnwidth]{pic/lifecycle-3}
-        \subcaption{Local and work file systems}\label{fig:lifecycle3}
+        \subcaption{Local and \texttt{work} file systems}\label{fig:lifecycle3}
     \end{minipage}
@@
 Data might be replicated by enabling the system to rerun parts of the workflow in case of a data loss.
-In addition, the system may combine the replication of data by \textbf{transforming} the data into a different representation allowing to achieve better performance for a variety of access patterns.
+In addition, the system may combine the replication of data by \textbf{transforming} data into a different representation allowing to achieve better performance considering a variety of access patterns.
 
@@
 As an example, in \cref{pic/workflow-s3-n}, the outcome of Task\,1 may be used directly by Task\,2.
-To achieve some level of independence between producer and consumer, data may also be kept in memory and cached.
+Data may also be kept in memory and cached to achieve a certain level of independence between producer and consumer.
+
 \end{description}
@@
 
-Data reduction reduces the amount of data stored.
-We discuss here two potential optimisations: \textbf{data compression}, and \textbf{data recomputation}.
+Data reduction decreases the amount of data stored.
+We discuss here two potential optimisations: data compression and data recomputation.
 
@@
 \item[Data Compression] is the process of encoding information using fewer bits than the original representation.
-Knowing the characteristics of data production and usage makes it simpler for scientists to annotate the required precision of data in those workflows.
+Knowing the characteristics of data production and usage allows scientists to annotate the required precision of data in those workflows.
 The storage system can exploit such information by reducing the precision of data and automatically picking an appropriate compression algorithm.
@@
 If, for instance, $P_3 C_1$ is a large dataset, \texttt{checkpoint} is small, and computing time is short, it is easy to see that deleting and recomputing it may improve the costs of running the workflow.
-This is just an example, and, currently, scientists perform those optimisations manually.
+That is just an example, and, currently, scientists perform those optimisations manually.
 \end{description}
@@
 \item[Abstraction] Providing the abstraction that enables a separation of concerns.
-Once the IO characteristics for a workflow is defined, the user does not have to know the architecture of the target system on which the workflow will run.
-Thus, removing the specialist from the decision-making process of individual workflows.
+Once the I/O characteristics of a workflow are defined, the user does not have to know the architecture of the target system on which the workflow will run.
+Thus, this level of abstraction can remove the specialist from the decision-making process of individual workflows.
 
-\item[Optimisation] The workflow will be optimised specifically for the available system infrastructure and extra information about the data.
-In particular, by exposing the heterogeneous architecture, potentially runtime characteristics can be considered.
+\item[Optimisation] The workflow will be optimised specifically for the available system infrastructure and information about data.
+In particular, by exposing the heterogeneous architecture, potential runtime characteristics can be considered.
 By using information about the value of data, policies for data management (storage resilience,  recomputation, replication, etc.) can be decided.
@@
 \item Deliver information about dataset life cycle together with the workflow, and
-\item Adapt the resulting workflow, individual scripts and application executions to consider the potential for data placement.
+\item Adapt the resulting workflow, individual scripts, and application executions to consider the potential for data placement strategies.
 \end{enumerate}
@@
 
-ESDM is used as I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
+ESDM is used here as the I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
 This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.
@@
 Each of them comes with a lightweight performance model and the maximum size of data fragments.
-The metadata section (Line 24) utilises here a POSIX interface to store the information about the ESDM objects.
-Internally, ESDM creates so-called containers and dataset objects to manage data fragments.
+The metadata section (Line 24) utilises here a POSIX interface to store the information about ESDM objects. Internally, ESDM creates so-called containers and dataset objects to manage data fragments.
 
@@
 "metadata": {"type": "POSIX",
-        "id": "md",
-        "target": "./metadata",
+        "id" : "md",
+        "target" : "./metadata",
         "accessibility" : "global"
@@
 
-ESDM manages a pool of threads that should be created per compute node to achieve good performance and delegates the assignment of optimal block sizes to the storage backend.
-The number of threads is defined in the configuration file.
-As an example, based on the number of Object Storage Targets (OSTs) provided by both Lustre systems at DKRZ, performance tests already developed using ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
-To clarify the behaviour, ESDM distributes a single dataset across multiple storage devices depending on their characteristics.
-Since ESDM also supports several (non-POSIX) storage backends, an application can utilise all available storage systems without any modifications to the code.
-
-The configuration file is inquired by an application that utilises ESDM and steers the distribution of data during I/O.
-While the current system information and performance model are based on latency and throughput only, it shows that automatic decision making can be made on behalf of the user.
+ESDM manages a pool of threads that should be created per compute node to achieve better performance and delegates the assignment of optimal block sizes to the storage backend.
+Since ESDM supports several (non-POSIX) storage backends, an application can utilise all available storage systems without any modifications to the code.
+%The maximum number of threads is defined in the configuration file.
+The configuration file is inquired by the application utilising ESDM and steers the distribution of data during I/O.
+To elucidate the system's behaviour, ESDM distributes a single dataset across multiple storage devices depending on their characteristics.
+While the current system information and performance model are based on latency and throughput only, ESDM shows that automatic decision making can be made on behalf of the user.
+%As an example, based on the number of Object Storage Targets (OSTs) provided by both Lustre systems at DKRZ, performance tests already developed using ESDM~\cite{2019_3361225} shows that no more than 200 threads in total should be used to perform I/O to extract the best performance.
 
@@
 
-The user now has to provide information about datasets required for input and the generated output for each workflow task in a separate file similarly to Cylc's workflow configuration file.
-An example of an I/O-workflow configuration file (\texttt{io.cylc}) is shown in \cref{lst:cylc}.
-In this file, information about Task\,1 is given by example, and we expect the extra information about all tasks in the same file.
-Ultimately, this could be integrated into the workflow specification file of Cylc.
+The user now has to provide information about the datasets required as input and the generated output for each workflow task in a file called I/O-workflow configuration file (\texttt{io.cylc}).
+An example of an \texttt{io.cylc} file is shown in \cref{lst:cylc}.
+In this file, information about Task\,1 is given as an example, and we expect the extra information about all tasks in the same file.
+This file could define a cycle flexibly to be a month or a year according to the file \texttt{flow.cylc}.
+The notation is similar to the specification of Cylc workflows using a nested INI format and, ultimately, files \texttt{io.cylc} and \texttt{flow.cylc} can be merged.
+
+For each task, inputs and outputs are defined.
+In the \textbf{input} section, each entry specifies the virtual name that is used by ESDM as a filename inside NetCDF.
+Line~5, for example, it defines that the filename \texttt{topography} is mapped to a specific input file and that this dataset does not depend on any previous step of the workflow.
+The next line specifies that the input filename \texttt{checkpoint} should be mapped to the output of Task~1 \texttt{checkpoint} dataset from the previous cycle (e.g., the checkpoint generated after completing the last year's production).
+For the initial cycle, the checkpoint file will be empty, and the application will load \texttt{init} data.
+In the \textbf{output} section, the datasets are annotated with their characteristics more precisely.
+For each variable, a pattern defining how frequently data is output according to the workflow must be provided.
+Most data is input and output in the periodicity of the cycle. Still, we can have variables with different patterns, such as \texttt{varA}, which is output per day regardless of the cycle.
 
@@
 [Task 1]
+
   [[inputs]]
+
     topography = "/pool/input/app/config/topography.dat"
@@
   [[outputs]]
+
     [[[varA]]] # This is the name of the variable
@@
 
-In this example, the \texttt{io.cylc} file could define a cycle flexibly to be a month or a year according to the \texttt{flow.cylc} file.
-The notation is similar to the specification of Cylc workflows using a nested INI format.
-For each task, inputs and outputs are defined.
-In the input section, each entry specifies the virtual name that is used by ESDM as a filename inside NetCDF.
-Line~3, for example, defines that the filename \texttt{topography} is mapped to a specific input file.
-This dataset does not depend on any previous step of the workflow.
-The next line specifies that the input filename \texttt{checkpoint} should be mapped to the output of Task~1 checkpoint dataset from the previous cycle (e.g., the checkpoint generated after completed the last year's output).
-For the initial cycle, the checkpoint file will be empty, and the application will load the init data.
-In the output section, the datasets are annotated with their characteristics more precisely.
-For each variable, a pattern defining how frequently the data is output according to the workflow must be provided.
-Most data is input and output in the periodicity of the cycle, except for \texttt{varA}, which is output per day regardless of the cycle.
-Next, we formally define the expected annotations in all the fields expected in the I/O-workflow configuration file:
+Next, we formally define the expected annotations in all the fields envisioned in the I/O-workflow configuration file:
 
@@
 
-  \item[Name] A basic name for the field/data generated. It is extended by a pattern defined in a variable (Lines: 8, 16, 23).
+  \item[Name] A primary name for the field/data generated. It is extended by a pattern defined in a variable (Lines: 11, 19, 26).
 
-  \item[Pattern] The frequency the data is output (Lines: 9, 17).
+  \item[Pattern] The frequency of data output (Lines: 12, 20).
 
-  \item[Lifetime] How long the data must be retained on storage (if at all) (Lines: 10, 18).
+  \item[Lifetime] How long data must be retained on storage (if at all) (Lines: 13, 21).
 
-  \item[Type] The class type of data, i.e., checkpoint, diagnostics, temporary (Lines: 11, 19, 24).
+  \item[Type] The class type of data, i.e., checkpoint, diagnostics, temporary (Lines: 14, 22, 27).
 
-  \item[Datatype] The data type of the data (Lines: 12, 20, 25).
+  \item[Datatype] The data type of data (Lines: 15, 23, 28).
 
-  \item[Size] An estimate of the data size\footnote{This field can be inferred if dimension and datatype are provided.} (Lines: 13, 26).
+  \item[Size] An estimate of data size\footnote{This field can be inferred if dimension and data type are provided.} (Lines: 16, 29).
 
-  \item[Dimension] The data dimension (Line: 21).
+  \item[Dimension] The data dimension (Line: 24).
 
-  \item[Accuracy] Characteristics quantifying the required level of data precision (Line: 14).
+  \item[Accuracy] Characteristics quantifying the required level of data precision (Line: 17).
 
@@
 
-Note that the user may not be able to provide all required information.
-This can be handled by assuming a default safe behaviour.
-For instance, in the case of missing data precision, data should be retained in original form.
+Note that the user may not be able to provide all required information which can be handled by assuming a default safe behaviour.
+For instance, in the case of missing data precision, data should be retained in the original form.
 Knowing the dimension or size a priori might be difficult for scientists, e.g., the log file size is unclear.
-In this case, the user may insert relevant information like small or big, reflecting that any information is better than no information at all.
-In future, we will explore how to automatically infer the output volume from the input or by using monitoring.
-By allowing to run using an empty I/O workflow specification and monitoring I/O accesses for one cycle, we can propose an I/O description to the user to simplify the specification.
+In this case, the user may insert relevant information like small or big, indicating that any information is better than no information at all.
+In future, we will explore ways to infer the output volume from the input automatically.
+For instance, by running the workflow without I/O specification and monitoring I/O accesses for one cycle, we can propose an I/O description to the user to simplify the specification and generate an experimental I/O configuration file.
 
@@
 We are working together with Cylc Team in developing how EIOS interfaces with Cylc.
-% Components of it are involved in different steps of the workflow and the I/O path.
-While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which provide the opportunity for keeping data in local storage.
-Our design imposes only minor changes to Cylc as the core requirements are covered by normal functionalities:
+While Cylc schedules the workflow, EIOS can provide hints about colocating tasks which generate the opportunity for keeping data in local storage.
+Our design imposes only minor changes to Cylc as normal functionalities cover the core requirements:
 
@@
 
-\ \\ \vspace{-0.5cm}
+\ \\ \vspace{-0.6cm}
 
 This will be achieved by calling an external command (run on the Cylc scheduler host) which adds additional directives to be used by the job.
-This command, provided by EIOS, will determine some attributes of previous tasks through simple SQL queries to the Cylc database.
-We plan on using the Cylc broadcast functionality to change the directives used by a task by running an external program prior to any task where we need to be able to alter the directives.
+This command, provided by EIOS, will determine attributes of previous tasks through simple SQL queries to the Cylc database.
+We plan on using the Cylc broadcast functionality to change the instructions used by a task by running an external program before any task where we may want to alter the directives.
 
@@
 
-\ \\ \vspace{-0.5cm}
+\ \\ \vspace{-0.6cm}
 
 This will be achieved by defining environment variables in the job script which are set to the output of another external command (run on the job host).
-This command, also provided by EIOS, will have access to all the normal Cylc environment variables with details about the current task.
+This command, also provided by EIOS, will have access to all the standard Cylc environment variables with details about the current task. \vspace{-0.1cm}
+
 \end{description}
 
-We plan on utilising DDN's IME API to pin data in IME and to trigger migrations between IME and a storage backend explicitly.
+We plan on utilising DDN's IME API to pin data in IME and trigger migrations between IME and a storage backend explicitly.
 Decisions about data locality will not be made for a whole (and potentially big) workflow.
@@
 
-The steps to execute a workflow enriched with I/O information and perform smarter scheduling is depicted in \cref{pic/stages-io}.
+The steps to execute a workflow enriched with I/O information and perform smarter scheduling are depicted in \cref{pic/stages-io}.
 Components of EIOS are involved in different steps of the workflow and the I/O path.
@@
   \item \textbf{Slurm}
-  Cylc may now have added an optimisation identified by EIOS which is now handled by a modified Slurm.
+  Cylc may now have added an optimisation identified by EIOS which is promptly handled by a modified Slurm.
   Also, if migrations have to be performed, Slurm will administer them according to the specification in the job script.
@@
   The application may either use XIOS, NetCDF with ESDM support or ESDM directly to access datasets.
-  Hence, in \cref{pic/layers-xios}, the HDF5 layer is replaced with ESDM.
-  ESDM loads the \texttt{esdm.conf} file that contains the information about the available storage backends and their characteristics.
-  ESDM extracts the long-term schedule information from the generated pseudo filenames and employs it during the I/O scheduling to optimise the storage considering data locality between tasks.
-  Basically, ESDM can now change the priorities for data placement of the different storage locations that would normally be encoded in its configuration file.
+%  Hence, in \cref{pic/layers-xios}, the HDF5 layer is replaced with ESDM.
+  ESDM loads the file \texttt{esdm.conf} that contains the information about the available storage backends and their characteristics.
+  ESDM extracts the long-term schedule information from the generated pseudo filenames and employs them during the I/O scheduling to optimise the storage considering data locality among tasks.
+  Basically, ESDM can now change the priorities for data placement in the different storage locations that would typically be encoded in Cylc's configuration file.
 \end{enumerate}
+
+\section{Potential Benefit}
+\label{sec:evaluation}
+
+In this section, we discuss the potential benefit that our vision for I/O-aware workflows may have considering DKRZ Mistral supercomputer, ESDM current version and a hypothetical workload related to the workflow in \cref{pic/cylc1}.
+In our scenario, we compare the usage of the node-local file system with a globally shared Lustre file system to store intermediate data.
+We focus on the model execution and subsequent verification and postprocessing steps.
+When the subsequent jobs require the whole data to generate the data product, they can be run on the same nodes.
+
+\fig{width=0.6\columnwidth}{pic/data}{Lustre performance for 100, 200, and 500 nodes}
+
+\cref{pic/data} shows the read/write performance when using ESDM to store a time series of 10 steps of a variable with 200k $\times$ 200k dimension (about 1\,km global resolution of the model, equivalent to 3\,TB of data in total).
+Note that, while we only consider the volume data for one variable with one level and one timestep, this value could be multiplied by a sensible number of levels and timesteps of the model.
+Data is stored on either Lustre02 or both Lustre file systems -- ESDM splits data of a single variable internally and distributes them across the file systems.
+While the mapping is not yet optimal, the figure shows that the write performance benefits from this approach.
+% Note that each Lustre file system has 128 Object Storage Targets (OSTs)
+% (250 in total)
+% which means that performance will not improve beyond 500 nodes.
+
+Another improvement can be achieved by using local storage.
+Each local SSD of Mistral is a Micron M600 MTFD (256\,GB), which has a nominal sequential read/write performance slightly above 500\,MB/s.
+Hence, with 500 nodes, we could achieve 250\,GB/s, which surpasses the Lustre performance observed in \cref{pic/data}.
+Even more, our 3\,TB of data would be about 6\,GB per node, which could be cached in memory and overlap with the computation phases.
+An additional benefit of using local storage is that the interference with I/O activities of other jobs would be minimised.
+Actually, for all reasonable sizes of the experimental data with 500+ nodes, the observed performance of node-local storage would be higher, and thus, improving the workflow execution time.
+Since DKRZ has more than 3,000 nodes, using the local SSD would sum up to 1.5\,TB/s.
+For model runs with 1\,km resolution, such configurations would be reasonable.
+It might also be suitable to couple the model with a parallel analysis process directly using an in-memory file system such as \texttt{tmpfs}.
+In this case, the performance per node can be assumed to be 3\,GiB/s, making it a viable option for smaller runs.
+
+Users can always use the node-local storage manually and create specific run-scripts to reproduce similar behaviours utilising the local storage.
+However, this would be tedious and error-prone.
+The purpose of our proposal is to establish an abstraction layer to allow for semi-automatic decision making and reduce, or even remove manual intervention.
 
@@
 
-In the domain of climate and weather, organising the data placement on storage tiers is performed by the users or via policies, often leading to suboptimal decisions.
-Additionally, the manual optimisation and hard-coding of storage locations is non-portable and an error-prone task.
+In the domain of climate and weather, organising data placement on storage tiers is performed by the users or via policies, often leading to suboptimal decisions.
+Additionally, manual optimisation and hard-coding of storage locations are non-portable and an error-prone task.
 We believe users must be able to express their workflows abstractly.
@@
 With knowledge about the data pattern, the runtime system could generate optimised execution plans and monitor their execution.
-In this work, we describe the general vision and a specific design for the software stack in the domain in climate and weather that we work on in the ESiWACE project.
-The proposed changes increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system architecture in a workflow.
+In this work, we describe the overall vision and a specific design for the software stack in the domain of climate and weather that we work on in the ESiWACE project.
+The proposed changes increase the opportunity for smarter scheduling of storage in heterogeneous storage environments by considering the characteristics of data and system architecture in the workflow.
 

### 2020-05-25 paper.tex
Bigger size for fig 1.
Inserting ref for the SSD

@@
 
-\fig{width=0.8\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file \cite{8675433}}
+\fig{width=0.9\columnwidth}{pic/cylc1}{Example of a Cylc workflow with its configuration file \cite{8675433}}
 
@@
 Another improvement can be achieved by using local storage.
-Each local SSD of Mistral is a Micron M600 MTFD (256\,GB), which has a nominal sequential read/write performance slightly above 500\,MB/s.
+Each local SSD of Mistral is a Micron M600 MTFD\footnote{https://www.anandtech.com/show/8528/micron-m600-128gb-256gb-1tb-ssd-review-nda-placeholder} (256\,GB), which has a nominal sequential read/write performance slightly above 500\,MB/s.
 Hence, with 500 nodes, we could achieve 250\,GB/s, which surpasses the Lustre performance observed in \cref{pic/data}.

### 2020-05-25 paper.tex
Incorporated Luciana's comments.

@@
 
-  \item \textbf{Script} executes the commands with one or multiple (potentially parallel) applications to run sequentially.
+  \item \textbf{Script} starts the commands sequentially (a command can be a parallel application).
   During the creation of the script, Cylc has included variables that describe the task in the workflow.
@@
 
-ESDM is used here as the I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
+The system information of the design is realized using already available capabilities in the ESDM middleware.
+We assume ESDM is used as the I/O middleware in the parallel application (with NetCDF or directly) and orchestrates the I/O according to a simplified ESDM configuration file (\texttt{esdm.conf}).
 This file contains information about the available technology in the data centre, its I/O characteristics, and will be used to make decisions about how to prioritise I/O targets.
@@
 \cref{pic/data} shows the read/write performance when using ESDM to store a time series of 10 steps of a variable with 200k $\times$ 200k dimension (about 1\,km global resolution of the model, equivalent to 3\,TB of data in total).
-Note that, while we only consider the volume data for one variable with one level and one timestep, this value could be multiplied by a sensible number of levels and timesteps of the model.
+Note that, while we only consider the volume data for one variable with one level and ten timesteps, this value could be multiplied by a sensible number of levels and timesteps of the model.
 Data is stored on either Lustre02 or both Lustre file systems -- ESDM splits data of a single variable internally and distributes them across the file systems.
 While the mapping is not yet optimal, the figure shows that the write performance benefits from this approach.
-% Note that each Lustre file system has 128 Object Storage Targets (OSTs)
-% (250 in total)
-% which means that performance will not improve beyond 500 nodes.
+In our observations, performance will not improve beyond 500 nodes, which might be due to the fact that each Lustre file system has 128 Object Storage Targets (OSTs) (250 in total).
 

### 2020-05-25 paper.tex
Expanded Sec 5.

@@
 
-In this section, we discuss the potential benefit that our vision for I/O-aware workflows may have considering DKRZ Mistral supercomputer, ESDM current version and a hypothetical workload related to the workflow in \cref{pic/cylc1}.
-In our scenario, we compare the usage of the node-local file system with a globally shared Lustre file system to store intermediate data.
+In this section, we discuss the potential performance benefit that our vision for I/O-aware workflows may have considering DKRZ Mistral supercomputer, ESDM current version and a hypothetical workload related to the workflow in \cref{pic/cylc1}.
+In our scenario, we compare the usage of the node-local file system\footnote{%
+We assume the availability and fault-tolerance of the nodes is non-uncritical for the particular workload -- typically nodes can be repaired and returned to the pool within days.} with a globally shared Lustre file system to store intermediate data.
 We focus on the model execution and subsequent verification and postprocessing steps.
-When the subsequent jobs require the whole data to generate the data product, they can be run on the same nodes.
+Firstly, checkpoints of a long model execution chain could be stored locally and restart from there.
+When the subsequent jobs require the whole data to generate a product, they can be run on the same nodes.
 
@@
 While the mapping is not yet optimal, the figure shows that the write performance benefits from this approach.
-In our observations, performance will not improve beyond 500 nodes, which might be due to the fact that each Lustre file system has 128 Object Storage Targets (OSTs) (250 in total).
+In our observations, performance will not improve beyond 500 nodes, which might be due to the fact that each Lustre file system has 128 Object Storage Targets (256 in total).
 
@@
 Actually, for all reasonable sizes of the experimental data with 500+ nodes, the observed performance of node-local storage would be higher, and thus, improving the workflow execution time.
-Since DKRZ has more than 3,000 nodes, using the local SSD would sum up to 1.5\,TB/s.
+Since DKRZ has more than 3,000 nodes, using the local SSD would sum up to 1.5\,TB/s speeding up the IO phase by 7x.
 For model runs with 1\,km resolution, such configurations would be reasonable.
